### 2023-10-21 Manuscript/Manuscript.qmd
Edits to "the lay of the land"

---
title: "Equity conceptualizations and standards within transportation literature: a scoping review"
bibliography: All_papers_in_extracted.bib
format: docx
---

```{r, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r setup}
library(huxtable)
library(sf)
library(ggplot2)
library(tidyverse)
library(readxl)
library(DiagrammeR)
library(scales)
library(corrplot)
library(patchwork)
library(flextable)
library(ftExtra)
```

```{r PRISMA-flow-chart}
Pflow_chart <- grViz("digraph flowchart {
      node [fontname = Helvetica, shape = rectangle, style = filled, color = grey90]

      tab1 [label = '6832 references imported for screening']
      tab1_1 [label = '611 duplicates removed']
      tab2 [label = '6221 studies screened - title and abstract (step 1)']
      tab2_1 [label = '4511 studies excluded']
      tab3 [label = '1710 studies assessed for full-text eligibility (step 2 and 3)']
      tab3_1 [label = 
      <1545 studies excluded:<br ALIGN = 'LEFT'/> 
      &#8226; 1061 No ES (only EC)<br ALIGN = 'LEFT'/> 
      &#8226; 285 No full-text<br ALIGN = 'LEFT'/> 
      &#8226; 147 No EC and no ES<br ALIGN = 'LEFT'/> 
      &#8226; 38 No EC (only ES)<br ALIGN = 'LEFT'/> 
      &#8226; 13 Other<br ALIGN = 'LEFT'/> 
      &#8226; 1 Wrong study type<br ALIGN = 'LEFT'/> >]
      tab4 [label = 'Data from 165 studies extracted']

      blank1[label = '', width = 0.01, height = 0.01]
      tab1 -> blank1[dir=none];
      blank1 -> tab1_1[minlen=10];
      {{ rank = same; blank1 tab1_1 }}
      blank1 -> tab2

      blank2[label = '', width = 0.01, height = 0.01]
      tab2 -> blank2[dir=none];
      blank2 -> tab2_1[minlen=10];
      {{ rank = same; blank2 tab2_1 }}
      blank2 -> tab3

      blank3[label = '', width = 0.01, height = 0.01]
      tab3 -> blank3[dir=none];
      blank3 -> tab3_1[minlen=10];
      {{ rank = same; blank3 tab3_1 }}
      blank3 -> tab4
      }
      ")
```

```{r, eval=FALSE}
Pflow_chart %>%
  DiagrammeRsvg::export_svg() %>% 
  charToRaw() %>% rsvg::rsvg_png(file="Pflow_chart.png")
```

```{r import-extraction-data}
Consensus_all <- read_excel("ALL_Extraction_Results_Feb22_edited.xlsx") |> filter(`Reviewer Name` == "Consensus")

Consensus_only_analysis <- read_excel("ALL_Extraction_Results_Feb22_edited.xlsx") |> 
  filter(`Reviewer Name` == "Consensus") |> 
  filter(`Included in analysis` == "Yes")
```

```{r by-year-and-continent-plot}
year_cont_plot <- ggplot(data = Consensus_only_analysis,
                         aes(y = `_Publication year`, 
                         fill = `_Case study continent`)) +
  geom_bar(aes(x = (..count..)/sum(..count..)),col="black") + 
  scale_x_continuous(labels=scales::percent,
                     breaks = seq(0,0.2, by=0.05),
                     limits = c(0,0.2)) +
  xlab("% of papers") +
  theme_classic() + 
  theme(text=element_text(size=10),
        axis.title.y = element_blank(),
        legend.position = "bottom") + guides(fill = guide_legend(nrow = 1,
                                                                 title=NULL))
```

```{r modes-by-type-plot}
df_mode <- strsplit(Consensus_only_analysis$`_mode`, "; ", fixed = T)

df_mode <- data.frame(
  id=rep(Consensus_only_analysis$`Covidence #`, lengths(df_mode)),
  mode=unlist(df_mode))

df_mode <- df_mode |> group_by(mode) |> 
  summarize(count = n()) |> 
  mutate(count = count/sum(count))

mode_plot <- ggplot(data = df_mode, aes(x = forcats::fct_reorder(mode,count), y= count))+ 
  geom_bar(stat="identity", fill = "coral2", col = "black") + 
  scale_y_continuous(labels=scales::percent,
                     breaks = seq(0,0.4, by=0.05),
                     limits = c(0,0.4)) +
  coord_flip() +
  labs(y = "% of mentions in papers", subtitle = "Modes") + 
  theme_classic() + 
  theme(text=element_text(size=10),
        axis.title.y = element_blank())
```

```{r population-groups-plot}
df_pop <- strsplit(Consensus_only_analysis$`_population`, "; ", 
                       fixed = T) 
df_pop <- data.frame(
  id=rep(Consensus_only_analysis$`Covidence #`, lengths(df_pop)),
  pop=unlist(df_pop)) |> 
  mutate(pop = ifelse(pop == "Vulnerable populations (composite)", "Comp. vul", pop))

df_pop <- df_pop |> group_by(pop) |> summarize(count = n()) |> 
  mutate(count = count/sum(count))

#adding grouping var so they print on plot in the order we want
df_pop$group <- c(7,9,6,8,2,3,5,10,1,4)

pop_plot <- ggplot(data = df_pop, aes(x = pop, y = count)) +
  geom_bar(stat="identity", fill = c(rep("skyblue2",each=8), "grey", "skyblue2"), col = "black") + 
  scale_y_continuous(labels=scales::percent, limits = c(0,0.25), breaks = seq(0,0.25, by=0.05))+
  scale_x_discrete(limits=(df_pop$pop)[order(df_pop$group)]) +
  coord_flip() +
  labs(y = "% of mentions in papers", subtitle = "Population groups") + 
  theme_classic() + 
  theme(text=element_text(size=10),
        axis.title.y = element_blank())
```

```{r concs}
df_conc <- strsplit(Consensus_only_analysis$`_equity conceptualization`, "; ", 
                                                   fixed = T)

                            df_conc <- data.frame(
                              id=rep(Consensus_only_analysis$`Covidence #`, lengths(df_conc)),
                              conc=unlist(df_conc)) %>% 
                              mutate(conc = ifelse(conc == "Inequitable exposure to externalities", "Ineq. externalities", conc),
                                     conc = ifelse(conc == "Sufficientarism / Capabilities approach", "Emerg. theories", conc),
                                     conc = ifelse(conc == "Transport-related exclusion", "TR exclusion", conc))

df_standard_mcat <- strsplit(Consensus_only_analysis$`_type of equity standard2`, "; ", 
                       fixed = T) 
df_standard_mcat <- data.frame(
  id=rep(Consensus_only_analysis$`Covidence #`, lengths(df_standard_mcat)),
  std_mcat=unlist(df_standard_mcat))

standards_conc <- merge(df_conc, df_standard_mcat, by="id", all = T)

NUM_standards_conc <- standards_conc %>% group_by(id) %>% summarize(count = n())
NUM_ENTRIES_standards_conc <- NUM_standards_conc$count %>% sum()

standards_conc <- merge(standards_conc, NUM_standards_conc, by="id", all = T)

# standards_conc_plot <- ggplot(data = standards_conc,
#                               aes(x = `conc`, fill = `std_mcat`)) +
#   geom_bar() +
#   labs(x = "Conceptualizations (summarized)", fill = "Standards (main categories)") +
#   theme_classic() + 
#   theme(text=element_text(size=10),
#         axis.title.y = element_blank(),
#         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   coord_flip()

# c("Wellbeing","Vertical equity","Utilitarian","TR exclusion", "Spatial equity", "Rights", "Ineq. externalities", "Horizontal equity", "Emerg. theories")
```

```{r, message=FALSE}
standards_conc <- standards_conc %>% 
  group_by(conc, std_mcat) %>% summarize(total = n())

standards_conc <- standards_conc %>%
  mutate(total_perc_in_conc = total/sum(total),
         total_perc_in_all = total/NUM_ENTRIES_standards_conc)
sums_concs <- standards_conc %>% group_by(conc) %>% summarize(
  total_perc_in_all = sum(total_perc_in_all),
  total_perc_in_conc = 1) %>% mutate(std_mcat = "TOTALS")

new_rows <- data.frame(conc = c("Emerg. theories", "Utilitarian", "TOTALS"),
                       total_perc_in_all = c(NA, NA, NA),
                       total_perc_in_conc = c(NA, NA, NA),
                       std_mcat = c("Environmental +","Environmental +", "TOTALS"))

sums_std_mcat <- standards_conc %>% group_by(std_mcat) %>% summarize(total_perc_in_all = sum(total_perc_in_all),
  total_perc_in_conc = 1) %>% mutate(conc = "TOTALS")

totals_standards_conc <- rbind(standards_conc, sums_concs)
totals_standards_conc <- rbind(totals_standards_conc, sums_std_mcat)
totals_standards_conc <- rbind(totals_standards_conc, new_rows)

#adding the grouping numbers, so the order that vars are printed are in geom_tile() are chronological
totals_standards_conc$group <- c(12,13,14,
                    41,42,43,44,
                    31,32,33,34,
                    21,22,23,24,
                    51,52,53,54,
                    71,72,73,74,
                    02,03,04,
                    81,82,83,84,
                    61,62,63,64,
                    15,45,35,25,55,75,05,85,65,
                    91,92,93,94,
                    11,01,95)

#adding long definitions, for table under heatmap

totals_standards_conc_table <- data.frame(
  Conceptualizations = c("The distribution of transport system benefits/costs should favourably prioritize disadvantaged groups.",
                         "Concerning the links between transport system interaction and social exclusions.",
                         "Evaluation of transport system benefits/costs through wellbeing measures (e.g., livability, quality of life, subjective well being).",
                         "Equal distribution of transport system benefits/costs across space.",
                         "Equal distribution of transport system benefits/costs across population."),
  Conceptualizations_cont = c("Concerning the equitable distribution of externalities (e.g., road safety, congestion, pollution).",
                         "Concerning legal/human rights of access (e.g., mobilty rights, rights to the city, discriminatory application of laws).",
                         "E.g., Sufficientarism: minimums informed by individual capabilities. Capabilities approach: the capability of living the life one values (rather than having the right to do so)",
                         "Maximizing benefits/welfare overall, without consideration of distribution.",
                         " "),
  Standards = c("Related to user characteristics (e.g., affordability levels, perceived equity).",
                "Access to destinations indicators (e.g., # of parks within 30-min transit trip), the use of travel impedance thresholds (e.g., 300m to bus stop).",
                "E.g., Level of service, quality, universal design standards.",
                "E.g., noise, air, pollution criteria, safety.",
                " ")
)

```

```{r}
set_flextable_defaults(font.size= 7,
                       font.family = NULL,
                       text.align = 'justify')

totals_conc_table_flex <- 
  # compose(totals_conc_table_flex,
  #         i = 1, j = c(1,2), part = "header", value = as_paragraph(c("Conceptualizations", "Conceptualizations"))) %>%
  prepend_chunks(flextable(totals_standards_conc_table %>% select(-c("Standards"))),
                 part = "body", i = c(1,2,3,4,5), j = 1, as_b(c("Vertical equity: ", "TR exclusion: ", "Wellbeing: ", "Spatial equity: ", "Horizontal equity: "))) %>% 
  prepend_chunks(part = "body", i = c(1,2,3,4), j = 2, as_b(c("Ineq. externalities: ", "Rights: ", "Emerg. theories: ", "Utilitarian: "))) %>% 
  # merge_h(part="header", i=1) %>% #header merge
  merge_at(i=5, j=1:2) %>%
  # align( align="center", part="header") %>%
  delete_part("header") %>%
  border_outer() %>%
  border_inner_h() %>%
  autofit()

totals_standards_table_flex <- 
  # compose(totals_standards_table_flex,
  #         i = 1, j = c(1), part = "header", value = as_paragraph(c("Standards"))) %>% 
  prepend_chunks(flextable(totals_standards_conc_table %>% select("Standards")),
                 part = "body", i = c(1,2,3,4), j = 1, as_b(c("Population: ", "Opportunity: ", "Infrastructure: ", "Environmental +: "))) %>%
  merge_at(i= 4:5, j= 1) %>%
  delete_part("header")%>%
  border_outer() %>%
  border_inner_h() %>%
  autofit()
```

```{r}
standards_conc_percentages_plot <- ggplot(totals_standards_conc, 
       aes(x=std_mcat, y=conc, fill= total_perc_in_all 
           #width = rep(10, each=50)
           ))+ 
  geom_tile() + 
  scale_fill_gradient2(low = "#AFAFED", mid= "#3F3FD4", high = "#191970", midpoint = 0.2, limits = c(0,.4), na.value = "#FCB5AC") +
  scale_x_discrete(limits=(totals_standards_conc$std_mcat)[order(totals_standards_conc$group)][1:5]) + 
  scale_y_discrete(limits=(totals_standards_conc$conc)[order(totals_standards_conc$group)] ) +
  geom_text(aes(label = total_perc_in_all %>% percent(accuracy = 0.1)), color = "white", size = 2.5) +
  geom_vline(xintercept=4.5, linewidth=1.5, color = "#191970") +
  geom_hline(yintercept = 43.5, linewidth=1.5, color="#191970")+
  theme_classic() +
  coord_flip()+
  theme(legend.position = "none",
        axis.ticks = element_blank(),
        axis.text.x=element_text(angle=20,hjust=1, colour = "black"),
        axis.text = element_text(size=7, colour = "black"),
        axis.title = element_blank())+
  ggtext::geom_richtext(aes(x=5.385, y=23.5, label="Conceptualizations"), fill="grey", size=2) +
  ggtext::geom_richtext(aes(x=3, y=49, label="Standards"), angle = 270, fill="grey", size=2) 
# standards_conc_percentages_plot
# ggsave("standards_conc_percentages_plot.png")
```

```{r, warning=FALSE}
#the heatmap on top and table of stand/concept definitions on the bottom.
standards_conc_percentages_plot_and_table <- cowplot::plot_grid(
  gen_grob(totals_standards_table_flex, scaling="fixed", fit="auto"),
  standards_conc_percentages_plot,
  NULL,
  gen_grob(totals_conc_table_flex, scaling = "fixed", fit="auto"),
  nrow = 2, ncol=2, rel_heights = c(1.5,1), rel_widths = c(1,4) )
#standards_conc_percentages_plot_and_table
# ggsave("standards_conc_percentages_plot_and_table.png")
```

```{r echo=FALSE}
#function for creating comments
word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Anastasia"',
           'date="2023-08-12T15:00:00Z"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

# Introduction

> "_To be wealthy and honored in an unjust society is a disgrace._"
> — Confucius, The Annalects

The fiery wake of a rocket could be seen ascending, piercing the morning sky above West Texas. It was July 20, 2021, and on board was a small group of four passengers that included Jeff Bezos, the founder of Amazon and then the world's richest person [@harwoodBezos2021]. The mission that day was among the first-ever private suborbital passenger flights, and the adventure (described as "intense" by one of the passengers) lasted a total of 10 minutes and 10 seconds [@harwoodBezos2021]. In addition to intense, the undertaking was expensive: a seat for the flight had previously been auctioned for no less than $28 million USD [@griffinPerson2021]. Before even 10:00 am (EDT), when Bezos was back from his excursion, he took time to declare that this was the "[b]est day ever". To reporters covering the event he said "I want to thank every Amazon employee and every Amazon customer, because you guys paid for all of this" [@johnsonRoadToSpace2021].

Meanwhile, firmly grounded on planet Earth, the employees that Bezos thanked for his suborbital escapade were struggling with some very mundane problems of their own, and none as lofty as conflicting schedules that prevented them from flying in rockets [@griffinPerson2021]. According to reports, people employed directly or indirectly by Amazon for warehouse or delivery work had, for years, been treated to "inhumane" conditions [@fungEuropean2018; @scottAmazon2019; @greeneAlgorithm2021], and subjected to surveillance on the job, degrading schedules that drove drivers to urinate in bottles, crushing demands for productivity quotas that led injury, facing little or no job security, and treated less with gratitude than as disposable inputs to feed Amazon's earnings and consumers' demands [@tungAmazon2020; @BBCNews2021; @reeseTeamsters2022; @middletonRobot2023].

Coverage of the July 20 launch by the mainstream media was in many cases uncritical. "We're going to build a road to space so that our kids and their kids can build a future" Bezos declared, before adding "...we need to do that to solve the problems here on Earth" [@johnsonRoadToSpace2021]. Few reporters saw it fit to ask what problems the billionaire planned to solve on Earth, or what kind of future Bezos was trying to build, and for whose children. In other words, the billionaire was not confronted with questions about what did his trip do _for whom_ and _to whom_. A less dispassionate observer might have been excused for wondering (possibly aloud) about the basic _fairness_ of a man amassing a nigh unimaginable fortune that allowed him to build and fly his own spaceship, while masses of his employees were treated as throwaway cogs in the vast apparatus of his logistics empire.

The question of _fairness_, however, is not a simple one, and most of us would probably have been stumped to explain in a clear and precise way, beyond a disquieting intuition, just why the above picture was disturbing. _Justice_ is a political ideal based on the principle that individuals should be treated in a fair and equitable manner [@gosslingUrban2016, p. 2], giving and receiving whatever they are _due_ [@jaggarPhilosophicalChallengesGlobal2009, pp. 1-2]. The political [and contested; @vanoutriveHow2019] nature of the concept presents challenges that are only narrowly amenable to scientific inquiry. For starters, the notion that people are "due" something depends on the values of a society, as embodied in its systems and institutions [@karnerFrom2020]. Values,in turn, are not subject to natural laws, but rather are the result of intersubjective realities, which is to say illusions whose legitimacy derives from a collective will to believe. For example, justice would likely mean something very different to a person in a democratic society, than to another in a society where they owed their all—perhaps including their own time or bodily autonomy—to some collective illusion (e.g., the state, a monarch). As well, the meaning of "justice" would likely differ in yet another society where very few owned most, and most owned very little due to a different illusion (e.g., that wealth equates merit). In these two hypothetical cases, elucidating the meaning of "fair" would in all certainty be beyond the dreams of most, since fair would be what the state, the monarch, or the extremely wealthy said it was. In contrast, in democratic societies^[Paraphrasing, democracy is the worst of all collective illusions, except for all others.] individuals rights are valued above the whims of the few. In such a setting, the task of defining a "just" distribution of the burdens and benefits of things—from income, to roads, to space travel—quickly becomes muddled, encumbered even, by the necessity to pay attention to a multitude of voices, not all of them equally loud.

The rocket that took Bezos to the edge of space is a somewhat rare example of a transportation technology, a tool of space-time convergence. By enabling movement at very high speeds, rockets might—one day, at some indeterminate point in the future—prove essential to the expansion of the human species beyond our home planet. But in the present moment, the benefits of a private suborbital flight (e.g., the joy of movement, the sense of adventure, the awe of seeing Earth from space) are for a few, whereas the burdens (e.g., the use of non-renewable resources, the climate-altering emissions) affect us all, and not evenly at that. The benefits of public transportation, a much more common transportation technology, are for most, but in many cases we have penny-pinched these systems while concentrating the costs on those who have less [@allenSuburbanization2020; @kaeorueanAnalysis2020]. In a plutocratic society, those with most can (and often do argue^[See the so-called [Techno-optimist Manifesto](https://a16z.com/the-techno-optimist-manifesto/)]) that this distribution of burdens and benefits is fair since both benefits and burdens are earned. The fact that the likes of Bezos travel to space is proof that the likes of Bezos are due those trips. In a democratic society, the members of the collective might actually agree that large rewards (e.g., space flight) must be offered to highly qualified individuals (e.g., Bezos) to entice them to take important responsibilities (e.g., founding and leading Amazon). The fact that some must give up jobs because it takes them too long to reach them by public transportation would constitute proof that those people should have studied more, been earlier risers, worked harder [@SpurrWaiting2015; @greismanLittle2017]. Again, the members of the collective might think this a fine state of affairs, having come to this conclusion of their own accord or after being persuaded by billionaires. Or, contrariwise, the members of the collective might decide that this state of affairs is _unjust_: the _values_ of those in charge of defining what is "fair" matter.

Multiple national and cross-national studies suggest that people in many societies do indeed have some tolerance for inequality [@kiatpongsanHow2014]: it would appear that _some_ stratification, as suggested by @davisSome1945, is perceived as fulfilling a valuable function. However, extreme inequality is often frowned upon, and can lead to social dysfunction and other ills [@acemogluWhy2000; @taydasCan2012; @duIncome2019; @houleInequality2022]. But the perceptions of what is "fair" are neither universal or static. Instead, they are malleable, and can be affected by the existence of opportunities for social mobility [@shariffIncome2016; @artigeWhat2019], by exposure to inequality [@schroderIncome2017; @garciaChanging2023], by learned helplessness [@kimSocioeconomic2022], and even by how information about inequality is communicated to the public [@walkerPeople2021]. It follows that, in general terms, there are at least three different manners of thinking about inequality^[@martensTransport2016, in the introduction of his landmark text, talks about explanatory and prescriptive theories of justice; despite the saying "the arc of history is long and bends toward justice", it is unlikely that a predictive theory of justice exists.]: 1) in a _positive_ (or descriptive) manner, as the current state of the distribution of benefits and burdens of things; 2) also descriptively, as the _perceptions_ about the distribution of those benefits and burdens; and 3) normatively (or prescriptively), as the desired or ideal state of the same. Clearly, the three may coincide (e.g, if the perceived levels of inequality matched actual inequality and also how much of it the public desired). However, they do not necessarily have to, and in many cases will differ from one another. Measuring inequality as it is, in the most objective way possible is an essential task to decide whether there is a need to develop inequality-related policy to increase fairness. In turn, measuring the perceptions of inequality, in the most accurate way possible, may be important to achieve sufficient public buy-in in order to enhance the chances that a given policy will succeed.

Transportation systems, as a class of essential technologies that facilitate or impede social inclusion and activity participation [@churchTransport2000; @lucasTransport2001; @socialExclusion2003; @cassSocial2005; @casasSocial2007; @prestonAccessibility2007; @paezMobility2009], have increasingly come into focus from the perspective of equity. In response to this focus, a lively and rapidly growing literature has emerged on the topic [see _inter alia_ @martensTransport2016; @diciommoTransport2017; @guoSystematicOverviewTransportation2020; @karnerFrom2020; @vecchioTransportEquityLatin2020; @pereiraTransportationEquity2021; @vanweeEvaluating2021; @zhangLiterature2021; @desjardinsExaminingEquityAccessibility2022; @karnerAdvancesPitfallsMeasuring2023]. A cynical rationale for this interest could be that keeping track of actual and perceived inequalities can serve _at the very least_ as a gauge of social discontent [as Chilean authorities discovered to their sorrow when unrest erupted over the increase of metro fares in 2019; @BBCNews2019; @diazpabonInequality2021]. More optimistically, in democratic systems, there could be a genuine interest in being ready to respond to popular demands for fairness. Several challenges arise when approaching this endeavor, caused by the notorious complexity of transportation systems: simultaneously, they move people, goods, and information. As well, emerging technologies and service models can swiftly change the balance of benefits and burdens among a population [@guoSystematicOverviewTransportation2020]. Unlike income, which is as unidimensional as it gets, the benefits and burdens of transportation systems are diffuse over space and time. For example, transportation systems engineered to offer higher mobility for people _somewhere_, can simultaneously cut others off from essential opportunities _elsewhere_, as @rajeEngineering2004 poignantly illustrated with examples of infrastructure in the UK. Furthermore, the shades of policies past can continue to haunt a region and even the planet for decades or longer, as shown by the legacy of displacement and decay caused by urban highways all across the US [@archerWhite2020] and the time horizon for the impacts of climate change to be fully felt [@markolfTransportation2019].

But complexity is no excuse to shirk the task.

The objective of this report is to scan the state of knowledge in terms of defining and operationalizing "fairness" in the transportation domain. Much research has been devoted to the issues of _measuring _ equity in transportation, including (among many others) @ramjerdiEquity2006, @delboscUsing2011, @welchMeasure2013, @martensMeasuring2019, and @pritchardNew2022. Further, there are multiple sources that discuss the conceptual and philosophical foundations of equity and fairness in transportation [e.g., @martensTransport2016; @pereiraDistributiveJusticeEquity2017a; @vanoutriveHow2019]. Finally, previous reviews of planning documents have investigated equity from narrowly scoped perspectives, such as accessibility [@boisjolyHowGetThere2017] or a particular mode of transportation [e.g., cycling; @doranPursuitCyclingEquity2021]. These inquiries are valuable to scholar, planning agencies, and decision-makers alike, and the present review will tread similar, but not completely overlapping ground. In our estimation, there remains a gap in the literature in terms of understanding how standards for equity are developed and implemented in the transportation domain. In contrast, we do know that adoption of equity concepts in planning practice has lagged developments in academic work [@pereiraTransportationEquity2021; @boisjolyHowGetThere2017; @doranPursuitCyclingEquity2021; @linovskiEquityTransportationPlanning2020; @litmanEvaluatingTransportationEquity2022].

To illustrate this gap, we note how @beilerExploring2016, in their exploration of transport equity, cite the following strategies identified by the US DOT to address justice [p. 287]:

- Reduce adverse human health and environmental effects on minority and low-income populations.
- Include all potentially affected communities in the transportation decision-making process.
- Ensure that minority and low-income populations receive equitable benefits.

While commendable, the strategies are too vague, which makes it possible to implement them in a myriad ways, either genuinely to comply with the spirit of justice, or else performatively to deceive it [@mcculloughPerformative2023]. Some relevant questions include: how much should the adverse effects be reduced? To zero? Or to some tolerable level of adversity? What should that level be? What are the criteria for deciding that a community is "potentially affected"? What benefits should be distributed? Should the benefits be based on simple population weights? Or, contrariwise, should more deprived individuals be eligible for a larger share of the benefits?

These questions boil down to the development and use of _standards_ for transportation justice. The term "standard" is used to mean "something set up and established by authority as a rule for the measure of quantity, weight, extent, value, or quality"^[https://www.merriam-webster.com/dictionary/standard]. How much pollution is allowed to be generated, and where, depends on who is affected, and how much health is valued overall, as well as by whom. For example, firms may or may not adopt lower standards for the emission of pollutants in poorer areas; it might be that poor people end up being relegated to areas that adopt lower standards [@gouldsonFirms2006]. Regardless of the cause, the result is the same: pollution tends to be worse were poorer people are [@delucaCode2012].

Supporting the creation of (more) just transportation systems involves understanding the production, distribution, and management of transportation benefits and costs; how they are distributed; and what values are implemented (and by whom) in the form of standards [@pereiraDistributiveJusticeEquity2017a; @shellerMobilityJusticePolitics2018; @pereiraTransportationEquity2021]. Thus, for this review we engage the literature with the following questions in mind: 1) what is our current understanding of the things that transportation systems do, for whom, and to whom; 2) what does the literature say about the distribution of the benefits and costs of transportation systems; 3) and what values are embodied in normative statements about said distribution. Ultimately, this review aims to collate the existing academic knowledge on the matter, and present it in a manner useful to support the development and implementation of standards for equity in transportation planning and policy. In this, we aim to provide relief even, or particularly, to planners in those places where legislation explicitly calls for justice^[[As recently as -@martensFair2021, Martens and Golub note that Federal directives related to Title VI in the US "do not provide guidelines that can help agencies develop explicit standards to assess the distribution of accessibility benefits from projects or plans."]].

The rest of this report is structured as follows.

After this introduction, we set the stage for our investigation by laying out some important definitions. We then describe the methods used for searching, selecting, and reviewing the relevant literature. 
<!--
In Section 2, we set the stage by defining what do we mean by equity dimensions, conceptualizations and types of standards used in the literature search and throughout the paper.
In Section 3, we outline the methods used in the scoping review conducted.
In Section 4, we summarize the findings of the literature.
In Section 5, we discuss the methods connects to standards, conceptualizations and dimensions.
In Section 6, we summarize the findings and link them to future transport planning agendas in addition to visioning ways the findings can be used to build *just* transportation planning processes.
-->

# Setting the stage

> "_Man's capacity for justice makes democracy possible, but man's inclination to injustice makes democracy necessary._"
> ― Reinhold Niebuhr

> "_Never forget that justice is what love looks like in public_."
> ― Cornel West

## Justice, equity, fairness, and standards

In the introductory paragraphs we used the terms "justice", "fairness", and "equity" relatively loosely. This was done purposefully. As we noted, people often have strong intuitions of what is "fair", "just", and "equitable". These conceptions may or may not match those of the authorities who set the standards of fairness. But in democratic societies, the authority of political leaders, bureaucrats, planners, and all those charged with the business of governing, derives from the will of the people. It is therefore important to explicitly state how we plan to use these words, to clearly spell out our intuitions of "justice", "fairness", and "equity". A clear mutual understanding of these concepts is essential for constructive debate, and for participants of a democratic society to be effective arbiters of what is just.

Let us begin by stating that justice is an end goal, that is, a desirable state of affairs that we strive to achieve.

It is said that justice is attained when people "give and receive whatever they are due" [@jaggarPhilosophicalChallengesGlobal2009, pp. 1-2], and it ceases to exist when there are persons or groups that are denied "access to the opportunities they need to lead a meaningful and dignified life" [@karnerTransportationEquityTransportation2020, p. 440].  Justice is a fluid concept, because it depends on the desirability of different states of affairs, which may change between populations or over time. That said, it is possible to distinguish several forms of justice, including [see @jaggarPhilosophicalChallengesGlobal2009; @pereiraDistributiveJusticeEquity2017a; @karnerTransportationEquityTransportation2020]:

**Retributive justice**. It is concerned with the retribution due to people who do wrongs.

**Reparative (or restorative) justice**. This form of justice relates to the proper way to correct or rectify past wrongs even if the wrongdoers can no longer be meted retribution.

**Procedural justice**. The main concern of this form of justice is to ensure that the views, opinions, and preferences of all stakeholders are properly accounted for when decisions that affect their lives are made.

**Distributive justice**. This is perhaps the most commonly studied form of justice [see @jaggarPhilosophicalChallengesGlobal2009, p. 2; @pereiraDistributiveJusticeEquity2017a], and its main concern is the way the benefits and burdens of the tangible and intangible products of society are collected by different segments of a population.

It might be argued that all of the above are particular forms of distributive justice. Retributive justice, for example, is often concerned with the distribution of the benefits and burdens of being a member of society: the way it is usually achieved is by distributing intangibles like "freedom" (e.g., of movement, of association) as benefits, and/or claiming tangibles like money as burdens (e.g., fines). Reparative justice often distributes current benefits and burdens to redress past wrongs, for example by asking those who have benefited from past wrongs, even if unwittingly, to shoulder a bigger fiscal burden in order to cover programs that mete benefits to those who are still harmed by past wrongs. Procedural justice is the distribution of the benefits (e.g., the right to voice an opinion) and burdens (e.g., the effort required to develop an educated opinion) of the processes that lead to decisions with collective consequences.

We can then speak of the _purposes_ of distributive justice: to mete out retribution _fairly_, to repair past harms, and to ensure that procedures offer _equitable_ opportunities to influence outcomes. Equity and fairness are the instruments of justice, the tools by which society advances towards the end goal of justice.

The term "equity" tends to encompass various tools to understand the distribution of benefits and burdens of things among a population. In the transportation domain, the term is somewhat loaded because it is perceived as stemming from the authority of the state, and meant to assist with decisions about regulating and financing transportation spending [@karnerTransportationEquityTransportation2020]. Here, we are in agreement with @karnerTransportationEquityTransportation2020 that equity analysis should not be seen as an end in and of itself, but rather as a means to gather information about actual and perceived inequities. In this respect, the analytical tradition of equity, at least in transportation planning, means that the relevant models become embedded in the "political ecology of the estimated truth" [@kingModels1993]: in principle their assumptions and scope must be open and transparent, but they are also vulnerable to misuse and even abuse, as tools of subjugation.

Fairness, in contrast to equity, is somewhat more complicated to define. The concept does not have the same history of development as an analytical tool, and can be interpreted in numerous, and possibly discordant ways. That this is the case is convincingly demonstrated by @martensFair2021 in their study of the application of Title VI of the Civil Rights Act of 1964 in accessibility planning practice in the US. Title VI explicitly talks about the distribution of benefits derived from Federal funding: "[n]o person in the United States shall, on the ground of race, color, or national origin, be excluded from participation in, be denied the benefits of, or be subjected to discrimination under any program or activity receiving Federal financial assistance." However, as @martensFair2021 show, there are several ways to comply with regulations while achieving different outcomes, ranging from the banal (do not _knowingly_ discriminate), to the substantive (compensation for past discrimination, i.e., reparative justice). What kind of justice does fairness serve in each case? It depends on what was the reason for seeking justice in the first place. Our reading of @martensFair2021 is that fairness is a yardstick that is best deployed _a priori_ than _a posteriori_, for doing the latter risks rationalizing the outcomes rather than driving them.

The last concept that we discuss in this section is that o a standard. Briefly, standards are a way of making concrete statements about fairness. Returning to the ambiguities in Title VI discussed by @martensFair2021, the attainment of justice depends on the standard used to indicate fairness. For example, explicit non-discrimination constitutes a very weak standard that takes aim at the actions of agencies instead of the recipients of the benefits; accordingly, any distribution of benefits would be considered fair, as long as the agency does not explicitly and knowingly targets or denies the benefits to particular groups. The standard provides conditions to determine whether a situation is _fair_. A similarly weak standard is a _Pareto improvement_, whereby it is possible to concentrate the benefits as long as no group is worse off compared to the status quo. A policy is fair as long it does no harm. A somewhat more strict standard, a Pareto-Plus improvement, stipulates that an intervention is fair when all groups receive at least _some_ (non-trivial) benefits; the size of the benefits for each group is irrelevant. In contrast to the notion of "do no harm", such a standard embodies the ideal that no one is denied benefits. An egalitarian standard would weigh the benefits or burdens by population, and fairness is achieved when each group give or receive in proportion to their size. An affirmative action standard is even stricter, since it requires the benefits to be distributed in a non-egalitarian way that favors those who are still harmed by past or present discriminatory practices.

To quickly recap the discussion so far, justice is an end. But to understand what that end is, we must clearly define standards of fairness. Equity analysis is tool to measure where the actual or perceived distribution of the burdens and benefits of the products of transportation systems stand with respect to the standard, in other words, instruments to see how close or far we are from a just situation.

In the following section we will discuss the analytical apparatus that we will use to interrogate the literature on equity standards in transportation.

## A framework to analyze questions of justice

For this report, we are inspired by the framing of @jaggarPhilosophicalChallengesGlobal2009 for philosophical questions of justice^[Similar questions are found peppered throughout the literature. This is done either explicitly, as for example in @karnerTransportationEquityTransportation2020, who ask "of what", "for whom," and "how much" in reference to equity; or implicitly, as in @gosslingUrban2016, who asks of the outputs of transportation "what?" (exposure, space, access) and "for whom?" (gender, age, ethnicity).]. According to -@jaggarPhilosophicalChallengesGlobal2009, Western philosophy has approached the issue of justice by asking "Where?", "When?", "Who?", "What?", and "How?". Conventionally, discussions about justice have been aspatial, or rather, seen from the point of view of social space instead of geographical space, despite an early interest of geographers on the matter [@pirieSpatial1983]. The texture of the questions becomes more immediate and crisp when talking about transportation, which is inherently about space and time.

- **"Where?"** Traditionally related to the applicable domain or sphere of life relevant for justice. Conventionally this meant the in-group [e.g., members of the same nation state; @jaggarPhilosophicalChallengesGlobal2009, p. 3], a notion since expanded (e.g., with the Universal Declaration of Human Rights) to include all members of the human species, but has yet to cover other living creatures. <!-- SOUTH AMERICAN CONSTITUON THAT DOES INCLUDE IT NOTE FOR ANASTASIA--> In the case of justice in transportation, the question of "where?" is paramount; it might be argued that, by their very nature, transportation generates inequalities. By concentrating the effects of space-time convergence (for instance, by providing access to a transit system or a highway), an inequality is automatically generated. The burdens of transportation, in contrast, are often diffuse. They are incrementally paid, for example by a distributed population in the form of taxes by a spatially distributed population, or in poor health by people who were not party to transportation decisions made elsewhere.

- Conventionally, the question of **"When?"** refers to the situation, that is, the social circumstances within which the demands of justice have application. In our case, we might as well ask about the temporal aspects of transportation systems, which include _when_ to invest in transportation [e.g., @rabelloEvaluation2015] (and as a result when to generate a spatial inequality), for how long the burdens and benefits can still be associated to a specific intervention, and whether claims to reparative justice expire over time. <!-- WHEN IN TIME SOMETHGING HAPPENS TOO - WHAT INVESTMENTS HAPPENED IN THE PAST THAT CAST SHADOWS-->

- When we ask **"Who?"**, we inquire about which entities should be regarded as subjects of justice <!-- OF INEQUITIES / MEASURING THE INEQUITIES OF 'WHO'. LIKE INCOME, THE MOBILITY TOOL IS MODIFIABLE -- THE MOBILITY TOOL IS NOT MODIFIABLE IN MANY CASES, IT IS WHO YOU ARE . T IS ALMOST LIKE A CHARACTERISTICS OF THE PERSON. PART OF THE SOCIAL CONTRACT THAT IDENTIFIES EOPEL WITH THEIR MOBILITY TOOL-->, meaning who are entitled to make claims of moral consideration from the perspective of justice. To make it tractable, this question is often approached through the filter of population groups, which may include the several traits, including gender identity, ableness, ethnicity, age, caste, and income. Often, it is appropriate to reflect on the intersections between traits, given evidence that the lived experiences of, say, a White woman and a Black women, can be markedly different. A possible complication in the case of transportation is that disentangling the "who" from the tools available to them is not always straightforward. Clearly, a person is not their mode of transportation; in practice, though, there are large segments of the population who live in situations where they cannot extricate themselves from the tools for mobility that they can use, either because they have driven themselves out of choices [see @laveryDriving2013], or have been driven out of choices by factors beyond their control, effectively ending up as captive users of a single mode [@jacquesRescuing2012; @cherancheryInvestigatingPerceptionCaptive2018]. In societies that have grown into transportation monocultures with a predilection for automobility [@millerCollaborative2011] there may actually be less choice about mobility tools than we would like to assume. So, while it is important to avoid conflation of the "who" with the "what",  for analytical purposes we need to be mindful of the connection between person and their mobility tools. In the case of transportatin, in addition to members of the public who use transportation systems, there is another category of **who**, that stand possibly in opposition to users, namely the entities charged with providing services, maintaining infrastructure, and so on. These could be ministries or departments of transportation, transit agencies, public works departments and others. Identifying these entities is relevant to ellucidate who is responsible for apportioning the benefits or mitigating the burdens of transportation.

- **"What?"** asks which entities should be regarded as objects of justice, meaning which kinds or categories of things should be distributed in a just manner. To understand the distributional implications of transportation systems, it is essential that we are clear about what they do. In other words, what do transportation systems _produce_? At their most fundamental, transportation systems are space-time convergence tools, tools that improve the rate at which time is traded for space. They usually do this by increasing the speed of movement: sidewalks facilitate walking, traffic lights facilitate the ordered flow of vehicular traffic, and a launching pad makes it possible for a rocket to takeoff. With complex interlocking parts (sidewalk, road, traffic sign, parking regulations), transportation systems produce _mobility_, the potential for movement.The realization of this potentials happens through travel but, as the adage goes, travel is derived demand, which seems to hold for most (even if not all) situations [e.g., @mokhtarianUnderstanding2001; @redmondPositive2001; @paezEnjoyment2010; @whalenMode2013]. For this reason, we cannot stop at considering mobility, but the ultimate goal of mobility, which is to reach destinations. In combination with land use systems (the spatial distribution of opportunities on the landscape), mobility produces _accessibility_, the potential to reach destinations. For this reason, we can think of the objects of transportation justice as being _proximate_ (the tools of mobility, mobility itself), and _ulterior_ (accessibility, opportunities for activity participation). The burdens of transportation are also many and varied. Some are direct and paid directly by the traveler (e.g., travel time, out-of-pocket costs), but many others are indirect and related to network externalities (e.g., exposure to pollution).

- The next question is **"How?"**, and it relates to the allocation of various objects of justice ("what") to various subjects of justice ("who") in various circumstances. **Equity standards** are a tool for answering this distributive question: how do we allocate burdens and benefits and to whom? Standards are thresholds that when operationalized effectively define what is in/equitable. The thresholds can be quantitative (e.g., square meters of green space per capita), or they can be qualitative descriptions (e.g., do not knowingly discriminate), or a mixed of the two. Some examples include: maximum travel distance/cost/time to or from key destinations, levels of maximum exposure to externalities (i.e., noise or air pollution),un/fulfilled needs, and dis/satisfaction with travel. A number of theoretical and conceptual frameworks exist to support us when approaching this question, and we can draw from concepts in transport-related social exclusion, transport disadvantage, and/or transport poverty, which are typically based on equity principles, such as utilitarianism, Sen's capabilities approach, and sufficientarism.

- Lastly, convincing answers to the above questions require a supporting rationale: a **"Why?"** [@jaggarPhilosophicalChallengesGlobal2009]. This is perhaps the most slippery of all the questions posed here. Justice is an inherently social construct <!--, and does not exist outside of human society-->. Asking **why?** amounts to asking what sort of social contracts regulate human interactions. These contracts can be defined by constitution, but there are often unwritten and possibly contested variants. To give an example drawing from the Canadian Bill of Rights, a number of rights and fundamental freedoms [including the liberty to move freely; see @chartopediaArt72023] are recognized to exist in Canada "without discrimination by reason of race, national origin, colour, religion or sex". Notice that this declaration includes several individual traits that define the subjects of justice, but does not consider age as one aspect of the person. Does this mean that they apply universally irrespective of age? Surely not, since no reasonable person would consider a toddler's demands for freedom of association as an absolute. So, when do these rights fully apply in the life of a person? In 2017, Adrian Crook of Vancouver, B.C., was warned by the province's Ministry of Children and Family Development that his kids could not be out of home in the community, alone or with other kids the same age, without supervision. The Ministry's argument was that children riding the bus unsupervised compromised the parent's ability to provide care and placed them at risk [@brendHere2017]. Crook argued that the goal in teaching his kids (then aged 7, 8, 9 and 11) to use transit to go to school was to raise self-reliant children [@stueckYoung2017]. While both sides discovered that there were not clear rules about kids riding transit alone [@brendHere2017], statistically there is evidence that travel by car is riskier than travel by bus [@morencyTraveling2018]. If the public good is to reduce risk, should not children be banned from riding in cars until they are ten? In a functioning democracy like Canada, we have chosen to care about fundamental freedoms and rights. This provides an answer as to "why" we would even consider children as subjects of justice. But the devil is in the details, and the other questions discussed above are needed to pin the devil by the tail. Are children under the age of 10 subjects of justice (i.e., are they a "who")? If so, is the ability to use transit unsupervised an object of justice (i.e., the "what")? When and where do these questions apply (e.g., riding the bus to school, or to a social event; during the day, or late at night)? Eventually, in 2020, a court of appeals ruled in Crook's favor, allowing his children to continue to ride the bus, thus providing much needed clarity with respect to the use of transit by children who desire independent mobility [@stueckCourt2020]. In actuality, the courts provided a standard of equity: starting at a certain age, children are objects of justice from the perspective of freedom of movement, and they are due the benefits of independent mobility by transit.

<!-- 
IN THE PARAGRAPH ABOVE, ANASTASIA SUGGESTS DISCUSSING WHY DO WE THINK THAT OUTSIDE IS DANGEROUS, THAT CARS ARE SAFER, THAT CERTAIN PEOPLE ARE TO BE AVOIDED.

THIS IS WHY WE ONLY FOCUS ON INEQUITIES. JUSTICE IS FLUID AND SLIPPERY. WHAT HAS BEEN DONE WHAT HAS NOT. 
-->

The focus of this report is on standards of fairness that, combined with the use of equity analysis, can help us understand how better to move towards just transportation systems.

# Scanning the lay of the land: Methods

> "_Not everything that is faced can be changed. But nothing can be changed until it is faced._"
> — James A. Baldwin

<!--
NOTE: This text was in the Definitions, but we decided that these definitions were a thin conceptual layer that could be removed. Instead of "equity dimensions", "conceptualization", and "standards" we will stick to using the where, when, who, what, how, and why conceptual framework. For example, the EQUITY DIMENSIONS are simply the "what" or "objects of justice". The "conceptualizations" are the "how". And so on.

For all literature reviewed in the transportation domain, equity *dimensions*, *conceptualizations*, and *standards* are discussed through the "Where?", "When?", "Who?", "What?", and "How?" framings.
As follows, we define each of these categories and how they inter-connect.

**Equity dimensions**: the assessment of transportation equity encompasses various dimensions, mirroring the multifaceted nature of transportation systems. These dimensions can be distinguished as the **"What?"**, specifically: (1) mobility/accessibility, (2) traffic-related pollution, (3) traffic safety, and (4) human-health as provided in a conceptual framework by @lucasMeasuringTransportEquity2019.

All papers reviewed include dimensions that cover the **"What?"** of the transport equity issue and can be further disaggregated:

-   *When* and *Where*: includes the article's publication year, the geographic context (i.e., urban, suburban and rural areas, countries) and location.

-   *Who*: includes the type of population group, mode type, activity/opportunity type, and transport network status.

papers that include equity **dimensions**, **conceptualizations**, **standards** and substantive answers to the "When?", "Where?", and "Who?" in relation to equity in transportation systems (i.e., the users, the infrastructure, the service, the outcomes, or a combination of all) are used to answer the following research questions:

-   What are the equity dimensions, equity standards, and equity conceptualizations and how are they applied in the existing transportation equity academic literature?

-   Based on how equity has been applied in the literature, what may be paths forward in theorizing justice in the context of transportation systems
-->

This review examines the breadth and depth of the academic literature on transportation to identify the extent to which standards for equity are defined and employed. In this task, we follow the Joanna Briggs Institute (JBI) approach to the conduct of scoping reviews. This approach builds upon the Arksey and O'Malley (2005) framework [@petersUpdatedMethodologicalGuidance2020]. The JBI is .

Further, the review is guided too by the Preferred Reporting Items for Systematic Reviews and Meta-Analyses extension for scoping reviews (PRISMA-ScR), which is consistent with the JBI approach [@triccoPRISMAExtensionScoping2018]. The use of these methods allows us to explore, in a consistent and organized manner, a relatively specialized topic within the broader transportation literature. In this way, we aim to collate the current knowledge as found in the landscape of published research.

The primary research question and the protocol were initially defined in consultation among the authors of the report. In other words, the starting point was the level of knowledge of experts in the subject matter. The initial draft of the protocol was refined from preliminary searches of related-reviews [e.g., @iglesiasHowUnevenUrban2019; @sagarisUsingPARFrame2020; @vecchioAccessibilityCapabilitiesApproach2021], and in consultation with a University of Toronto Research Services Librarian and Liaison Librarian in City Studies.

The methods are described in two parts: (i) development of the search strategy and (ii) selection of evidence and data extraction. The framework for analyzing questions of justice discussed in the preceding section was used in both stages, but was particularly valuable for selecting the evidence and for analysis for data extraction.

## Search strategy

<!-- SELECTION OF DATABASES?? -->

To guide the selection of search terms within the search query, **inclusion** and **exclusion** criteria were developed [@petersUpdatedMethodologicalGuidance2020]. For the inclusion criteria, the mnemonic PCC (population, concept, and context) was adopted (see Appendix @fig-A1: for details).

Next, the inclusion and exclusion criteria were deployed to develop the search strategy. The search strategy was refined iteratively, adding topic search terms by stages (e.g., terms in the title, abstract or key words). The terms were bundled by means of logical connector terms "AND" and "OR". These stages are summarized next. The full search term queries can be consulted in Appendix @fig-A1.

1. An initial limited search of Web of Science (WoS) Core Collection was undertaken to identify key documents This collection contains documents in journals, conference proceedings, and books published all over the world. Separate sarches using the terms 'transportation' and 'equity' were generated. From these searches, we examined the text contained in the titles and abstracts of relevant papers, the index terms used to describe the papers, and subject heading searches when available. As we developed a clearer outline of the literature, we continued with this process by adjusting the terms used for the search. This took the form: ("Transport" OR "Transit" OR "Car\*" OR "Walk" OR "Bike"...**1**) AND ("Equity" OR "Justice" OR "Fair"...**2**), where **1** and **2** signify additional terms relating to 'transportation' and 'equity', respectively.

1. Upon inspection of the preliminary search results and after achieving a consensus among the authors, the set of search terms related to 'equity' was expanded into three sets of terms. The first set describes theories and concepts of equity, the second describes the object of justice (i.e., the "what" in our analytical framework), and the third describes terms referring to standards (i.e., the "how"). All three sets of terms were augmented following an iterative process of refinement. The final search query took the following general form: ("Transport" OR "Transit" OR "Car\*" OR "Walk" OR "Bike"...**1**) AND ("Equity" OR "Justice" OR "Equity" OR "Fair"...**2**) AND ("Accessibility" OR "Mobility" OR ...**3**) AND ("Standard" OR "Threshold" OR ...**4**) where **1**,**2**,**3**, and **4** signify additional terms included in the sets combined with "OR" logical connectors.

After testing the search strategy on WoS Core Collection, we proceeded to apply to an augmented list of databases, that expanded on our search using the Core Collection of WoS. The databases used are: WoS General Collection-Science Citation Index Expanded, WoS Social Sciences Citation Index, and Transportation Research International Documentation (TRID). The definitive version of the search was completed and exported by the lead author on March 21st, 2021.

## Evidence selection and data extraction

The semi-automated nature of the search strategy tends to be overly inclusive, which serves our purpose well, since we aim to begin with more documents than are strictly needed, and so reduce the risk of omitting relevant material. The next stage is to trim the corpus of documents, a task that can no longer be automated, and requires expert knowledge. Selection of evidence is where this expert knowledge really comes to bear, and it consists of scanning the literature retrieved by the search strategy to retain in the corous only those papers that fit the inclusion and exclusion criteria. This process was pilot-tested with a subset of papers before being implemented on the full set. *Covidence*^[https://www.covidence.org/], an online application for literature screening, was used for all steps of selection and data extraction on the full export of literature. Covidence is designed for collaborative work, and helps to document the work of multiple reviewers. The steps of this process are shown in @fig-fig1.

```{r methods_framework-fig1}
#| label: fig-fig1
#| fig-cap: "Evidence selection process framework. Step 1 (orange) is title and abstract screening, step 2 (green) is full-text review, and step 3 (purple) is data extraction."
knitr::include_graphics("figures/Methods_framework.png")
```

The following steps summarize the process:

1. The first step (orange box in @fig-fig1) included screening all titles and abstracts of papers on whether they included transportation equity as defined by the PCC. Each paper was screened by two independent reviewers who then voted for inclusion, exclusion, or uncertain inclusion. All uncertain papers, conflicting papers, and papers missing abstracts were reviewed by a third person for inclusion or exclusion.

1. The second step (green box in @fig-fig1) included scanning all full-text papers which passed step 1. These papers were reviewed to determine if they included a relevant "how", i.e., an standard and/or relevant theoretical or conceptual discussion. At this stage, papers were evaluated again by two independent reviewers who voted for inclusion or exclusion. If an article was voted to be excluded, it was tagged with one of five possible reasons for exclusion, namely (1) no standards included; (2) no relevant conceptual elements included; (3) no standard and no conceptual elements included; (4) send back -- QA issue; or (5) other. Discrepancies were resolved by a third reviewer.

1. In the last step, a data extraction template for each record was filled by one reviewer (purple box in @fig-fig1). The data extraction template was created with the aim of striking a balance between the complexity of categories and the simplicity of summary; information related to "What?" (the object of justice), "Who?" (the subject of justice), "Where?" (the geographical context and sphere of life), and "How?" (equity standard(s) and concepts) was filled out for each study. @fig-A2 contains the template that was input into *Covidence* and used throughout.

```{r PRISMA-fig2}
#| label: fig-fig2
#| fig-cap: "PRISMA flow diagram for the evidence selection process. ES signifies equity standard and EC signifies equity conceptualization."
knitr::include_graphics("figures/Pflow_chart.png")
```

The evidence selection process is also represented using the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) flow diagram [@pagePRISMA2020Statement2021] in @fig-fig2. Notably, two rounds of exclusion occurred during the assessment for full-text eligibility. 1710 studies entered step 2, 1223 were excluded and the remaining 487 papers entered step 3. The data extraction template used by the reviewers (authorship team) in step 3 revealed that, as expected, inclusion was initially too generous, and some papers were not sufficiently relevant, because of a lack of content on standards and/or conceptual/theoretical elements. In this fashion, 322 papers were further excluded and data extraction was completed to give a final corpus of 165 papers. A summary of the reasons for exclusion of the 1545 papers (between steps 2 and 3) are included in @fig-fig2.

# The lay of the land: a summary of findings  `r word_comment("NOTE: in this section, only papers that are included in the review (i.e., have a full template) are cited.", "findings")`

> "_Give me knowledge, so that I may have kindness for all._"
> — Native American Proverb

A synthesis of the findings from the data extraction process (based on the template shown in @fig-A2) is detailed in this section. The presentation of findings is less granular than the template to highlight the key trends in the literature.

## "When" and "Where": the context for justice

```{r counts-case-study-and-urban}
case_cont_count <- strsplit(Consensus_only_analysis$`_Case study continent`, "; ",
                       fixed = T)
case_cont_count <- data.frame(
  id=rep(Consensus_only_analysis$`_Case study continent`, lengths(case_cont_count)),
  case_cont=unlist(case_cont_count))
case_cont_count <- case_cont_count |> group_by(case_cont) |> summarize(count = n())

urban_form_count <- strsplit(Consensus_only_analysis$`_Case study urban form`, "; ",
                       fixed = T)
urban_form_count <- data.frame(
  id=rep(Consensus_only_analysis$`_Case study urban form`, lengths(urban_form_count)),
  urban_form=unlist(urban_form_count))
urban_form_count <- urban_form_count |> group_by(urban_form) |> summarize(count = n()) 
```

@fig-fig3 displays the papers included in this review by year of publication and geographical provenance. The literature related to transportation equity has grown evidently voluminous, particularly since 2019. Of note is the geographic scope of the relevant literature. The majority of papers (`r ((case_cont_count$count[4] + case_cont_count$count[5]+ case_cont_count$count[6])/sum(case_cont_count$count)) |> percent()`) contain case studies based in the Global North, with many studies from North America (particularly USA and Canada), Europe (particularly UK, France, Spain and Scandinavia), Oceania (Australia and New Zealand), and Asia (Japan and Israel). For the reasons previously discussed, setting standards (that is, making statements of fairness) is a highly context-specific endeavor. Publication patterns, being what they are, display a disproportionately low number of items from the Global South where equity issues are perhaps as, or even more pressing, than in the Global North [e.g., @BBCNews2019]. While our search strategy and review tries to be as faithful to the literature as possible, it does not allow us to lay claim to reporting about on-the-ground transportation justice efforts from a truly global perspective.

```{r plotting-year-cont-plot-fig3}
#| label: fig-fig3
#| fig-cap: "Papers included in the review by year of publication and case study continent."
year_cont_plot
ggsave("year_cont_plot.png")
```

The smattering of studies from the Global South are predominately from **Asia**, notably China, but also India, Thailand, Iran, Philippines, and Indonesia. <!-- These countries add to our knowledge stemming from research reported from South Korean, Israel, and Japan. These studies focus on a variety of modes, population groups, and equity conceptualizations and standards that cannot be succinctly summarized. <!-- develop further, if there's room?-->The next most common focus within the literature from the Global South is from **South America**. Many of these studies mention a systemic absence of evidence relevant to the region [@vecchioTransportEquityLatin2020]. Despite the growing recognition in the literature of the interconnections between transport development, social exclusion, and poverty [@benevenutoExaminingTransportNeeds2020], a number of studies underscore an ongoing neglect of the social dimension of transport during the planning stage [@benevenutoExaminingTransportNeeds2020; @boisjolyAccessibilityMeasurementsSao2020]. Many studies also point at affordability as one of the main mobility barriers in the region [@falavignaAssessingInequalitiesPublic2016; @rivasHowAffordableTransportation2018], while some highlight multi-dimensional concerns such as public transport accessibility and quality of walking environments that contribute to mobility inequalities [@tiznadoaitkenRoleAccessibilityPublic2018].

Studies pertaining to **Africa** are even less numerous compared to the South American literature.
A shared characteristic among the studies from these two continents is a scarcity of official transport data [@friedMeasuringSustainableDevelopment2020] and a reliance on policy guidelines developed by international organizations. These studies also incorporate the use of informal transportation options and the pressures to develop road network infrastructure (which tends to support car dependency) over meeting the mobility/accessibility needs of citizens [@thondooSmallCitiesBig2020]. To address these challenges, researchers compile databases based on open and geo-referenced data, calculate objective and/or subjective measures [@berheAdaptationDissonanceQuality2014], and focus on advancing transport justice for low to medium income countries (LMIC) by aligning their goals with external policy guidelines such as the Sustainability Development Goals (SDGs), particularly those related to universal accessibility [@friedMeasuringSustainableDevelopment2020].

Of all the studies reviewed, `r (urban_form_count$count[3]/sum(urban_form_count$count)) |> percent()` focus on urban and suburban settings and are highly varied in their research aims. To give an example, @coxQualitativeStudyAccessibility2020 qualitatively examine cycling as a mode of travel for people with disabilities in a typical mid-size town in the UK. @ampeImpactChildBike2020, on the other hand, work to identify the lateral clearance that motorists should maintain when passing cyclists with children seats. While both studies refer to the same objects of justice ("what", i.e., the right to the road and cycling), the subjects of justice (the "who") are different (people with disabilities and children), and the "where", that is, the geographical context for the examination of fairness, differs as well.

The remainder of the studies focus on rural regions (`r (urban_form_count$count[2]/sum(urban_form_count$count)) |> percent()`). To illustrate, we highlight the work of @caoIndicatorsSocioSpatialTransport2017, who examined transportation disadvantage in remote places which rely on inter-island ferry trips in the rural Philippines. Similarly, @parrySocialVulnerabilityClimatic2018 studied remote communities in the Amazonian region, and suggest that "increasing accessibility through road building would be maladaptive, exposing marginalized people to further harm and exacerbating climatic change by driving deforestation" (pp. 125).

## "Who": the subjects of justice

@fig-fig4 showcases the population group types that are the focus of the papers in our review. From this tally, papers that consider income groups is the most widely represented in the literature. Particularly, most papers that focus on income pay usually focus on the lowest-income groups, as they experience the least benefits in terms of mobility and accessibilty, and usually the greatest burdens as well, in terms of cost, exposure, and so on [@peungnumsaiGridBasedSpatialAnalysis2020; @zhaoUnequableSpatialAccessibility2020; @falavignaAssessingInequalitiesPublic2016].

```{r plotting-population-group-plot-fig4}
#| label: fig-fig4
#| fig-cap: "The proportion of papers that focus on each type of population group. Categories for population groups were generated upon data extraction."
pop_plot
ggsave("pop_plot.png")
```

For example, there is an abundance of evidence to suggest that low-levels of household income is a significant determinant of transport-related inequities (e.g., access to public transport supply in Bangkok region (Thailand) [@peungnumsaiGridBasedSpatialAnalysis2020], access to employment opportunities in various cities in Brazil [@boisjolyAccessibilityMeasurementsSao2020], and unfavorable rates of environmental noise, air pollution, and green space per resident in Rijnmond region (Netherlands) [@kruizeEnvironmentalEquityRole2007]. However, this should be kept in context, as low-income is not universally associated with lower transport-related benefits for every object of justice. For instance, in Sheffield (UK), @mearsUnderstandingSocioeconomicEquity2019 demonstrates that historically working-class neighbourhoods (i.e., lower income working population) have more access to green space than other neighbourhoods due to urban planning approaches during the Victorian-era. However, the quality of green spaces are less than average. Similarly, @bertrandMeasuringMappingDisparities2008 finds that lower income groups do not always have below average accessibility depending on the granularity of analysis (i.e., the distance-to-food threshold for the cumulative opportunity measure).

Age is the second most common category of population group of focus within the reviewed literature. Many papers that focus on this category highlight differences in age-related capabilities; for instance, @martinez-jimenezAccessibilityCultureEducation2019 and @arranz-lopezSocialSpatialEquity2019 investigate travel distances and times to various opportunities based on specific age groups, acknowledging that age is an important consideration, and associated with differences in access to opportunities. The most commonly age groups are school-aged children and older populations. Research on school-aged children analyzes their wellbeing [@laszkiewiczChildrenGreenWalk2020], exposure to green space [@corazzaMethodologyEvidenceCase2020], access to schools [@sharmaSpatialSocialInequities2022], and aims at understanding and encouraging active travel journeys [@mackieOvercomingBarriersCycling2009; @mehdizadehWalkingTimeSchool2017]. Papers that focus on older adults often have similar aims as those on children, and try to understand transport-related impacts on wellbeing [e.g, @chenSpatialAnalysisFramework2020], measuring accessibility to key destinations by population group [e.g. @chengInvestigatingWalkingAccessibility2019], and seeking to understand how to better meet travel needs [e.g., @nordbakkeTransportUnmetActivity2015].

The third population group most commonly studied is what might be called 'composite vulnerable population groups' that exist at the intersection of several individual traits. These papers use some sort of composite vulnerability index that captures multiple population characteristics, including low-income, unemployment, immigrant status, family household characteristics, and so on. These indices are typically generated from official government sources or author-informed census data creation, and produced using a variety of methodologies. For instance, @awuorInfluenceEnvironmentalHealth2019 use the Neighbourhood Equity Index (NEI) to investigate disparities in premature death in Toronto (Canada). The NEI is a composite index that was developed by the city to capture differences in the City's neighbourhoods by ranking them based on socio-economic characteristics (e.g., social assistance, unemployment, income) and physical environmental characteristics such as green space availability. Other works use national census indicators such as the social and housing deprivation index [e.g., @pucciInequalitiesJobrelatedAccessibility2019 or explore transport disadvantage, equity in policy implementation, or transport-related mortality burden by means of census measures (e.g., household poverty) and transport-related indicators [e.g., accessibility; @@aldredEquityNewActive2021; @iungmanImpactUrbanTransport2021; @sunPublicTransportAvailability2021; @scheurerSpatialAccessibilityPublic2017]. Similarly, Environmental Justice (EJ) indicators have been used in the US literature to identify neighbourhoods that have a higher than average proportion of low-income and non-white populations (i.e., a composite vulnerable population group'). Numerous studies have used EJ analysis to evaluate the equity impacts of transportation projects [e.g., @rowangouldIdentifyingEnvironmentalJustice2016; @parkTransitParksEnvironmental2021; @reddySafeguardingMinorityCivil2010].

`r word_comment("Papers that focus ONLY on (Dis)ability, or ONLY on Gender, or ONLY on Race/Ethnicity, or ONLY on Employment/Education level... the only focus", ".")` Papers that exclusively focus on populations with (dis)abilities e.g., [@parkJourneyVisuallyImpaired2017; @chiscanoImprovingDesignUrban2021; @orellanaWalkRollMapping2020] are relatively common in the literature. They often assess universal design guidelines and the capabilities of people who face disabilities to travel. However, from another perspective, papers with an exclusive focus on gender, race/ethnicity, or education level/employment are less common in the literature reviewed. Only two papers focus on gendered differences in cycling/active transportation (e.g., @adlakhaMindGapGender2020's case study in Chennai (India) and @xieWonCycleRoute2018's case study in Cardiff (UK)). Only two papers focus on race/ethnicity exclusively focusing on how minority ethnicity communities are in proximity to green space [@silvaEnvironmentalJusticeAccessibility2018] and culturally diverse family physicians in Toronto (Canada) [@wangModelingSpatialAccessibility2011]. Furthermore, papers that focus *solely* on education/employment status are not present in the reviewed papers. This is to say, papers that feature gender, race/ethnicity, or education level/employment population groups often feature them alongside other population group characteristics (as a trait in 'composite vulnerability measures'). This contrasts the prominence of studies that exclusively center on disabilities.

We include a catch-all category of papers (*Other*) that include group populations that less commonly have been the subject of research. Examples include: veterans and access to specific-healthcare needs [@mooneyTravelDistanceBarrier2000], pregnant people and access to services [@vadrevuMeasuringSpatialEquity2016], and youths who live in foster care [@batscheUsingGISEnhance2012]<!-- rural populations people who use transit and their service needs relative to urban needs @sharavWhatTransitService2019, only cyclists in the context of health-implications of cycling in Delhi (India) @apparicioCyclingOneMost2021 ) -->. Overall, the diversity of the *Other* population group classification demonstrates the diversity of transportation-equity concerns across population groups in the reviewed literature and the interplay of characteristics in the literature [@vecchioFairTransportPolicies2022].

## "Who" and "What": the tools of mobility

```{r plotting-mode-plot-fig5, fig.width=5}
#| label: fig-fig5
#| fig-cap: "The proportion of papers that investiage each type of mode. Categories for modes were generated upon data extraction."
mode_plot
ggsave("mode_plot.png")
```

As previously noted, mode of travel is not a trait of the individual, but in principle is a modifiable factor. That said, the existence of captive users by mode makes it challenging at times to extricate the mobility tool from the individual. In this section we choose to view the mode of travel primarily as an object of justice (i.e., a "what"), but occassionally this will be enjoined to the subject of justice (i.e., the "who").

It is interesting from this perspective to notice that the primary emphasis within the literature reviewed centers on public transit (@fig-fig5)., a mode that often is perceived by users as less than ideal [@paezEnjoyment2010; @mellaDrivers2021] and that yet is often reckoned as the only realistic mobility tool for many [@jacquesRescuing2012; @welchMeasure2013; @cherancheryInvestigatingPerceptionCaptive2018]. The variety of topics assessed from the perspective of public transit are varied. For instance, @mckeyCrowdsourcedMappingHealthy2020 identifies 'food deserts' in Dallas (USA) considering public transit accessibility. Other contributions intersect public transport and individual needs. Examples include the study of universal design and barrier-free transportation for people who face disabilities [e.g., @jimenez-espadaResearchProblemUniversal2021; @liuStatusQuoChallenges2019] or possible improvements to public transport for people with autism [@limFacilitatingIndependentCommuting2021; @feeleyValidationParatransitSkills2019]. The strong focus on public passenger transport systems could be due to their policy importance. They are a clear public good, and seen as a natural object of justice. Accordingly, public transportation is seen as adaptable to meet the demands of justice, for example by funding it sufficiently to provide barrier-free transport for most. In this respect, the literature points at a variety of factors that conspire against it, including low population and opportunity density (the outcome of decades of deliberate policy choices), fiscal prudence (the public face of the choice to lower taxes), and political will [to maintain or further unsustainable socio-technological systems, from SUVs to space tourism; @markardUnsustainabilities2023].

Transit is also often central to multi-modal or holistic comparisons that may serve transport equity analysis. As an example, @brusselAccessAccessibilityCritique2019 compares three different approaches to measure accessibility in the context of the Sustainability Development Goals (11.2) for the case of Bogota (Colombia), all of which capture some/all of the public transit system while others capture road and/or pedestrian systems. From a different perspective, @renneWhatHasAmerica2018 reviews natural disaster emergency evacuation plans from the lens of car-less (and oftentimes vulnerable) households in regions across the USA, paying particular attention to transit and pedestrian networks.

In contrast, only a few papers in the literature frame transit as a 'car-free' option and compare transit to car access. This framing is notable as car travel can be seen as a direct competitor to transit or as a benchmark for travel times and accessibility levels [@golubUsingPrinciplesJustice2014; @martensJusticetheoreticApproachDistribution2012]. As an example, @warrenDevelopingEquitableSustainable2015 propose a standard for per capita car ownership for Havana (Cuba), in recognition that car mobility is needed to alleviate transportation disadvantage in the short-term where public transit has yet to be sufficiently addressed. In this context, @warrenDevelopingEquitableSustainable2015 acknowledge the tension between household vulnerability and their need for mobility against GHG emission reduction goals and car-dependency cycles. However, not all papers see transit as a direct competitor, but as a mode that can be used to satisfy individual capabilities. For instance, @smithAccessibilityCapabilityMinimum2012 focus on rural areas in the UK and various types of households (e.g., retired, no-children, with children, single, etc.). In this investigation, perspectives about minimum transport needs and costs for a variety of living standards are explored. The papers reviewed vary in the importance they place on climate urgency, with some focusing more on satisfying *all* sufficient individual needs while planning for less car-dependent cities in the future.

Following a focus on transit, a focus on pedestrian travel is the second most common object of justice considered in the literature. Pedestrians are the most extreme case of confluence between the "what" and the "who", since the mobility tool is the person's own body, with its inherent range of abilities. In papers that focus exclusively on walking, many use or develop walkability scores to explore perceptions of neighborhood quality [@evansAccessibilityUserNeeds2015] or pedestrian mobility with a focus on middle-aged and older adults [@towneUsingWalkScore2016], gender [@kimNeighborhoodEnvironmentWalkability2016], or urban peripherical regions [@blecicCapabilitywiseWalkabilityEvaluation2021].
These papers use 'walkability' as a way to measure the equity in its distribution. Other papers use walkability as an indicator for public health and urban vitality [@sungResidentialBuiltEnvironment2015; @mccormackRelationshipClusteranalysisDerived2012].

Additionally, papers that focus on pedestrians also often see walking as a bridge to connect multiple modes: they often discuss 'walkability' as part of active transportation, which focuses on both walking and bicycle and/or transit. Concepts discussed include how active transport contributes to children's physical activity levels [@mammenSchoolTravelPlanning2014], walkability as an alternative to car predominance [@bertrandMeasuringMappingDisparities2008] or tension that exists between modes, creating unsafe conditions for walking [@siuAssessmentPhysicalEnvironment2019; @ferenchakEquityAnalysisProactively2019].

In contrast to transit and walking, cars are seldom the only mode examined within a paper. When car as a mobility tool is studied, it is often used as a comparison with transit or as the only mode of transport for areas with sub-standard transit systems e.g., [@kimmelStructuralBarriersComprehensive2018; @aljoufie2016]. Similarly, a number of car studies focus on externalities such as air pollution or safety [@fengTradeoffsMobilityEquity2014; @houstonProximityLicensedChild2006].

```{r calculating-destinations-of-interest}
dest_count <- strsplit(Consensus_only_analysis$`_activity`, "; ",
                       fixed = T)

dest_count <- data.frame(
  id=rep(Consensus_only_analysis$`_activity`, lengths(dest_count)),
  dest_count=unlist(dest_count))

dest_count <- dest_count |> group_by(dest_count) |> summarize(count = n()) 
```

## "What": the benefits of mobility and accessibility

The most fundamental benefits of transportation systems are mobility (enabling or impeding movement) and accessibility (the ease of reaching destinations). These benefits, however, are not necessarily valued by themselves, but rather are seen as instrumental to achieve an ulterior goal (e.g., activity participation). For example, although vehicle kilometers travelled (VKT) is still sometimes seen as a useful policy instrument [@zhaoRethinking2021], travelling more is not necessarily a sign of advantage when accessibility is low [@paezAccessibilityHealthCare2010], and similarly short trips may be a sign of advantage [@parkTransitParksEnvironmental2021]. For this reason, although the right to the road (and to transportation systems more generally) is important, the literature leans heavy on the ulterior object, namely accessibility to destinations.

In this respect, we find that most papers tend to be wide sighted, in that they do not focus on a particular destination (e.g., `r (dest_count$count[8]/sum(dest_count$count)) |> percent()` of studies). Within these papers, a variety of equity dimensions, often in combination with different modes of transportation. Typically, they are multi-modal and focus on factors that impact the **trip itself** (e.g., the trip experience, the quality of infrastructure, the level of service) or on the **people and relevant destinations that can be accessed** (e.g., a bundle of trips made by specific population groups, and whether they are enough for 'sufficient' quality of life). For the first, the focus is on the quality of infrastructure, safety issues, perceived accessibility and facets of the level of service such as frequency [@zheEvaluationSharedUse2008; @prasertsubpakijEvaluatingAccessibilityBangkok2012; @furstBestBadPractices2013 ; @lattmanPerceivedAccessibilityPublic2016]. For the second, the focus is on bundles of trips tailored for specific population groups, such as women [@russellPedallingEquityExploring2021] or people who face physical disabilities [@wilkinson-meyersLiveOrdinaryLife2015]. More broadly there is the question of what is enough for the quality of life to be 'sufficient' [@churchillTransportPovertySubjective2019]. These papers further demonstrate the multi-dimensional role of transportation systems: they provide a utilitarian service that can be used to get from A to B but they too are experienced by the people that use them. Papers that examine 'all trips' best exemplify this trend in the transportation equity literature.

In terms of papers that study specific destinations, the activities most commonly studied include healthcare services (`r (dest_count$count[6]/sum(dest_count$count)) |> percent()`), followed by employment (`r (dest_count$count[5]/sum(dest_count$count)) |> percent()`). Papers that exclusively focus on healthcare typically originate from the healthcare planning literature, and look to inform planners about disparities in services and corrective actions. For instance, @wangModelingSpatialAccessibility2011 model access of Chinese-language speakers in Toronto (Canada) to Mandarin-speaking family physicians. These authors infer an inequity from the spatial mismatch between language competent healthcare providers and Chinese speakers. Papers that exclusively focus on employment typically focus on these trips as they are the most common trip purpose and are often correlated with other trip activities like shops, recreation, and other services. For instance, @allenSizingTransportPoverty2019a propose a low employment-based accessibility threshold and a composite population vulnerability index to identify transportation-poor neighborhoods in urban Canada. Papers that focus on healthcare and employment typically source data from representative travel surveys/diaries, census data, and points-of-interest databases. In other words, these studies often benefit from well developed and institutional data that represents 'typical trips', especially in the Global North where these data are more readily available.

<!-- DATA AVAILABILITY AS A MATTER OF JUSTICE: IS THIS AN ADDITIONAL CALL FOR THE NEXT SECTION? IN FACT, THE NATIONAL TRANSPORT POVERTY SURVEY EXPLICITLY AIMS TO ADDRESS THE LACK OF DATA FOR PARTICULAR POPULATIONS IN CANADA! The relevant issues are not new, and have at least two parts (i.e., who counts? Who is measured? Who is a subject of justice? And, who gets to use the data?). With respect to the first part, who we count and how we count them has implications for our understanding of just who is a subject of justice, as illustrated by the history of racial classifications in the US [see @leeRacial1993]. In the case of transportation, the issue of data (collection/availability) as a matter of justice is gaining traction as digitalization of data casts a starker light on these questions...for examples, see the papers by @sourbatiSmart2021 and also, yes, by @behrendtData2023 (with Mimi Sheller).
-->

But what about non-healthcare and non-employment activity types? Papers that focus on other activities are not framed as 'typical travel patterns' and they often have different intentions. For instance, papers that focus on shopping destinations, such as grocery stores or markets (`r (dest_count$count[7]/sum(dest_count$count)) |> percent()`), frequently aim to identify food deserts [e.g., @choiFoodDesertsActivity2013; @jiaoHowIdentifyFood2012; @mckeyCrowdsourcedMappingHealthy2020; @kimAssessingSocialSpatial2020].
Papers that focus on educational facilities, including primary, secondary, and post-secondary schools, represent only `r (dest_count$count[3]/sum(dest_count$count)) |> percent()` of the studies, and examine children's active transportation to school [e.g.,@laszkiewiczChildrenGreenWalk2020] or universal design [e.g., @larkinsAccessibleTransportationBuilt2011].
When green space or other places of leisure is the exclusive focus (`r (dest_count$count[6]/sum(dest_count$count)) |> percent()` of papers), studies examine different accessibility questions such as the spatial distribution of green space e.g., [@xuSocialInequalitiesPark2017], for whom its accessible to e.g., [@mavoaAreaLevelDisparitiesPublic2015], and why e.g., [@mearsUnderstandingSocioeconomicEquity2019]. Very few reviewed papers include 'community' destinations (including public service centres, places of community support, and places of worship (`r (dest_count$count[2]/sum(dest_count$count)) |> percent()` of studies) or childcare activity types (`r (dest_count$count[1]/sum(dest_count$count)) |> percent()` of studies). In the few papers that do include them, these destinations are considered in a holistic representation of activity participation [@albertsRebuildingWomenLivelihoods2016; @smithAccessibilityCapabilityMinimum2012]. The lack of information about community destinations is particularly pronounced in the case of children, as per @desjardinsChildren2022.

The literature offers a variety of conclusions and recommendations that span across the benefits of transportation (e.g., destinations types), often comingled with population groups and the tools of mobility. In addition, as we will discuss next, there is a variety of perspectives in terms of the philosophical foundations and the standards of fairness.

<!--

EDITED UP TO THIS POINT.

-->

## "How": concepts of justice and standards of fairness

```{r calculating-concepts-and-standards}
conc_count <- strsplit(Consensus_only_analysis$`_equity conceptualization`, "; ",
                       fixed = T)
conc_count <- data.frame(
  id=rep(Consensus_only_analysis$`_equity conceptualization`, lengths(conc_count)),
  conc_count=unlist(conc_count))
conc_count <- conc_count |> group_by(conc_count) |> summarize(count = n())

stand_count <- strsplit(Consensus_only_analysis$`_type of equity standard2`, "; ",
                       fixed = T)
stand_count <- data.frame(
  id=rep(Consensus_only_analysis$`_type of equity standard2`, lengths(stand_count)),
  stand_count=unlist(stand_count))
stand_count <- stand_count |> group_by(stand_count) |> summarize(count = n()) 
```

```{r plot-standards-concs-perc, warning=FALSE}
#| label: fig-fig6
#| fig-cap: "The proportion of equity standards (vertical axis) within each type of equity conceptualization (horizontal axis) category."
standards_conc_percentages_plot_and_table
ggsave("standards_conc_percentages_plot_and_table.png")
```

`r word_comment("Answering: what conceptualizations occur in papers with opportunity standards? what types of methods?", ".")` Broadly summarizing how equity standards connect with conceptualizations, some trends emerge in methods used. **Opportunity** and **population** standards appear in the literature at similar proportions ( `r (stand_count$count[3] /sum(stand_count$count)) |> percent(accuracy=0.1)` vs. `r (stand_count$count[4] /sum(stand_count$count)) |> percent(accuracy=0.1)`), but frequently correspond to different conceptualizations. Referring to @fig-fig6, over 40% of papers that focus on **horizontal equity** and **spatial equity** conceptualizations suggest **opportunity** standards. Within these papers, travel impedance standards (a type of opportunity standard) are often suggested and accessibility indices are operationalized (i.e., a measure of the potential to interact with opportunities for populations located at each spatial unit within a region). Examples include: @chenImpactHighspeedRail2017 use a travel time threshold of 4 hours or less on high-speed rail from one municipality to another to be considered "comfortably connected", @yenisettyMeasuringAccessibilityVarious2020 assumes that populations living in areas where the travel distance to a public transit station is less than 1,200m is sufficient to interact with the transit system, and @shenMeasuringSpatialAccessibility2020 identifies regions where populations cannot access hospitals within 1 hour by car. Papers suggesting opportunity standards often employ disparity analysis through a variety of quantitative approaches such as inequality measures (e.g., Gini coefficient and Lorenz curves, poverty measures [@vanderveenOperationalizingIndicatorSufficient2020; @tiznadoaitkenRoleAccessibilityPublic2018]), spatial descriptive analysis, and the comparison to benchmarks (e.g., equal supply to demand of public transit in a spatial unit as done by @peungnumsaiGridBasedSpatialAnalysis2020) to determine which locations are spatially and horizontally (in)equitable relative to other locations in the studied region. Another branch of quantitative research conceptualize transportation system externalities as trade-offs, maximizing transport-related benefits (i.e., time savings, emissions reductions, congestion reductions, user fares) through optimization/location-allocation methodologies e.g., [@fengMulticriteriaEvaluationAccessibilitybased2014; @fakhrmoosaviIncorporatingTravelTime2021; @zhengAreabasedEquitablePricing2020; @wismadiSpatialPreferenceModelling2014]. These papers that focus primarily on **horizontal equity** and/or **spatial equity** seldom use exclusively qualitative methods.

`r word_comment("Answering: what conceptualizations occur in papers with **population** standards? what types of methods?", ".")` Whereas, papers that propose  **population** standards frequently utilize other methods.
Over 40% of papers that suggest **population** standards focus on **well-being** conceptualizations.
These papers typically ask *what is enough to lead a satisfactory life (as related to transportation)*, and the standards that are suggested include population benchmarks for comparison such as: questionnaires and relative comparisons to physical activity per week recommendations [@adlakhaMindGapGender2020; @auchinclossDesignBaselineDescription2020; @mccormackRelationshipClusteranalysisDerived2012; @kimNeighborhoodEnvironmentWalkability2016; @towneUsingWalkScore2016], summative per capita benchmarks (e.g., energy consumption for a 'decent living' is suggested in [@raoDecentLivingEmissions2012], and region-relative comparisons in health-related outcomes e.g., premature mortality rates [@awuorInfluenceEnvironmentalHealth2019], spatial access to hospitals [@pereiraGeographicAccessCOVID192021], spatial access to supermarkets, active-mode-usage, and Body Mass Index (BMI) [@murphySupermarketAccessTransport2017].
The majority of these papers use quantitative/mixed-methods to identify inequities in **wellbeing**, however, a minority do use exclusively qualitative methods to distill themes e.g. the exploration of *perceived* quality of life in [@berheAdaptationDissonanceQuality2014].

`r word_comment("Answering: what conceptualizations occur in papers that conceptualize BOTH population and opportunity standards? what types of methods?", ".")` Papers that conceptualize *both* **population** and **opportunity** standards often conceptualize **vertical equity** and **transport-related social exclusion** (note the similar proportions in these standards in @fig-fig6).
Expectedly, these papers often include a combination of methods: Questionnaires and other qualitative methods related to population standards and quantitative methods such as accessibility indices for opportunity standards are usually deployed.
For instance, census data and the estimated proportion of households within some travel distance/time/availability to/of key destinations is used to identify a variety of social exclusions e.g., [@mackettIncreasingAccessibilityCostEffectively2010; @chenExploringTravelCharacteristics2010; @danielsProposalAccessibilityPlanning2011; @sunPublicTransportAvailability2021; @sharmaPublicTransitAccessibility2021], transport-related social exclusion's link to wellbeing e.g., [@delboscTransportProblemsThat2011; @churchillTransportPovertySubjective2019], areas more likely to experience transport poverty [@allenSizingTransportPoverty2019a], food deserts [@mckeyCrowdsourcedMappingHealthy2020], or transport-related energy poverty [@robinsonDoubleEnergyVulnerability2020; @berryInvestigatingFuelPoverty2016; @berryDistributionalEffectsCarbon2019].

Similar to papers that conceptualize **wellbeing**, the majority of **social-exclusion**-conceptualizing papers use quantitative/mixed-methods to identify areas, households, and/or populations at risk. They use a variety of methods to identify _where_ populations at risk may be located, such as clustering methods [@mohriClusteringMethodMeasuring2021]. The minority of papers that employs exclusively qualitative methods use surveys to understand population travel willingness/barriers e.g., [@chenExploringTravelCharacteristics2010; @mehdizadehWalkingTimeSchool2017] and interviews/focus groups related to topics of unmet activity needs [@nordbakkeTransportUnmetActivity2015].

`r word_comment("Answering: what conceptualizations occur in papers that conceptualize infrastructure standards? what types of methods?", ".")` The largest group of papers that suggest **Infrastructure** standards conceptualize **Rights**. Recall, **Infrastructure**-standards suggesting papers represents `r (stand_count$count[3]/sum(stand_count$count)) |> percent()` of all papers and **Rights-**conceptualization is represented two times more than any other conceptualization type. These papers often focus on populations with mobility impairments and non-car users' inequities in their ability to access the transportation systems. Methods vary, but focus equally on audits of existing infrastructure relative to best-practice universal design principles [@odeckEconomicAppraisalUniversal2010; @larkinsAccessibleTransportationBuilt2011; @jimenez-espadaResearchProblemUniversal2021; @perezdelhoyoParticipatoryManagementImprove2021] and qualitative methods that interview/survey users about their perceived access to transport systems e.g., [@marquezFactorsAffectingPersonal2019; @mateo-babianoBicycleSharingAsia2017; @furstBestBadPractices2013; @velhoEffectTransportAccessibility2016; @parkJourneyVisuallyImpaired2017; @limFacilitatingIndependentCommuting2021; @stjernborgAccessibilityAllPublic2019] and experiment the suitability of existing best practice standards e.g., [@daamenAssessingGapPublic2008; @velhoEffectTransportAccessibility2016; @bharathyRevisitingClearFloor2018].

More broadly, papers that suggest **Infrastructure** standards are often multi-dimensional, and extend beyond **Rights** conceptualizations. They often suggest **Opportunity** and **Population** standards as well and often apply **Vertical**, **Horizontal** and **Spatial** equity lenses as well. These multi-dimensional papers can refer to established guidelines and suggest composite indices e.g., @racheleNeighbourhoodSocioeconomicTransport2017a combines transport network properties such as street connectivity, cul-de-sac length, street block length, traffic volume, public transport stops and service frequency inputs to define an indicator of transport design that is supportive of walkability and access to public transport. Other works include assessing the quality of infrastructure [@xuSocialInequalitiesPark2017], the severity and frequency of accidents on the system  [@benevenutoExaminingTransportNeeds2020; @appleyardTransitCorridorLivability2017], user-groups (particularly disadvantaged groups in the case of vertical equity conceptualizations) [@prasertsubpakijEvaluatingAccessibilityBangkok2012] as part of the multi-criteria indicators. Another branch of literature explicitly focuses on affordability or other barriers to the transport system, and suggests improvements to the infrastructure such that all groups (especially the most disadvantaged) can sufficiently interact with the system e.g., [@basuPracticalFrameworkBenchmarking2019; @songWomenWheelsGender2019; @welchEquityTransportDistribution2013], dabbling into conceptualizations of **transport-related social exclusion** e.g., [@kentPrioritizingLowstressEquitable2019] and **sufficientarian/capabilities** conceptualizations e.g., [@smithAccessibilityCapabilityMinimum2012].

`r word_comment("Answering: what conceptualizations occur in papers that conceptualize Environmental+ standards? what types of methods?", ".")` Though papers that conceptualize **Environmental +** standards are not common in the literature review (`r (stand_count$count[1]/sum(stand_count$count)) |> percent()` of all papers), they most frequently occur in papers focused on **Inequitable externalities**. These papers often use traffic-related air pollution, noise pollution, green-space, urban design elements, urban air temperature, health related outcome, and physical activity guidelines to quantify transport-related externalities. Methods used are almost all quantitative or mixed-methods, and the identification in inequalities is spatial clustering, the use of Gini coefficient [@fengMulticriteriaEvaluationAccessibilitybased2014], comparisons to established environmental thresholds or health guidelines [@agostfelipInclusiveModelAssessing2021; @agostfelipInclusiveModelAssessing2021; @kruizeEnvironmentalEquityRole2007; @iungmanImpactUrbanTransport2021; @apparicioCyclingOneMost2021; @khomenkoLiveableCityHealthy2020; @muellerSocioeconomicInequalitiesUrban2018], creation of composite multi-dimensional indices [@agostfelipInclusiveModelAssessing2021; @mirandaBenchmarkingSustainableUrban2012; @corazzaMethodologyEvidenceCase2020], and/or in addition to spatial analysis [@jephcoteGeospatialAnalysisNaturally2013; @carrierApplicationThreeMethods2014].

See Table XX for detailed examples of papers conceptualizations and standards across each equity dimension.

# Moving forward: discussion and recommendations

> "_Justice is itself the great standing policy of civil society; and any eminent departure from it, under any circumstances, lies under the suspicion of being no policy at all._"
> — Edmund Burke

We discuss trends in how the literature has conceptualized and used standards (the "How") to evaluate the dimensions (the "What") of transportation equity. A noteable point is the trends are context-specific: they concern the "When" and "Where" (differences across publication year, across case studies continent, and between urban and rural contexts) and the "Who" (differences in the focus on the population sub-groups, transport modes, destinations of interest) as discussed in the preceding sections following the four equity dimensions that structure this review (mobility/accessibility, traffic-related pollution, human-health related, and traffic safety).

To broadly summarize what we found in the literature:
<!--
- the "When" and "Where": 
  - most studies focus on case studies in north american and europe
  - south american/africa papers have key differences in focus compared to global north: they are more focused on affordability-as-a-barrier, tensions in infrastructure dev vs. service prioritization, data availability limitations
  - 
- the "Who":
  - low income most popular
   - (dis)abilities are exclusive focused population category in many studies; other 'vulnerabilities' are typically composite (aside from income, which is part of both exclusive-focus and composite vulnerabilities)
  - transit, with cycling and walking becoming more popular
  - cars are seldom the focus, and if they are, it is as a comparison to transit and/or focus on their negative externalities
  - destinations of interest are mostly employment and healthcare; other destinations are less studied. The many 'destination' however, is the trip itself -- the experience, the infrastrucutre, the mode-service, a bundle of trips taken a month to all locations, etc. 
- the "What" and "How":
  - accessibility/mobility -> opportunity standards are typically travel impedance standards, and affordability thresholds (e.g., more than 10% of monthly income should not be spent on transport [@rivasHowAffordableTransportation2018] to work and/or other necessary destinations). Oftentimes spatial equity, disparity analysis is operationalized. Infrastructure standards are also suggested -> oftentime Rights conceptualizations, i.e. access for all.
  - transport-related environmental externalities -> pollution/noise standards are suggested, often using WHO guidelines or local neighbourhood averages as a comparision.
  - health  -> health standards, the reliance on population transport health literature such as physical activity guideline recommendation per week or excess mortality burden. 
  - safety -> the use of engineering standards and best practice
  - cross-dimensional -> measures such as urban livability, etc. a more holistic conceptualization of transport equity
-->

**The "When" and "Where"**:

The majority of papers (`r ((case_cont_count$count[4] + case_cont_count$count[5]+ case_cont_count$count[6])/sum(case_cont_count$count)) |> percent()`) focus on case studies in the Global North. Though their subject matter is varied, their spatial context is one situated in case studies in North American and Europe, and thus more often than not, in conversation with more developed and formal government transport planning apparatuses and technologies (e.g., planning for equitable high-speed rail [@monzonEfficiencySpatialEquity2013], autonomous vehicle technology [@eppenbergerOpportunitySharedAutonomous2021], public consultation processes [@reddySafeguardingMinorityCivil2010]).

However, studies from the Global South, specifically South America and Africa, have key differences in focus compared to Global North: they are more focused on affordability-as-a-barrier, public spending tensions in developing new transport infrastructure compared to transport service prioritization and more intense data availability limitations. In this way, they do not engage with emerging mobility technologies with large capital costs. Informality in transport planning is more present. In the present day, these countries, compared to the Global North, are underdeveloped for reasons of past colonialism under Northern states, heavily reliant on primary sector exports (lower efficiency, lower national GDP) under growing global financial markets, and have more fragile democracies. Because of lower data availability and more extreme needs for 'sufficient' transport, analysis of transportation inequities are more focused along economic lines. For instance, public transit network being improved such that those that at least 60% below the poverty line have at least the same amount of accessibility (60 min travel time to destinations) to jobs, educational facilities and services as the total population using public transit [@basuPracticalFrameworkBenchmarking2019].

Temporally, the Global North and Global South studies can be generally seen on different transport inequity continumums. Global North -> concerned with particularly disadvantaged groups and creating indicators that may be used to guide the remedy of formal processes that ensure access to societal benefits [@cuiSpatialAccessPublic2020]. Global North countries often surpass international sufficiency thresholds related to transport and public health (e.g., @carrierApplicationThreeMethods2014 measure NO2 road-side air pollution and find no levels are below the WHO guideline) but advocate for the reduction in disproportional impact of air pollution on lower-income groups. New standards are needed that are context-specific. What's also missing? We call for: _discussions of equality_. The Global North inequities are rising -- the richest have the most mobility, this is missing from the discussion. We need new standards that reflect this.

In reference to transport inequity continuum, papers on the Global South tend to operationalize international standards as guidelines. Though not covered in this review, Global South nations' formal processes for planning are newer (relatively), more fragile, and operating under more strict financial constraints than the Global North. Informal processes are thus more important to account for equity (e.g, informal transit [@friedMeasuringSustainableDevelopment2020], populations living in informal residential locations [@sharmaPublicTransitAccessibility2021]). Comparisons to thresholds set by governing international bodies are relevant as minimums are relatively lower (e.g., traffic related pollution, access to basic healthcare). Global South, along some dimensions, are lagging Global North development, but have the opportunity to plan better and not repeat Global North mistakes (e.g., car-centric development [@warrenDevelopingEquitableSustainable2015]). More work on this is needed. Additionally, what's missing? Similarly, _discussions of equality_. Inequities look different than in Global North, but systems thinking is required as global forces influence the capacity of formal transport planning processes?

**The "Who"**:

Though the literature reviewed most commonly considers low-income as a socio-demographic characteristics of populations deserving equity as either by itself or alongside other characteristics (`r prettyNum(round(df_pop$count[8]+df_pop$count[4] * 100, 2))`\% of the papers). Many employ accessibility methods (e.g., walking accessibility to open space [@tangWalkingAccessibilityNeighbourhood2021]) but a others use qualitative or mixed methods to gain perspectives of inequalities that are typically cross-dimensional (e.g., @milanLiftingPeripheralFortunes2017 asks Medellin residents living in TOD areas compared to non-TOD areas about the impact of TOD on their wellbeing based on income group and gender). The use of qualitative methods allows for a more nuanced reflection of residents perspectives, and different inequities emerge.

From a modal perspective, transit is also the primary focus of the literature; cycling and walking the next most popular focuses. These modes use public-space and users can interact with the public when using them (unlike car modes, which are seldom the focus of the reviewed literature). Upon the rise of automobility, transit was seen as a public service and something getting in the way of automobility. But historically, mass transit and active modes are how people get around - and automobility is in direct tension with the movement of other modes as it requires a lot of pubic land and public subsidy. This may be one reason why transport inequity research focuses on these modes, often car modes as driving inequalities in many dimensions (e.g., access/mobility, traffic-related pollution, traffic-safety, and human-health).

In the literature reviewed, destinations of interest are mostly employment and healthcare; other destinations are less studied. For many papers, the 'destinations' however, are the trip themselves -- the experience, the quality of the infrastrucutre, the mode-service, a bundle of trips taken a month to all locations, etc.

So what's missing from the "who"? More nuisances perspectives that capture the "who" of inequities. As shown in the variety of methods used, dimensions analysis, and conceptualizations and standards, transport inequalities are multi-dimensional. How decision makers define an equity-deserving community will impact results (e.g., [@rowangouldIdentifyingEnvironmentalJustice2015]). In this sense, standards that are defined need to be sensitive to changing community-based definitions of inequities. Are issues of economic-inequity at the root of transport inequities for a specific community? Are (dis)abilities? Are inequities in the service of transit the focus of a study because it can be improved to address transport inequities and are alternative modes driving these inequities? Access to what sorts of opportunities is driving transport inequities? How do populations, transit modes, and opportunities saught intersect to define the "who" of inequities? Community-based informed understanding of inequities are needed, and tracking how they change are needed.

**The "What" and "How"**:

Inequities that focus on accessibility/mobility -> opportunity standards are typically travel impedance standards, and affordability thresholds (e.g., more than 10% of monthly income should not be spent on transport [@rivasHowAffordableTransportation2018] to work and/or other necessary destinations). Oftentimes spatial equity, disparity analysis (spatial mismatch e.g., @mulleyAreNetworkPlanning2015) are operationalized. Infrastructure standards are also suggested -> often times access to destinations is conceptualized as a "right" as such being able to enter the transportation network and physically use it the pre-requisite that accessibility measures do not capture (e.g., (dis)abilities focused).

Inequities that focus on transport-related environmental externalities -> pollution/noise standards are suggested, often using WHO guidelines or local neighbourhood averages as a comparison.

Inequities that focus on human-health  -> health standards, the reliance on population transport health literature such as physical activity guideline recommendation per week or excess mortality burden.

Inequities that focus on traffic-safety -> the use of engineering standards and best practice

Of course, some papers focus on cross-dimensional transport inequity dimensions. Measures such as urban livability, etc. a more holistic conceptualization of transport equity. This is what we think is missing from the literature, and what's needed to move forward to justice.

## Call 1: The need for explicit conceptualizations and grounded standards

Some conceptualizations are implicit within the literature: for example, @muellerSocioeconomicInequalitiesUrban2018 relative risk of mortality as related to transport-related air pollution should not be higher in deprived groups than the general population. It is not evident from the paper what conceptualization drives this focus on human-health inequities but can be implicitly inferred to be general "well being". We believe, more explicit conceptualizations support inequities are needed in the literature.

Some standards are also seemingly arbitrary. For example, @caoIndicatorsSocioSpatialTransport2017 proposes 20 ferries per day to avoid social exclusion for inter-island transport planning in the Philippines in their analysis, but admits that a standard should be politically determined. It is unclear if, for instance, 10 or 30 ferries would make a difference in a specific quality-of-life outcome or if that number is tied to funding/resource constraints. If literature is to recommend standards, researchers need standards linked to grounded outcomes. For decision-makers setting standards, measuring inequities and moving towards setting flexible guidelines for standards is the next step. As such, we believe, firmer justification of standards are lacking in the literature and are needed.

Relative measures are often used: context, temporal, and group-specific. E.g. a certain group should have more access in certain situations. Context-matters, but this presents a challenge in policy planning.

Equity is not justice. Some studies conflate the two. Justice should not be mixed up with equity. More justice work is needed, e.g., Restorative justice -- repairing harm is missing from the discussion of addressing transport inequities.

## Call 2: Need for creative methods for systems-thinking approach to inequities

On the methodological side, more mixed methods are needed in transport equity research.
Conceptualizations and standards are usually discussed from purely qualitative or quantitative approaches, a missed opportunity to combine the strengths of both approaches, whether by deep diving into some particular experiences or perceptions through qualitative methods or tailoring more meaningful quantitative analysis after qualitative explorations. For instance, @xieWonCycleRoute2018 finds through interviews and go-alongs with women cyclists that the standard Cycling Level of Service (CLS) tools used by engineers to plan cycling infrastructure misses the critical gendered perspective. Further, @somenahalliAgingTransportMobility2007 surveys older adults to understand their mobility issues, revealing factors that are unseen in standard daily travel surveys.

Furthermore, this is plenty of disparity analysis, without engaging explicitly with equity conceptualizations. For example, within the mobility/accessibility dimension, metrics of accessibility (usually 15 to 60 minutes) are used to show differences among areas and groups but with scarce policy and practice implications of those results.
Aiming at specific goals and standards tied to conceptualizations is the ideal case.
When these analyses engage with metrics that may be tied to conceptualizations (like Gini coefficient or Theil index), they fall short in assessing if the result's good or bad e.g., [@mijaresEquityAnalysisUrban2013].
If a Gini coefficient of 0 means that all people have the same access to public transport stops, what does it mean a 0.2, 0.3, or 0.4? Is this good or bad news for decision-makers? Are new policies needed to reduce that number to a certain threshold, orienting future interventions? These questions usually remain unanswered despite its importance. These measurements can also bring some challenges and pitfalls, as recently summarized by @karnerAdvancesPitfallsMeasuring2023.

## Call 3: Data justice

<!-- see my comments above -->

What are the sources of data and what are the motivations for some categories?
POI databases typically include education, health and aggregated categories for leisure and community.
Within 'community' are organizations, government services, visiting friends/family -- very broad, grandma's house is not there, social networks rarely incorporated.
Childcare, typically daycare or facilities -- domestic work, mobilities of care, mobility interdependence, are unrepresented relative to the presence of work destinations. Transport systems' focus is more than just to work or as a source of economic development, though in underdeveloped regions, transport systems as a force of economic development e.g., high-speed rail [@chenImpactHighspeedRail2017; @monzonEfficiencySpatialEquity2013; @kimImpactsHighspeedRail2015].

Data availability matters, especially when operationalizing emerging theories (i.e., sufficientariansm [@vanderveenOperationalizingIndicatorSufficient2020]):

- Leisure destinations (e.g., green space, parks, recreation) are less studied in this context.
- Some categories are missing all together -- mobilities of care.
- Issue of data availability

## Call 4: More direct and explicit links between standards and experienced outcomes

A more robust assessment of the implications of equity standards on life outcomes is still pending.
Estimating the benefits of increased mobility or accessibility, or reducing affordability burdens and transport externalities needs to be associated with outcomes like life and neighborhood satisfaction, subjective well-being, mental and physical health, social capital, among others.

## Call 5: Evaluations of interventions and policies

There is a need to evaluate more equity interventions or policies.
In our review, only 19 out of 155 studies assess specific projects with an equity lens <!-- ignacio says CHECK THIS, make sure that this is the correct number-->.
Examples include mode-shift from driving to active school travel [@mammenSchoolTravelPlanning2014], transit fare restructures [@hickeyUsingQuantitativeMethods2010] and spatial analysis of Low Traffic Neighborhoods [@aldredEquityNewActive2021].
This is a key step towards transport justice; assessing the effects of policies on different dimensions and populations groups and evaluating if a specific context is moving towards equitable standards over time.

<!--
THESE THREE PARAGRAPHS SOUND LIKE A CONCLUSION. PERHAPS A CALL TO ACTION - TO BE ABLE TO MAKE AN INTERNATIONAL COMPARISION ON JUSTICE. WE NEED TO KNOW THE WHY BUT LITERATURE SHOULD START TO OUTLINE THAT EXPLICILITY.
-->

For the purpose of this review focused on *equity*, the focus is on studies that substantively answer "Where?", "When?", "Who?", "What?", and "How?", while the answer to "Why?" may occasionally be incidental.

Transportation equity literature exists within cross-cutting categories.

Though critical, answers to "Why?" are not overwhelming common in the transportation equity academic literature that also address  "Where?", "When?", "Who?", "What?", and "How?". This is discussed as a next step in planning for *justice* transportation literature in Section 6.

# Concluding remarks

> "The way to right wrongs is to turn the light of truth upon them._" 
> — Ida B. Wells

"Justice is a constant struggle" we need equity frameworks that support the movement for justice. Understanding and tracking transport inequities over time and space; specifically "who", "what", "where", "how" and "why" are integral. Then set policies that address these inequities but stay flexible to changing demands of justice.

Flexible frameworks are important. As mentioned, transport is derived demand – what’s everyone’s demand look like? Past, current, and future? Creating transport equity frameworks that lend themselves to changing as the struggle for Justice moves forward and is resilient to backsliding is crucial. Setting standards from a systems approach: both considering the _positive_ rights (a right to have access to sufficient quality of essential services, for example) and _negative_ rights (mobility of cars must be limited as not to impact air quality, health, safety, and positive rights) must be both be conceptualized. Systems thinking is lacking transport equity literature.

We thus set out the following calls for transport system decision makers in defining equitable transportation standards, particularly in collecting data that can be used to highlight and track inequalities such that standards can be set.

1. **Understand what transport inequities you are measuring**. Common measures used in transport systems are accessibility measures. What does the measure mean and does potential access relate to experienced outcomes? Next, what does an un equitable and equitable arrangement look like ? This is context specific and locality sensitive. Global north has different standards than global south.

2. **Understand what institutional context we’re operating in**. For example, what is the structure of transport planning. What Does WHO guidelines make sense to use as a standard for the community's context?

3. **Understand what institutional context transfer looks like**. A standard set in one context, may not transfer to another community in the same region.

4. **And be clear with our terms – what is equity and for whom?**. If '15 minute city' is the policy goal, what conceptualization is driving this standard? Is it sufficientarism? Egalitariansm? The standard can be interpreted a multitude of ways. The conceptualization needs to guide the standard. Being explicit about the conceptualization is important.

# Appendix

## Search strategy

```{r Search-query-A1}
#| label: fig-A1
#| fig-cap: "The search query. TS = topic search (keywords, abstract, title). TASCA = subject categories. Green text area transportation system related terms, blue text are equity dimension related terms, purple text are equity/justice conceptualization related terms, and orange text are standards related terms. Hits corresponds the number of papers that the search yielded and was retained into the evidence selection process."
knitr::include_graphics("figures/Search-query.png")
```

Definitions of the population-concept context (PCC) used in the creation of the inclusion and exclusion criteria for the search strategy.

-   **Population**: the focus of the included studies should be on individuals, groups, communities, or entire regional areas that are impacted by passenger transportation infrastructure and systems (i.e., all modes and flows) from the perspective of equity (i.e., fair distribution, production, and re-production of burdens and benefits). This criteria is reflected in the creation of the first set of topic search terms that relate to transportation modes (e.g., "walking" OR "cycling" OR "transit" - see green text in @fig-A1 for the full list).
-   **Concept**: the included studies should also include equity dimensions and conceptualizes equity as discussed in the previous section. This inclusion criteria is reflected in the second and third set of topic search terms developed in the search strategy. These terms relate to types of equity dimensions (e.g., "accessibility" OR "mobility" or "transport-related air pollution" - see blue text in the @fig-A1 for the full list) and equity conceptualizations (e.g., "Justice" OR "equity" - see purple text in @fig-A1 for the full list).
-   **Context**: the included studies should also be limited to publications that include equity standards. Context can be more difficult to explicitly search for with key terms so synonyms for 'standards' were added to the query as a four set of topic search terms (e.g., threshold, indicator, criteria - see orange text in @fig-A1 for full list). Additionally, journal article and conference papers, English-language literature from any country, any study design (e.g., quantitative, qualitative, or mixed-method studies, or conceptual frameworks), and any record published within the past 30 years are included (January 1992 to March 2022). The time period is selected as the first (to the authors knowledge) peer-reviewed article which operationalized equity standards and equity conceptualization was published in 1996 [@khistyOperationalizingConceptsEquity1996]; we are broadening the search by a few years for completeness. English is selected as it is the common language spoken across the authorship team. Furthermore, papers that explicitly fall within the Transportation or related topic/category is included in the query (e.g., "Transportation", "Social Sciences", "Geography", "Civil Engineering", "Philosophy" - see the @fig-A1 for full query).

The **exclusion criteria** for the search are papers that are not within the inclusion criteria.
Specifically:

-   Literature published before January 1992.
-   Papers which do not include transportation equity dimensions.
-   Grey as concepts contained within are frequently published in a more developed form in journals.

## Example of the data extraction template:

```{r data-extraction-template-A2}
#| label: fig-A2
#| fig-cap: "The data extraction template with associated defintions."
knitr::include_graphics("figures/Data-extract-template.png")
```

## Examples of papers summarized by element of the analyitical framework

```{r}
dat <- data.frame(matrix(c("What?: **mobility and accessibilty**",
                           "Where? [@rivasHowAffordableTransportation2018] - South America (select cities)",
                           "Analyses how affordable urban public transportation is in select Latin American and Caribbean countries. They look at the estimated average monthly cost of transit trips and average monthly household income and conceptualize **transportation-related** **affordability**, especially for the most economically vulnerable (**vertical equity**).",
                           "How?: The financial burden of a basket of urban public transportation trips (60 trip fares, representing 30 round-trips per month) should not exceed 10% of household monthly income.",
                           "What?: **mobility and accessibilty**",
                           "Where?: [@bharathyRevisitingClearFloor2018] - North America (USA - National)",
                           "This study designed a web-based tool and took a representative sample of wheeled mobility device (WhMD) users anthropometry measurements to determine if the minimum standard suggested by the ADA is sufficient. We understand this conceptualization as a type of **Rights** conceptualization that WhMD should have minimum clear floor space (as described the guidelines in line with the American Disabilities Act) to access bus shelters, bus stop pads, and transit terminals.",
                           "How?: The clear floor area for wheelchairs: 760 mm (30 in.) wide by 1220 mm (48 in.) in length as described by the ADA standards. Of note, this minimum clear floor area is insufficient for a variety of the WhMD users.",
                           "What?: **mobility and accessibilty**",
                           "[@ryanWhatAreWe2021] - Europe (Stockholm, Gothenburg and Malmo cities in Sweden)",
                           "Investigates what the literature and planning process is missing when we measure accessibility by comparing objective and self-reported accounts of accessibility among older people. This paper conceptualizes accessibility as from the position of the **capabilities approach** and **vertical equity** (particularly acknowledging that older people have capabilities that differ from the general population).",
                           "How?: Specifically for older populations (aged 65+), the following travel distances are suggested as equitable trip lengths to grocery stores per mode: Walking: less than or equal to 1500m, Combined public transit and walking (less than or equal to 1000m (walking element)), Combined car and walking: less than or equal to 1000m less than or equal to 1000m (walking element)), Bicycle: less than or equal to 3000m in addition to travel time threshold of less than 15 mins.",
                           "What?: **mobility and accessibilty**",
                           "Where?: [@wismadiSpatialPreferenceModelling2014] - Asia (Yogyakarta, Indonesia)",
                           "Explores the equitable provision of transport infrastructure provision: an application of Sen's **capability approach**. Conceptualizes equity through Sen's capability approach and spatial equity.",
                           "How?: Areas below the relative poverty line (of its neighbours) can only be located transport resources (i.e., measure in person*kms that can be travelled at car speed, i.e., mobility) based on the following 2 benchmarks (they can be considered, together as the floor/minmum access): 1) Global: standard deviation (SD) distance to mean should be minimized. 2) Local: priority to minimise the differences with its neighbourhood",
                           "What?: **mobility and accessibilty**",
                           "Where?: [@zhengAreabasedEquitablePricing2020] - North America",
                           "This paper conceptualizies equity in the multimodal network (transit, car) being fair toll-pricing across differences in populatins value of time (VOT). VOT is determined based on household income, with lower income households having lower VOT and thus deserving of lower tolls (vertical equity). From this perspective, a utilitarian perspective that seeks to minimize multimodal traffic congestion through introducing toll-pricing based on VOT is implemented.",
                           "How?: suggest that a toll-pricing scheme based on individuals travel value-of-time (lower income people have a lower VOT) is equitable.",

                           "What?: **environmental pollution**",
                           "Where?: [@carrierApplicationThreeMethods2014] - North America (Montreal, Canada) ",
                           "This work examines the statistical association between different social groups and the concentration of air pollutants. They frame their work from the perspective of environmental equity. We interpret the conceptualizations to be along the lines of **inequitable externalities**, **spatial** and **vertical equity** - transport-related air pollution is a product of road transport and it impacts the air of residents in unequal spatial ways. The paper then frames this impact as unfair, particularly from the perspective of disproportionately disadvantaged residents",
                           "How?: The literature suggests that the health implications from the transport-related air pollution from major roadways is most acute at residential distance locations of 200 m or less. Residential locations should not be located within this distance threshold from the perspective of human health. **Environmental+** and **Population standards**: Uses the WHO NO² threshold as a point of comparison (annual concentrations of NO² should not exceed 40 μg/m-3). They argue that even through no neighbourhood, even those disproportionately low income, exceed the WHO limit in this case study, they still suggest that air pollution should not be disproportionately impacting disadvantaged neighbourhoods. It can be interpreted that they use the WHO threshold as a minimum threshold and suggest that air pollution levels should not be impacting disadvantaged populations disproportionately ( a relative population standard)",
                           "What?: **environmental pollution**",
                           "Where?: [@jephcoteGeospatialAnalysisNaturally2013] - Europe (Leicester, UK)",
                           "Geospatial analysis of naturally occurring boundaries in road-transport emissions and childrens respiratory health across a demographically diverse cityscape. Emperically identifies at what distance away from major roadways children are most impacted by transport-related pollution. This is framed in the perspective of children’s **well-being**. Children are at most risk for acute respiratory distress from elevated levels of air pollution, and as such planning should consider this point of public health.",
                           "How?: Finds that children (most vulnerable to air pollution - related to motoized traffic) are most impacted by air pollution within 283 m of a road way. This should be the distance threshold that schools and other childrens facilities are located.",
                           "What?: **health impacts**",
                           "Where? [@adlakhaMindGapGender2020] - Asia (Chennai, India)",
                           "From the perspective of disparity in gendered physical activity, this paper focuses on women's cycling as both transport and exercise. They advocate for all people achieving physical activity thresholds (**horizontal equity**) but prioritize women and especially women in neighborhoods with low-walkability and socio-economic status (**vertical equity**).",
                           "How?: All people should get 150 min of moderate activity a week or 75 min of vigorous physical activity per week.",
                           "What?: **health impacts**",
                           "Where? [@savingmothersDidSavingMothers2019] - Africa (Select urban and rural regions in Uganda)",
                           "The **well-being** of mothers, this paper examines the timely access to emergency obsteric and newborn care for child-bearing aged women in Uganda.",
                           "How?: 2 hours to the nearest facility with surgical capacity with anesthesia services - this threshold is determined through the onset of bleeding to death if a women with obstetric hemorrhage does not receive adequate treatment).",
                           "What?: **health impacts**",
                           "Where?: [@iungmanImpactUrbanTransport2021] - Europe (Madrid and Barcelona, Spain)",
                           "They use environmental pollution guidelines, but from the position of health. They investigate the impact of urban and transport planning on attributable mortality burden in Madrid and Barcelona and its distribution by socioeconomic status . Pre-mature mortality is linked to the exposure to pollution and motorized vehicles (**inequitable externalities**). These externalities should not be impacting people disproportionately (**vertical equity**) and should be even across space (**spatial equity**).",
                           "How?: All minimum thresholds, if exceeded this is inequitable: NO² concentration 40 μg/m³; PM 2.5 concentration 10 μg/m³; Noise 53dB for average 24 hours; Living with 300 m crow-flies distance from at least .5 hectares of greenspace; and a Change of air temperature of at least 1 ⁰C. ",
                           "What?: **health impacts**",
                           "Where?: [@mehdizadehWalkingTimeSchool2017] - Asia (Rasht, Iran)",
                           "From the perspective of children’s **well-being**, assesses the walking time to school. They frame walking to school as health-related.",
                           "How?: perceived walking time to school for students aged 7-9 yrs is 10 mins, and the longer the PWTS the less likely they were to use an active mode to travel to school.",
                           "What?: **health impacts**",
                           "Where? [@murphySupermarketAccessTransport2017] - Oceania (Melbourne, Australia)",
                           "Assesses the relationship between supermarket access and transport mode used, the body mass index (BMI) of the mode-user (**wellbeing**) and the equity in access distribution by income (**vertical equity**).",
                           "How?: all households should be sufficiently active (greater than 150 min and at least 5 sessions) and households should be within 1 km euclidean distance to supermarket (80-90% of the dwellings should meet this). Planners should prioritize socially disadvantaged areas to meeting these standards first.",

                           "What?: **transport-related safety**",
                           "Where?: [@ferenchakEquityAnalysisProactively2019] - North America (Denver, USA)",
                           "Operationalizes and compares an equity analysis of proactively- and reactively-identified traffic safety issues from the perspective of **Spatial equity**, **Vertical equity** and **Inequitable exposure to externalities**.",
                           "How?: standards are suggested for both reactive and proactive analysis. First, the lower the number of collisions on the road with pedestrians/cyclists (i.e., reactive safety analysis), the better. No/minimal inequalities for general population vs. equity seeking groups (high proportion of POC and/or low income in tract). Second, the lower the perceived safety, the better (i.e., if travel to school by ped. or bike is unsafe due to traffic conditions). No/minimal inequalities for general population vs. equity seeking groups (high proportion of POC and/or low income in tract).",
                           "What?: **transport-related safety**",
                           "Where?: [@zheEvaluationSharedUse2008] - Asia (Tokyo, Takamatsu, and Tokushima)",
                           "Evaluates the observed safety of  shared use pedestrian and bicycle paths from the perspective of **well-being**.",
                           "How?: the study suggests that the safety threshold for bicycles and pedestrians to coexist on shared infrastructure is less than 0.5 pedestrians/minute per metre of sidewalk (width) and less than 3.0 cyclists/minute per metre of sidewalk (width). The standard for pedestrian/bicycle share use in terms of hourly traffic volume is less than 26 pedestrians / hour and 108 cyclists / hour for 2m wide sidewalks.",

                           "What?: **mobility/accessibility and health impacts**",
                           "Where?: [@aldertonWhatMeaningUrban2019] - Asia (Bangkok, Thailand) – **Mobility/ accessibility** and **health**",
                           "Establishes short-, medium-, and long-term goals for the city in collaboration with technical leaders within the municipal government for the perspective of **well-being** (urban livability): the standards included in this table relate directly to transportation systems. Indicators are inspired by the Sustainable Development Goals (SDGs) as well other global planning standards.",
                           "How?: 1) Green space: % of residents living < 400 m from public open space, a large park (> 1.5ha), and/or local park, 2) Public transit access: % of residents living < 400 m of a local bus stop and <800 m of train station, 3) Facilities: % of residents living < 400 m of a community centre. The following **Infrastructure standard** is suggested: Canal water quality - dissolved oxygen content of equal to or less than 2.0 mL/L",
                           "What?: **Mobility/ accessibility** and **health impacts**",
                           "Where?: [@berheAdaptationDissonanceQuality2014] - Africa (Mekelle, Ethiopia)",
                         "Examines adaption and dissonance in the quality of life (QoL) of residents. QoL is conceptualized along the lines of **well-being** and aspects of QoL directly tie into transport systems. They conduct a qualitative QoL survey of residents on the topic of three QoL domains: housing quality, access to important destinations, and affordability. They also measure quantitative indicators associated with these domains. We assume the equity goal for this paper is that subjective and objective QoL measures should not be mismatched: as discussed by the authors of this study, subjective QoL is higher than objective QoL the participant is experiencing adaption and in the reverse scenario the participant is experience dissonance.",
                           "How: 1 & 2) Access to primary or secondary education facility, percentage of households living within 1 km or 2km (walking distance), respectively from a primary school or secondary school. 3) Access to health facility, percentage of households within 40 min walking time from a health facility. 4) Access to public transport, percentage of households within a distance of 500 m from a mini-bus stop. **Population standards**: 1) Adequate family income, percentage of households earning more than the official poverty line. 2) Subjective QoL is constructed based on the households level of satisfaction for each of the eight indicators using a six point Likert-scale (1=very satisfied to 6=very dissatisfied).",
                           "What?: **mobility/accessibility, health impacts, and safety**",
                           "Where? [@agostfelipInclusiveModelAssessing2021] - Europe (Castellon, Spain)",
                           "Conceptualizes equity through age-friendly urban spaces that reduce (and eliminate) conditions for **transportation-related social exclusion** for older populations and prioritize those who are economically vulnerable (**vertical equity**). These guidelines are inspired by the SDGs in addition to planning guidelines used national, regional, and local guidelines used in Spain.",
                           "How: 1) Access to facilities needed for old age health. Minimum distance thresholds from the geometric center of neighborhood are suggested: at least: 1000 m from health facilities (600 m or less is preferred), elderly-specific care facilities and shops should be 600 m (300 m or less is preferred). **Population standards**: 1) Certain neighbourhoods should be prioritized above others. From this papers focus on age-friendly urban environment, they suggest that if the neighbourhood has an average old age indicator (i.e., greater than 64 years, and/or greater than 79 years, and/or aging ratio of persons aged greater than 64 relative to 15 to 64 age) should be prioritized. 2) Economic vulnerable and non-civically engaged neighbourhoods should also be prioritized. If the neighbourhood has a lower percentage of civic associations within the neighbourhood than average, and/or household income, and/or a higher than average interventions for dependency and/or social subsidies, they should be priorized. **Infrastructure standards **: 1) Green space: should be at least 10 m2 per inhabitant in the neighbourhood, greater than 15 m2 per inhab. is the goal. 2) As related to sidewalk infrastructure at least 50% of all sidewalks (preferably 75% or greater) should: have a width of 1.5m or larger, ramps should have a grade of 8% or less, be well maintained (free from deficiencies), be paved for pedestrian use, and cover public transit stops. 3) Lighting is critical for traffic-safety and a sense of safety overall. As such, at least 50% roads should: have a min. of 35 lux (road traffic) and 20 lux (pedestrian streets), and adapted traffic lights. 4) Buildings should be age-friendly. As a proxy for the quality of residential living space quality, at least 50% of residential buildings in a neighbourhood should be built within the last 50 years (preferably 75% or more). In terms of physical access into the buildings, at least 10% should have elevators and accessible entrances (preferably 25% or more). **Environment + standards **: 1) Noise at the street level should be less than 55 dB and 45 dB (but preferably less than 50 dB and 40 dB) in the daytime and nighttime, respectively.",
                           "What?: **mobility/accessibility and safety**",
                           "Where?: [@mateobabianoPedestrianNeedsMatter2016] - Asian (Manila, Philippine)",
                           "The perception of pedestrians’ walking environments should be sufficient across 6 themes. Equity is conceptualized around **spatial equity** (equally fair walking environments for all locations) and **rights** (the right to mobility/accessibility for pedestrians)",
                           "How?: perceived pedestrian perception on protection, ease, equitable access, mobility, identity, and enjoyment must be met."
                           ), ncol = 4, byrow = T))
```

```{r}
set_flextable_defaults(font.size= 7,
                       font.family = NULL,
                       text.align = 'justify')

flextable(dat) %>%
  ftExtra::colformat_md() %>%
  compose(
    i = 1, j = c(1,2,3,4), part = "header", value = as_paragraph(c("Dimension", "Continent", "Conceptualization", "Standard"))) %>%
  merge_v(j=1) %>% #merge 'dimensions' column
  rotate(part = "body", rotation = "btlr", j=1, i=c(1:5,6:7,8:12, 13:14,15:18)) %>%
  valign(valign="center", part="body", j=1) %>%
  hline( i = c(5,7,12,14), part = "body") %>%
  set_table_properties(layout = "autofit")

```

<!-- we need to separate by qual and quant, karen asked-->

# References