### 2024-05-15 Journal-manuscript/Journal-manuscript.qmd
Antonio's edits before submission

---
title: "Searching for standards of fairness in the transportation justice literature"
format: docx
csl: apa7.csl
#   tandf-pdf:
#     keep-tex: true
#   tandf-html: default
# author:
#   - name: Anastasia Soukhov
#     affiliations:
#       - ref: MAC
#     orcid: 0000-0003-2402-304X
#     email: soukho@mcmaster.ca
#   - name: Ignacio Tiznado-Aitken
#     affiliations:
#       - ref: TORONTO
#     orcid: 0000-0003-2402-304X
#     email: i.tiznadoaitken@utoronto.ca
#   - name: Matthew Palm
#     affiliations:
#       - ref: UNC
#     orcid: 0000-0003-2402-304X
#     email: palmmatt@unc.edu
#   - name: Steven Farber
#     affiliations:
#       - ref: TORONTO
#     orcid: 0000-0003-2402-304X
#     email: steven.farber@utoronto.ca
#   - name: Antonio Páez
#     affiliations:
#       - ref: MAC
#     orcid: 0000-0003-2402-304X
#     email: paezha@mcmaster.ca
# 
# affiliations:
#   - id: MAC
#     name: McMaster University
#     department: School of Earth, Environment and Society
#     city: Hamilton, ON
#     country: Canada
#   - id: TORONTO
#     name: University of Toronto Scarborough
#     department: Department of Geography and Planning
#     city: Toronto, ON
#     country: Canada
#   - id: UNC
#     name: The University of North Carolina at Chapel Hill
#     department: Department of City and Regional Planning
#     city: Chapel Hill, NC
#     country: USA
# keywords:
  # - transportation
  # - equity
  # - literature review

bibliography: bibliography.bib 
filters:
  - docx-landscape.lua
---

```{r, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r setup}
library(huxtable) # Easily Create and Style Tables for LaTeX, HTML and Other Formats
library(sf) # Simple Features for R
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(gt) # Easily Create Presentation-Ready Display Tables
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(readxl) # Read Excel Files
library(DiagrammeR) # Graph/Network Visualization
library(scales) # Scale Functions for Visualization
library(corrplot) # Visualization of a Correlation Matrix
library(patchwork) # The Composer of Plots
library(flextable) # Functions for Tabular Reporting
library(ftExtra) # Extensions for 'Flextable'
```

```{r PRISMA-flow-chart}
Pflow_chart <- grViz("digraph flowchart {
      node [fontname = Helvetica, shape = rectangle, style = filled, color = grey90]

      tab1 [label = '6832 references imported for screening']
      tab1_1 [label = '611 duplicates removed']
      tab2 [label = '6221 studies screened - title and abstract (step 1)']
      tab2_1 [label = '4511 studies excluded']
      tab3 [label = '1710 studies assessed for full-text eligibility (step 2 and 3)']
      tab3_1 [label = 
      <1545 studies excluded:<br ALIGN = 'LEFT'/> 
      &#8226; 1061 No ES (only EC)<br ALIGN = 'LEFT'/> 
      &#8226; 285 No full-text<br ALIGN = 'LEFT'/> 
      &#8226; 147 No EC and no ES<br ALIGN = 'LEFT'/> 
      &#8226; 38 No EC (only ES)<br ALIGN = 'LEFT'/> 
      &#8226; 13 Other<br ALIGN = 'LEFT'/> 
      &#8226; 1 Wrong study type<br ALIGN = 'LEFT'/> >]
      tab4 [label = 'Data from 165 studies extracted']

      blank1[label = '', width = 0.01, height = 0.01]
      tab1 -> blank1[dir=none];
      blank1 -> tab1_1[minlen=10];
      {{ rank = same; blank1 tab1_1 }}
      blank1 -> tab2

      blank2[label = '', width = 0.01, height = 0.01]
      tab2 -> blank2[dir=none];
      blank2 -> tab2_1[minlen=10];
      {{ rank = same; blank2 tab2_1 }}
      blank2 -> tab3

      blank3[label = '', width = 0.01, height = 0.01]
      tab3 -> blank3[dir=none];
      blank3 -> tab3_1[minlen=10];
      {{ rank = same; blank3 tab3_1 }}
      blank3 -> tab4
      }
      ")
```

```{r, eval=FALSE}
Pflow_chart |>
  DiagrammeRsvg::export_svg() |> 
  charToRaw() |> rsvg::rsvg_png(file="figures/Pflow_chart.tiff")
```

```{r import-extraction-data}
Consensus_all <- read_excel("ALL_Extraction_Results_Feb22_edited.xlsx") |> filter(`Reviewer Name` == "Consensus")

Consensus_only_analysis <- read_excel("ALL_Extraction_Results_Feb22_edited.xlsx") |> 
  filter(`Reviewer Name` == "Consensus") |> 
  filter(`Included in analysis` == "Yes")
```

```{r by-year-and-continent-plot}
year_cont_plot <- ggplot(data = Consensus_only_analysis,
                         aes(y = `_Publication year`, 
                         fill = `_Case study continent`)) +
  geom_bar(aes(x = (..count..)/sum(..count..)),col="black") + 
  scale_x_continuous(labels=scales::percent,
                     breaks = seq(0,0.2, by=0.05),
                     limits = c(0,0.2)) +
  xlab("% of papers") +
  theme_classic() + 
  theme(text=element_text(size=10),
        axis.title.y = element_blank(),
        legend.position = "bottom") + guides(fill = guide_legend(nrow = 2,
                                                                 title=NULL))
```

```{r modes-by-type-plot}
df_mode <- strsplit(Consensus_only_analysis$`_mode`, "; ", fixed = T)

df_mode <- data.frame(
  id=rep(Consensus_only_analysis$`Covidence #`, lengths(df_mode)),
  mode=unlist(df_mode))

df_mode <- df_mode |> group_by(mode) |> 
  summarize(count = n()) |> 
  mutate(count = count/165,
         sub_ob_j = "What mode?")

mode_plot <- ggplot(data = df_mode, aes(x = forcats::fct_reorder(mode,count), y= count))+ 
  geom_bar(stat="identity", fill = "coral2", col = "black") + 
  scale_y_continuous(labels=scales::percent,
                     breaks = seq(0,0.4, by=0.05),
                     limits = c(0,0.4)) +
  coord_flip() +
  labs(y = "% of mentions in papers", subtitle = "Modes") + 
  theme_classic() + 
  theme(text=element_text(size=10),
        axis.title.y = element_blank())
```

```{r population-groups-plot}
df_pop <- strsplit(Consensus_only_analysis$`_population`, "; ", 
                       fixed = T) 
df_pop <- data.frame(
  id=rep(Consensus_only_analysis$`Covidence #`, lengths(df_pop)),
  pop=unlist(df_pop)) |> 
  mutate(pop = ifelse(pop == "Vulnerable populations (composite)", "Comp. vul", pop))

df_pop <- df_pop |> group_by(pop) |> summarize(count = n()) |> 
  mutate(count = count/165)

#adding grouping var so they print on plot in the order we want
df_pop$group <- c(7,9,6,8,2,3,5,10,1,4)

pop_plot <- ggplot(data = df_pop, aes(x = pop, y = count)) +
  geom_bar(stat="identity", fill = c(rep("skyblue2",each=8), "grey", "skyblue2"), col = "black") + 
  scale_y_continuous(labels=scales::percent, limits = c(0,0.25), breaks = seq(0,0.25, by=0.05))+
  scale_x_discrete(limits=(df_pop$pop)[order(df_pop$group)]) +
  coord_flip() +
  labs(y = "% of mentions in papers", subtitle = "Population groups") + 
  theme_classic() + 
  theme(text=element_text(size=10),
        axis.title.y = element_blank())
```

```{r concs}
df_conc <- strsplit(Consensus_only_analysis$`_equity conceptualization`, "; ", 
                                                   fixed = T)

df_conc <- data.frame(
                              id=rep(Consensus_only_analysis$`Covidence #`, lengths(df_conc)),
                              conc=unlist(df_conc)) |> 
                              mutate(conc = ifelse(conc == "Inequitable exposure to externalities", "Ineq. externalities", conc),
                                     conc = ifelse(conc == "Sufficientarism / Capabilities approach", "Emerg. theories", conc),
                                     conc = ifelse(conc == "Transport-related exclusion", "TR exclusion", conc))

df_standard_mcat <- strsplit(Consensus_only_analysis$`_type of equity standard2`, "; ", 
                       fixed = T) 
df_standard_mcat <- data.frame(
  id=rep(Consensus_only_analysis$`Covidence #`, lengths(df_standard_mcat)),
  std_mcat=unlist(df_standard_mcat))

standards_conc <- merge(df_conc, df_standard_mcat, by="id", all = T)

NUM_standards_conc <- standards_conc |> group_by(id) |> summarize(count = n())
NUM_ENTRIES_standards_conc <- NUM_standards_conc$count |> sum()

standards_conc <- merge(standards_conc, NUM_standards_conc, by="id", all = T)

# standards_conc_plot <- ggplot(data = standards_conc,
#                               aes(x = `conc`, fill = `std_mcat`)) +
#   geom_bar() +
#   labs(x = "Conceptualizations (summarized)", fill = "Standards (main categories)") +
#   theme_classic() + 
#   theme(text=element_text(size=10),
#         axis.title.y = element_blank(),
#         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   coord_flip()

# c("Wellbeing","Vertical equity","Utilitarian","TR exclusion", "Spatial equity", "Rights", "Ineq. externalities", "Horizontal equity", "Emerg. theories")
```

```{r}
case_cont_count <- strsplit(Consensus_only_analysis$`_Case study continent`, "; ",
                       fixed = T)
case_cont_count <- data.frame(
  id=rep(Consensus_only_analysis$`_Case study continent`, lengths(case_cont_count)),
  case_cont=unlist(case_cont_count))
case_cont_count <- case_cont_count |> group_by(case_cont) |> summarize(count = n())

urban_form_count <- strsplit(Consensus_only_analysis$`_Case study urban form`, "; ",
                       fixed = T)
urban_form_count <- data.frame(
  id=rep(Consensus_only_analysis$`_Case study urban form`, lengths(urban_form_count)),
  urban_form=unlist(urban_form_count))
urban_form_count <- urban_form_count |> group_by(urban_form) |> summarize(count = n())

```

```{r calculating-destinations-of-interest}
dest_count <- strsplit(Consensus_only_analysis$`_activity`, "; ",
                       fixed = T)

dest_count <- data.frame(
  id=rep(Consensus_only_analysis$`_activity`, lengths(dest_count)),
  dest_count=unlist(dest_count))

dest_count <- dest_count |> group_by(dest_count) |> summarize(count = n()) 
```

```{r}
df_when <- Consensus_only_analysis |> 
  mutate(pub_year_group = case_when(`Publication year` >= 1992 & `Publication year` <= 2007 ~ "1992 to 2007",
                                    `Publication year` >= 2008 & `Publication year` <= 2012 ~ "2008 to 2012",
                                    `Publication year` >= 2013 & `Publication year` <= 2017 ~ "2013 to 2017",
                                    `Publication year` >= 2018 & `Publication year` <= 2022 ~ "2017 to Mar. 2021")) |>
  group_by(pub_year_group) |>
  summarise(count = n())|>
  mutate(sub_ob_j = "1.\nWhen?",
         count = count/165) |>
  rename("sub_ob_j_type" = "pub_year_group")

df_where <- data.frame(case_cont = c("Global North", "Global South"),
                                count = c(sum(case_cont_count[2:3,2],case_cont_count[4,2], 2), 
                                          sum(case_cont_count[4:6,2], 2))) |>
  mutate(sub_ob_j = "2.\nWhere?",
         count = count/165) |>
  rename("sub_ob_j_type" = "case_cont")

# THE WHO'S ARE MODE AND POPULATION GROUP
df_mode_2 <- df_mode |> mutate(sub_ob_j = "3.b Who?\n(mode)") |> 
  rename("sub_ob_j_type" = "mode")

df_pop_2 <- df_pop |> mutate(sub_ob_j = "3.a Who?\n(pop group)")  |>
  rename("sub_ob_j_type" = "pop") |> select(-c("group"))

# THE WHAT IS ARE DESTINATION group
df_dest <- dest_count |> mutate(sub_ob_j = "4.\nWhat?",
         count = count/165) |> 
  rename("sub_ob_j_type" = "dest_count")

df_all_sub_ob_just <- rbind(df_when, df_where, df_mode_2, df_pop_2, df_dest) #|>
  # mutate(sub_ob_j_type = as.factor(sub_ob_j_type),
  #        sub_ob_j = as.factor(sub_ob_j))
```
```{r}
# all_sub_ob_just_plot_stacked <- ggplot(data = df_all_sub_ob_just, 
#                                        aes(x = sub_ob_j, 
#                                            y = count,
#                                            fill = sub_ob_j_type,
#                                            fill = forcats::fct_reorder(sub_ob_j_type, count),
#                                            label = sub_ob_j_type
#                                            ))+ 
#   geom_bar(stat="identity", col = "black") + 
#   geom_text(size = 3, position = position_stack(vjust = 0.5)) +
#   scale_fill_manual(values = rep(RColorBrewer::brewer.pal(6,"Pastel2"),10))+
#   scale_y_continuous(labels=scales::percent,
#                      breaks = seq(0,1, by=0.1),
#                      limits = c(0,1)) +
#   coord_flip() +
#   labs(y = "% mentions in papers", x = element_blank(), subtitle = "Corpus Breakdown: % mention of topics in When, Where, Who and What") + 
#   theme_classic() +
#   theme(text=element_text(size=10),
#         axis.text.x  = element_text(size = 15),
#         legend.position="none")
# 
# all_sub_ob_just_plot_stacked
```

```{r}
df_all_sub_ob_just_reordered <- df_all_sub_ob_just |>
  mutate(
    sub_ob_j_type = ifelse(sub_ob_j_type=="Unspecified" & sub_ob_j == "3.b Who?\n(mode)", "Unspecified 3.b",sub_ob_j_type),
    sub_ob_j_type = ifelse(sub_ob_j_type=="Unspecified" & sub_ob_j == "4.\nWhat?", "Unspecified 4.",sub_ob_j_type),
    sub_ob_j_type = ifelse(sub_ob_j_type=="Employment" & sub_ob_j == "4.\nWhat?", "Employment 4.",sub_ob_j_type),
    sub_ob_j_type = ifelse(sub_ob_j_type=="Employment" & sub_ob_j == "3.a Who?\n(pop group)", "Employment 3.a",sub_ob_j_type),
    sub_ob_j_type = ifelse(sub_ob_j_type=="Other" & sub_ob_j == "3.b Who?\n(mode)", "Other 3.b",sub_ob_j_type),
        sub_ob_j_type = ifelse(sub_ob_j_type=="Other" & sub_ob_j == "3.a Who?\n(pop group)", "Other 3.a",sub_ob_j_type),
    sub_ob_j_type = fct_reorder(sub_ob_j_type, count, .fun='median'))

df_all_sub_ob_just_reordered$ID = seq.int(nrow(df_all_sub_ob_just))

#adding color column 
df_all_sub_ob_just_reordered <- df_all_sub_ob_just_reordered |>
  mutate(colour = c("#FCBBA1", "#FC9272", "#FB6A4A", "#EF3B2C",
                     "#D9D9D9", "#BDBDBD",
                    "#E5F5E0", "#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45",
                    "#F7FBFF", "#DEEBF7", "#C6DBEF", "#9ECAE1", "#6BAED6", "#4292C6", "#2171B5", "#08519C", "#08306B","#08519C",
                    "#FCFBFD", "#EFEDF5", "#DADAEB", "#BCBDDC", "#9E9AC8", "#807DBA", "#6A51A3", "#54278F"))
#c(RColorBrewer::brewer.pal(9,"Blues"))
```

```{r, echo=FALSE}
df_all_sub_ob_just_reordered_plot <- ggplot(data = df_all_sub_ob_just_reordered, 
       aes(y = reorder(sub_ob_j_type, count),
           x = count, 
           label = sub_ob_j_type,
           fill = sub_ob_j_type))+ 
  geom_bar(stat="identity",  col = "grey", alpha=0.8, size=0.05) +
  facet_grid(rows=vars(sub_ob_j),
             scales="free_y",
             space="free_y",
             switch = "y") +
   scale_x_continuous(labels=scales::percent,
                      breaks = seq(0,0.8, by=0.2),
                      limits = c(0,0.65),
                      expand = c(0,0))+
  scale_fill_manual(values= c("1992 to 2007" = "#FCBBA1",
                              "2008 to 2012" =  "#FC9272", 
                              "2013 to 2017" =  "#FB6A4A",
                              "2017 to Mar. 2021" =  "#EF3B2C",
                              "Global North" =  "#D9D9D9",
                              "Global South" =  "#BDBDBD",
                              "(Dis)Abilities" =  "#DEEBF7", 
                              "Age" =  "#6BAED6", 
                              "All populations" =  "#F7FBFF",
                              "Comp. vul" =  "#C6DBEF", 
                              "Education level" =  "#6BAED6", 
                              "Employment 3.a" =  "#9ECAE1", 
                              "Gender" =  "#DEEBF7", 
                              "Income" =  "#4292C6", 
                              "Other 3.a" =  "#9ECAE1",
                              "Race/Ethnicity" =  "#C6DBEF",
                              "Bicycle" =  "#A1D99B", 
                              "Car" =  "#74C476", 
                              "Other 3.b" =  "#E5F5E0", 
                              "Pedestrian" =  "#41AB5D",
                              "Transit" =  "#238B45",
                              "Unspecified 3.b" =  "#C7E9C0",
                              "Childcare" =  "#BCBDDC", 
                              "Community" =  "#DADAEB", 
                              "Education" =  "#EFEDF5", 
                              "Employment 4." =  "#9E9AC8", 
                              "Healthcare" =  "#807DBA", 
                              "Leisure" =  "#DADAEB", 
                              "Shops" =  "#BCBDDC", 
                              "Unspecified 4." =  "#6A51A3")) + 
                      #c(rep(RColorBrewer::brewer.pal(9,"Pastel2"),4))) +
  geom_text(size = 2.5, position = position_stack(vjust = 1), col = "Black") +
  theme_classic() +
  theme(strip.text.y = element_text(size = 5),
        strip.background.y = element_rect(fill = "lightgrey", linewidth=0.3),
        strip.placement = "inside",
        axis.title = element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.y=element_blank(),
        axis.text.x=element_text(size=7),
        axis.line = element_line(linewidth=0.3),
        legend.position = "none",
        panel.grid.major.x = element_line(colour="grey", linewidth=0.3))
```

```{r, eval=FALSE}
df_all_sub_ob_just_reordered_plot
ggsave("figures/df_all_sub_ob_just_reordered_plot.tiff", dpi=300) 
#manually remove the 3.a, 3.b. and 4. from the figure.
```

```{r, message=FALSE}
standards_conc <- standards_conc |> 
  group_by(conc, std_mcat) |> summarize(total = n())

standards_conc <- standards_conc |>
  mutate(total_perc_in_all = total/165)

empty_standards_conc <- data.frame(conc = c("Emerg. theories", "Utilitarian"),
                                   std_mcat = c("Environmental +", "Environmental +"), 
                                   total = c(NA,NA),
                                   total_perc_in_all = c(0,0))

totals_standards_conc <- rbind(standards_conc,empty_standards_conc)

#adding the grouping numbers, so the order that vars are printed are in geom_tile() are chronological
totals_standards_conc$group <- c(13,12,11,
                    44,43,42,41,
                    34,33,32,31,
                    24,23,22,21,
                    54,53,52,51,
                    74,73,72,71,
                    03,02,01,
                    84,83,82,81,
                    64,63,62,61,
                    14,
                    04)

#adding long definitions, for table under heatmap

totals_standards_conc_table <- data.frame(
  Conceptualizations = c("The distribution of transport system benefits/costs should favourably prioritize disadvantaged groups.",
                         "Concerning the links between transport system interaction and social exclusions.",
                         "Evaluation of transport system benefits/costs through wellbeing measures (e.g., livability, quality of life, subjective well being).",
                         "Equal distribution of transport system benefits/costs across space.",
                         "Equal distribution of transport system benefits/costs across population."),
  Conceptualizations_cont = c("Concerning the equitable distribution of externalities (e.g., road safety, congestion, pollution).",
                         "Concerning legal/human rights of access (e.g., mobilty rights, rights to the city, discriminatory application of laws).",
                         "E.g., Sufficientarism: minimums informed by individual capabilities. Capabilities approach: the capability of living the life one values (rather than having the right to do so)",
                         "Maximizing benefits/welfare overall, without consideration of distribution.",
                         " "),
  Standards = c("Related to user characteristics (e.g., affordability levels, perceived equity).",
                "Access to destinations indicators (e.g., # of parks within 30-min transit trip), the use of travel impedance thresholds (e.g., 300m to bus stop).",
                "E.g., Level of service, quality, universal design standards.",
                "E.g., noise, air, pollution criteria, safety.",
                " ")
)

```

```{r}
set_flextable_defaults(font.size= 6,
                       font.family = NULL,
                       text.align = 'justify')

totals_conc_table_flex <- 
  prepend_chunks(flextable(totals_standards_conc_table |> select(-c("Standards"))),
                 part = "body", i = c(1,2,3,4,5), j = 1, as_b(c("Vertical equity: ", "TR exclusion: ", "Wellbeing: ", "Spatial equity: ", "Horizontal equity: "))) |> 
  prepend_chunks(part = "body", i = c(1,2,3,4), j = 2, as_b(c("Ineq. externalities: ", "Rights: ", "Emerg. theories: ", "Utilitarian: "))) |> 
  # merge_h(part="header", i=1) |> #header merge
  merge_at(i=5, j=1:2) |>
  # align( align="center", part="header") |>
  delete_part("header") |>
  border_outer() |>
  border_inner_h() |>
  autofit()

totals_standards_table_flex <- 
  # compose(totals_standards_table_flex,
  #         i = 1, j = c(1), part = "header", value = as_paragraph(c("Standards"))) |> 
  prepend_chunks(flextable(totals_standards_conc_table |> select("Standards")),
                 part = "body", i = c(1,2,3,4), j = 1, as_b(c("Population: ", "Opportunity: ", "Infrastructure: ", "Environmental +: "))) |>
  merge_at(i= 4:5, j= 1) |>
  delete_part("header")|>
  border_outer() |>
  border_inner_h() |>
  autofit()
```

```{r}
standards_conc_percentages_plot <- ggplot(totals_standards_conc, 
       aes(x=std_mcat, y=conc, fill= total_perc_in_all ))+ 
  geom_tile() + 
  scale_fill_gradient2(low = "#AFAFED", mid= "#3F3FD4", high = "#191970", midpoint = 0.1, limits = c(0,.21), na.value = "grey40") +
  scale_x_discrete(limits=(totals_standards_conc$std_mcat)[order(totals_standards_conc$group)][c(4,3,2,1)]) + 
  scale_y_discrete(limits=(totals_standards_conc$conc)[order(totals_standards_conc$group)] ) +
  geom_text(aes(label = total_perc_in_all |> percent(accuracy = 1)), color = "white", size = 2.5) +
  theme_classic() +
  coord_flip()+
  theme(legend.position = "none",
        axis.ticks = element_blank(),
        axis.text.x=element_text(angle=20,hjust=1, colour = "black"),
        axis.text = element_text(size=7, colour = "black"),
        axis.title = element_blank()) 
```

```{r calculating-concepts-and-standards}
conc_count <- strsplit(Consensus_only_analysis$`_equity conceptualization`, "; ",
                       fixed = T)
conc_count <- data.frame(
  id=rep(Consensus_only_analysis$`_equity conceptualization`, lengths(conc_count)),
  conc_count=unlist(conc_count))
conc_count <- conc_count |> group_by(conc_count) |> summarize(count = n())

stand_count <- strsplit(Consensus_only_analysis$`_type of equity standard2`, "; ",
                       fixed = T)
stand_count <- data.frame(
  id=rep(Consensus_only_analysis$`_type of equity standard2`, lengths(stand_count)),
  stand_count=unlist(stand_count))
stand_count <- stand_count |> group_by(stand_count) |> summarize(count = n()) 
```

```{r, warning=FALSE}
#the heatmap on top and table of stand/concept definitions on the bottom.
standards_conc_percentages_plot_and_table <- cowplot::plot_grid(
  gen_grob(totals_standards_table_flex, scaling="fixed", fit="auto"),
  standards_conc_percentages_plot,
  NULL,
  gen_grob(totals_conc_table_flex, scaling = "fixed", fit="auto"),
  nrow = 2, ncol=2, rel_heights = c(0.8,0.7), rel_widths = c(0.8,4) )
```

```{r, eval=FALSE}
standards_conc_percentages_plot_and_table
ggsave("figures/standards_conc_percentages_plot_and_table.tiff", dpi=300)
```

```{r echo=FALSE}
#function for creating comments
word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Anastasia"',
           'date="2023-08-12T15:00:00Z"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

# Introduction {#sect1}

Transportation systems are technologies essential for social inclusion and activity participation, and therefore important from an equity perspective [@martensTransport2016; @karnerAdvancesPitfallsMeasuring2024; @vecchioTransportEquityLatin2020]. Beyond ethical motivations, tracking objective and perceived inequalities is of interest for governing bodies to respond to popular demands for fairness. This is a challenging task: transportation systems are notoriously complex, with benefits and burdens that are diffuse over space and time. To compound matters, emerging technologies and service models can swiftly change the balance of benefits and burdens among a population [@guoSystematicOverviewTransportation2020]. Transportation systems that are engineered to offer higher mobility for people _somewhere_ can simultaneously cut others off from essential opportunities _elsewhere_, as @rajeEngineering2004 and others have poignantly illustrated. Over time, the shades of policies past long haunt a region and the planet, as shown by the legacy of urban highways all across the US [@archerWhite2020] and the time horizon for the impacts of climate change [@markolfTransportation2019].

The objective of this work is to scan the state of knowledge in terms of defining and operationalizing fairness in transportation. Much research has been devoted to the issues of _measuring_ equity in transportation, including (among many others) [@delboscUsing2011; @martensMeasuring2019; @pritchardNew2022]. Further, there are multiple works that discuss the conceptual and philosophical foundations of equity and fairness in transportation [e.g., @martensTransport2016; @pereiraDistributiveJusticeEquity2017a; @vanoutriveHow2019]. Previous reviews of equity in planning documents have been tightly scoped to cover accessibility [e.g., @boisjolyHowGetThere2017] or a particular mode of transportation [e.g., cycling, @doranPursuitCyclingEquity2021]. These efforts are valuable, but in our estimation there remains a gap in terms of understanding how standards for equity are developed and implemented for transportation systems.

To illustrate this gap, we note how @beilerExploring2016 in their exploration of transport equity, cite the following strategies identified by the U.S. Department of Transportation to address matters of justice [p. 287]:

- Reduce adverse human health and environmental effects on minority and low-income populations.
- Include all potentially affected communities in the transportation decision-making process.
- Ensure that minority and low-income populations receive equitable benefits.

While commendable, the strategies are too vague, which means it is possible to implement them in multiple ways, to genuinely comply with the spirit of justice, or performatively to deceive it [@mcculloughPerformative2023]. Relevant questions include: how much should adverse effects be reduced? To zero? Or to some tolerable level of adversity greater than zero? What should that level be? What benefits should be distributed? What share of the benefits should minority and low-income residents receive for a project to be sufficiently ‘equitable'? These questions touch on the development and use of _standards_. In this work, we choose to focus exclusively on the academic domain as we know that adoption of equity concepts in planning practice has lagged developments in academic work [@pereiraTransportationEquity2021; @boisjolyHowGetThere2017; @linovskiEquityTransportationPlanning2020; @doranPursuitCyclingEquity2021].

Supporting the creation of (more) just transportation systems involves developing standards to track the production and management of benefits and costs; how they are distributed; what values they implement, as well as by whom [@pereiraDistributiveJusticeEquity2017a; @pereiraTransportationEquity2021]. In this spirit, our work seeks to make two contributions. First, it outlines a conceptual and flexible framework for engaging questions of transportation fairness and equity analysis, based on the questions "Why?", "Where?", "When?", "Who?", "What?", and "How?" (5WH). Second, it applies the framework to collate the existing knowledge about equity standards. In contrast to previous reviews on measuring inequality in transportation systems, we are concerned with the implicit or explicit standards used to judge whether inequalities are fundamentally “fair” or unacceptable. To achieve this we scan the state of academic knowledge in defining standards of fairness in transportation.

# Background {#sect2}

In the introduction the terms "justice", "equity", and "fairness" were used relatively loosely. In the following, we describe justice as a target, equity as a set of measurement tools, fairness as a yardstick of justice, and standards as statements of fairness.

**Justice** is an end goal, a desirable state of affairs that we are morally obligated to achieve. Justice is attained when people "give and receive whatever they are due" [@jaggarPhilosophicalChallengesGlobal2009, pp. 1-2], and it ceases to exist when there are persons or groups that are denied "access to the opportunities they need to lead a meaningful and dignified life" [@karnerTransportationEquityTransportation2020, p. 440]. Justice is fluid because it depends on the desirability of different states of affairs, which may change between populations, places, and times. That said, several forms of justice can be distinguished [@jaggarPhilosophicalChallengesGlobal2009; @pereiraDistributiveJusticeEquity2017a; @karnerTransportationEquityTransportation2020].

- **Retributive justice** is concerned with the proportional retribution of wrongdoers relative to legitimate punishers and the innocent [@sep-justice-retributive].

- **Reparative (or restorative) justice** focuses on the reparation of caused harm; it centers the needs and voices of victims to restore wrongdoers and the community according to moral values [@tylerRestorativeJusticeProcedural2006; @braithwaite2003principles]. In planning and policy contexts, reparative justice often involves accountability mechanisms to compensate victims [@safransky2022grammars].

- **Procedural justice** strives to ensure that the views and preferences of all stakeholders are fairly accounted for in the decision-making and inter-personal procedures affecting their lives and communities [@tylerRestorativeJusticeProcedural2006].

- **Distributive justice** is perhaps the most studied form of justice in transportation [@pereiraDistributiveJusticeEquity2017a] and elsewhere [see @jaggarPhilosophicalChallengesGlobal2009, p. 2]. Its main concern is the collection of benefits and burdens of the tangible and intangible products of society by different segments of a population.

It might be argued that all forms of justice have a distributive facet. To illustrate, retributive justice is achieved by distributing societal intangibles like "freedom" (e.g., of movement, of association) as benefits, and/or tangible resources as burdens (e.g., fines as a punishment). Reparative justice deals with the distribution of benefits and burdens to redress past wrongs, for example by asking those who have benefited from said wrongs, even if unwittingly, to shoulder a bigger fiscal burden to benefit those who are still harmed by them. Procedural justice is the distribution of societal benefits (e.g., the right to voice an opinion as a recognized stakeholder in a process) and burdens (e.g., the effort required to develop an educated opinion) when making decisions with collective consequences.

From this perspective, we can speak of the _purposes_ of distributive justice: to mete out retribution _fairly_, to repair past harms, and to ensure that procedures offer _equitable_ opportunities to influence outcomes. Accordingly, equity and fairness are the "instruments" of justice, the tools by which society moves towards it.

**Equity**, as conceptualized alongside distributive justice, tends to encompass tools to understand the distribution of benefits and burdens of things among a population, often emphasizing those with the least advantage. In the transportation domain, the term is somewhat loaded because it is perceived as stemming from the authority of the state, and meant to assist with decisions about regulating and financing spending [@karnerTransportationEquityTransportation2020]. We are in agreement with @karnerTransportationEquityTransportation2020 that equity analysis should not be seen as an end in and of itself, but rather as a means to gather information about actual and perceived inequities. In this respect, the analytical tradition of equity, at least in transportation planning, means that the relevant models become embedded in a political ecology of estimated truth, and their assumptions and scope must be open and transparent to avoid misuse and even abuse as tools of subjugation.

**Fairness** is somewhat more complicated to define. The concept does not have the same history of development as an analytical tool, and can be interpreted in numerous, and possibly discordant ways. This is demonstrated by @martensFair2021 in their study of the application of Title VI of the US Civil Rights Act of 1964 in accessibility planning. Title VI explicitly talks about the distribution of benefits derived from Federal funding: "[n]o person in the United States shall, on the ground of race, color, or national origin, be excluded from participation in, be denied the benefits of, or be subjected to discrimination under any program or activity receiving Federal financial assistance." However, as @martensFair2021 show, there are wildly different paths to compliance, ranging from the banal (do not _knowingly_ discriminate), to the substantive (compensation for past discrimination within a societal context that recognizes harm was done, i.e., reparative justice). What kind of justice does fairness serve in each case? It depends on the rationale for seeking justice in the first place. Our reading of @martensFair2021 is that fairness is a yardstick that is best deployed _a priori_ than _a posteriori_, for doing the latter risks rationalizing the outcomes rather than driving them.

A **standard** is a way of making concrete statements about fairness. Returning to the ambiguities in Title VI, the attainment of justice _depends on the standard_. For example, explicit non-discrimination constitutes a very weak standard that takes aim at the actions of agencies instead of the recipients of the benefits. Accordingly, any distribution of benefits would be considered fair, as long as the agency does not explicitly and knowingly target or deny benefits to particular groups. The standard provides conditions to determine whether a situation is _fair_. A similarly weak standard is a Pareto improvement, whereby it is possible to concentrate the benefits as long as no group is worse off compared to the status quo [@tan2016pareto; @xuParetoimprovingPoliciesIdealized2018]. A policy based on this standard would be fair as long it does no harm. <!--A somewhat more strict standard, a Pareto-Plus improvement, stipulates that an intervention is fair when all groups receive at least _some_ (non-trivial) benefits; the size of the benefits for each group is irrelevant. In contrast to the notion of "do no harm", such a standard embodies the ideal that no one is denied benefits. --> A standard based on egalitarian principles (e.g., proportional equity) would weigh the benefits or burdens by population; in fairness, each group would give or receive in proportion to their size [@billsLookingMeanEquity2017;@martensJusticetheoreticApproachDistribution2012]. In contrast, an affirmative action standard (restorative justice) would be even stricter, requiring the benefits to be distributed in a non-egalitarian way to favor people still harmed by past or present discriminatory practices [@bierbaumMobilityJusticeLinking2021].

## A framework to analyze questions of justice: 5WH {#sect2.2}

Having defined "justice", "fairness", "equity" and "standards", we approach the literature with an analytical apparatus inspired by the framing of @jaggarPhilosophicalChallengesGlobal2009 for philosophical questions of justice.^[Similar questions are found peppered throughout the literature. This is done either explicitly, as for example in @karnerTransportationEquityTransportation2020, who ask "of what", "for whom," and "how much" in reference to equity; or implicitly, as in @gosslingUrban2016, who asks of the outputs ("what?") of transportation (exposure, space, access) and "for whom?" (gender, age, ethnicity).]. According to @jaggarPhilosophicalChallengesGlobal2009, Western philosophy has approached the issue of justice by asking "Why?", "Where?", "When?", "Who?", "What?", and "How?" (5WH), applying them to a particular domain or sphere of life relevant to justice.

In the case of transportation the question of **"Where?"** is paramount, as transportation by its very nature concentrates the benefits (e.g., access points to the system are not ubiquitous). The burdens, in contrast, are often diffuse: they are incrementally paid, for example by a distributed population in the form of taxes, or by a (possibly different) population in the form of poor health. As such, the answer to "Where?" is the definition of the spatial boundaries.

Conventionally, the question of **"When?"** refers to the temporal circumstances within which the demands of justice apply. When it comes to transportation, questions regarding temporailty are important, as for example: _when_ did the equity analysis take place and under what historical policy context; when is _the right time_ to invest in infrastructure and as a result when to generate a spatial inequality [@rabelloEvaluation2015]; for _how long_ the burdens and benefits can be associated to a specific intervention; or even the _timeline_ of reparative justice to reconcile the shadows of past transportation-related injustices.

When asking **"Who?"** we think about the entities that should be regarded as subjects/arbiters of justice. For tractability, this question is often approached through the filter of population groups, which may include several concurrent traits, including gender identity, ableness, ethnicity, age, caste, and income. Often, it is appropriate to consider the intersections between traits, given differences in a person's lived experiences. A complication in the case of transportation is that disentangling the "Who" from their mobility tools is not always straightforward. Although a person is not their mode of transportation, there are large segments of the population who cannot extricate themselves from the mobility tools they use, either because they have driven themselves out of choices [@laveryDriving2013], or have been driven out of choices by factors beyond their control [@jacquesRescuing2012]. While it is important to avoid conflating the "Who" with the "What", for analytical purposes we need to be mindful of the connection between a person and their mobility tools.

**"What?"** refers to which entities should be regarded as objects of inequities, meaning which categories of things should be distributed in a just manner. To understand the distributional implications of transportation systems, it is essential to understand what they _produce_. Transportation systems are technologies to improve the rate at which space is traded for time by increasing the speed of movement. But as the adage goes, travel is derived demand [@mokhtarianUnderstanding2001]. For this reason, we cannot stop at considering only mobility but must consider its ulterior goal--reaching destinations. Coupled with land use, mobility creates accessibility to opportunities as well as varied burdens. Transportation justice, thus, involves proximate (mobility tools and mobility) and ulterior (accessibility and activity opportunities) objectives.

The next question is **"How?"**, and it relates to the allocation of various objects of justice ("what") to various subjects of justice ("who") in various circumstances ("when" and "where"). Equity standards are a tool for answering this distributive question, by defining what is fair. The thresholds can be quantitative (e.g., square meters of green space per capita), or they can be qualitative descriptions (e.g., do not knowingly discriminate), or a mix of the two. Some examples include: maximum travel distance/cost/time to or from key destinations, levels of maximum exposure to externalities (i.e., noise or air pollution), un/fulfilled needs, and dis/satisfaction with travel. To support us when approaching this question we draw from concepts in transport-related social exclusion, transport disadvantage, and/or transport poverty, which are typically based on utilitarian or sufficientarian philosophies.

Above all, convincing answers to the above questions require a supporting rationale: a **"Why?"** [@jaggarPhilosophicalChallengesGlobal2009]. This is perhaps the most slippery of the questions posed here. Justice is inherently a social construct. Asking **Why?** amounts to asking what sort of social contracts regulate human interactions, that is, the self-imposed rules that result from our collective will to believe. These contracts can be defined by constitution, but there are often unwritten and possibly contested variants. In this way, analyzing the "Why?" in the reviewed literature is not the focus of this review, partly because answers to "Why?" are seldom explicitly stated. Instead, our focus is on the standards of fairness that, combined with the use of equity analysis, can help us understand how better to move towards just transportation systems and better formulate answers to "Why?".

## Methods {#sect2.3}

This review examines the academic literature on transportation to identify the extent to which standards for equity are defined and employed. In this task, we follow the Joanna Briggs Institute (JBI) approach to the conduct of scoping reviews, an approach that builds upon the Arksey and O'Malley (2005) framework [@petersUpdatedMethodologicalGuidance2020]. The review is also guided by the Preferred Reporting Items for Systematic Reviews and Meta-Analyses, particularly the extension for scoping reviews (PRISMA-ScR) which is consistent with the JBI approach [@triccoPRISMAExtensionScoping2018].

The primary research question and the protocol were initially defined by the authors, a group of experts in the field of transportation. The initial draft of the protocol was refined from preliminary searches of related reviews [e.g., @zhangLiterature2021; @pereiraDistributiveJusticeEquity2017a; @vecchioAccessibilityCapabilitiesApproach2021], and in consultation with research services librarians.

The search strategy was developed iteratively using **inclusion** and **exclusion** criteria [@petersUpdatedMethodologicalGuidance2020]. For the inclusion criteria, the mnemonic PCC (population, concept, and context) was adopted. The search strategy was refined by adding search terms. The terms were bundled by means of Boolean operators. These stages are summarized as follows (see Appendix section @sec-sect61 for details):

1. An initial limited search of Web of Science (WoS) Core Collection (containing journals, conference proceedings, and books published all over the world) was undertaken to identify key documents. Separate searches using the terms 'transportation' and 'equity' were generated. From these searches, we examined the text contained in the titles and abstracts, the index terms, and subject heading searches when available. As we developed a clearer outline of the literature, we refined the terms used for the search. This took the form: ("Transport" OR "Transit" OR "Car\*" OR "Walk" OR "Bike"...**1**) AND ("Equity" OR "Justice" OR "Fair"...**2**), where **1** and **2** signify additional terms relating to 'transportation' and 'equity', respectively.

2. Upon inspection of the preliminary search results and after achieving a consensus among the authors, the set of search terms related to 'equity' was expanded into three sets of terms. The first describes theories and concepts of equity, the second describes the object of justice (i.e., the "what" in our analytical framework), and the third describes terms referring to standards (i.e., the "how"). These sets of terms were augmented iteratively. The final search query took the following general form: ("Transport" OR "Transit" OR "Car\*" OR "Walk" OR "Bike"...**1**) AND ("Equity" OR "Justice" OR "Equity" OR "Fair"...**2**) AND ("Accessibility" OR "Mobility" OR ...**3**) AND ("Standard" OR "Threshold" OR ...**4**) where **1**,**2**,**3**, and **4** signify additional terms included in the sets combined with "OR" operators.

After testing the search strategy on WoS Core Collection, we applied it to an augmented list of databases. The databases used were: WoS General Collection-Science Citation Index Expanded, WoS Social Sciences Citation Index, and Transportation Research International Documentation (TRID). The definitive version of the search was completed and exported by the lead author on March 21st, 2021. The number of documents identified in this was was 6,382.

The semi-automated nature of the search strategy was overly inclusive, and reduced the risk of omitting relevant material. The next stage was to trim the corpus of documents, a task that could no longer be automated. The authors working with a group of undergraduate research assistants scanned the documents to assess their relevance based on titles and abstracts. Two research assistants voted on each document, and a third vote from the authorship team broke ties. After this step, 1,710 documents were assessed based on full-text, again with each being voted on by two research assistants and an authorship team member tie-breaker. Next, using a data extraction template and workflow that was pilot-tested with a subset of papers, the authorship team extracted data from the eligible documents using *Covidence* [@covidenceCovidence2023], an online application for literature screening. The evidence selection workflow, (PRISMA) flow diagram [@pagePRISMA2020Statement2021], data extraction template, and some sample data extractions can be consulted in Appendix @fig-figA2. The end result of this scan was a corpus of 165 documents retained for data extraction.

# An appraisal of the lay of the land {#sect3}

This section threads together the trends identified from the data extracted from the corpus (@fig-figA2)<!-- update this cross reference!-->. Specifically, the "How" and its application to the objects of justice (the "What"), the subjects of justice (the "Who"), and under which situations ("When" and "Where"). As an overview, @fig-fig1 shows the prominence of each category under "When?", "Where?", "Who?", and "What?". The "How?" category is shown in @fig-fig2.

```{r plotting-who-what-where-when-plot-fig1}
#| label: fig-fig1
#| fig-cap: "The proportion of category for each When, Where, Who and What; papers do often contain more than one topic and those papers are represented multiple times. Topic categories were generated upon data extraction."

knitr::include_graphics("figures/df_all_sub_ob_just_reordered_plot.tiff")
```

```{r plot-standards-concs-perc, warning=FALSE}
#| label: fig-fig2
#| fig-cap: "The prominence of Hows: the proportion of equity standards (vertical axis) within each type of equity conceptualisation (horizontal axis) category."

knitr::include_graphics("figures/standards_conc_percentages_plot_and_table.tiff")
```

## "When" and "Where" is transportation considered a sphere of justice {#sect3.1}

Within our corpus most papers (`r ((case_cont_count$count[4] + case_cont_count$count[5]+ case_cont_count$count[6])/165) |> percent()`) focus on studies in the Global North, with many studies from North America (particularly USA and Canada), Europe (particularly France, Spain and Northern Europe), Oceania (Australia and New Zealand), and Asia (Japan and Israel). Though their subject matter is varied, their spatial context mainly pertains to North America and Europe, and thus more often than not deal with more developed and formal government transport planning apparatuses and technologies e.g., planning for equitable high-speed rail [@monzonEfficiencySpatialEquity2013], autonomous vehicle technology [@eppenbergerOpportunitySharedAutonomous2021], or on the public consultation processes [@reddySafeguardingMinorityCivil2010].

The smattering of studies from the Global South are predominately from Asia, notably China, but also India, Thailand, Iran, Philippines, and Indonesia. The next most common focus within the literature from the Global South is from South America. Many of these studies mention a systemic absence of evidence relevant to the region [@vecchioTransportEquityLatin2020]. Despite a growing recognition of the interconnections between transport development, social exclusion, and poverty [@benevenutoExaminingTransportNeeds2020], a number of studies underscore ongoing neglect of the social dimension of transport during the planning stage [@benevenutoExaminingTransportNeeds2020; @boisjolyAccessibilityMeasurementsSao2020]. Many studies also point at affordability as one of the main mobility barriers in the region [@falavignaAssessingInequalitiesPublic2016; @rivasHowAffordableTransportation2018], while some highlight multi-dimensional concerns such as public transport accessibility and quality of walking environments that contribute to mobility inequalities [@tiznadoaitkenRoleAccessibilityPublic2018]. Studies pertaining to Africa are even less numerous.

A shared characteristic among studies from Africa and South America is a scarcity of official transport data [@friedMeasuringSustainableDevelopment2020]. These studies also incorporate the use of informal transportation options and tensions in developing road network infrastructure (which tends to support car dependency) over meeting the mobility/accessibility needs of citizens more equitably and sustainably [@thondooSmallCitiesBig2020]. To address these challenges, researchers compile databases based on open and geo-referenced data, calculate objective and/or subjective measures [@berheAdaptationDissonanceQuality2014], and focus on advancing transport justice for low to medium income countries by aligning their goals with external policy guidelines such as the Sustainability Development Goals (SDG), particularly those related to universal accessibility [@friedMeasuringSustainableDevelopment2020].

Of all the studies reviewed, `r (urban_form_count$count[3]/sum(urban_form_count$count)) |> percent()` focus on urban and suburban settings and are highly varied in their research aims. To give an example, @coxQualitativeStudyAccessibility2020 qualitatively examine cycling as a mode of travel for people with disabilities in a typical mid-size town in the UK. @ampeImpactChildBike2020, on the other hand, work to identify the lateral clearance that motorists should maintain when passing cyclists with children seats. The remainder of the studies focus on rural regions (`r (urban_form_count$count[2]/sum(urban_form_count$count)) |> percent()`), or remote regions, such as those that rely on inter-island ferry trips in the Philippines [@caoIndicatorsSocioSpatialTransport2017]. Similarly, @parrySocialVulnerabilityClimatic2018 studied remote communities in the Amazonian region, and suggest that "increasing accessibility through road building would be maladaptive, exposing marginalized people to further harm and exacerbating climatic change by driving deforestation" (pp. 125).

Overall, studies from the Global South often have some key differences in focus compared to the Global North:

- affordability as a barrier at the user- or policy maker- level is more often the motivation in work from the Global South;

- the expression of greater tensions in investing in new transportation infrastructure, such as roads in rural/under-developed areas, compared to prioritization of non-car modes. Studies centered in the Global North often focus on the later; and

- more significant data availability limitations and reliance on 'informally' collected data.

These differences center on the direct economic outcomes tied to transport infrastructure. Often, work from the Global South does not engage as intimately with emerging mobility technologies. As well, the informal aspects of transport planning are more overbearing. Clearly, countries in the Global South still struggle with the consequences of past colonialism, which has left them more reliant on primary sector exports (lower efficiency, lower national GDP) under growing global financial markets, and with more fragile democracies. Due to lower data availability, reliance on crowd-sourced or ‘informally’ collected data, and more extreme needs for ‘sufficient’ transport, analysis of transportation inequities is often cast along economic lines in the Global South.

The literatures on the Global North and South literature seem to be situated at different stages in the process of addressing transport inequities. As formal planning processes in the South operate under greater financial precariety, they rely on more informal processes to address unmet needs. This turns out to be a significant source of equity concerns e.g, informal transit [@friedMeasuringSustainableDevelopment2020], populations living in informal settlements [@sharmaPublicTransitAccessibility2021]. Lack of resources makes international standards highly relevant e.g., using the WHO noise and air pollution standard [@apparicioCyclingOneMost2021]. In contrast, these minimum standards may not be relevant in some areas of the Global North e.g., $NO_2$ pollution near roadways in Montreal are consistently below the WHO standard though populations are disproportionately impacted [@carrierApplicationThreeMethods2014]. Overall, the Global South lags behind the Global North, but this situation presents surprising opportunities for the Global South, to adopt potentially successful enhancements from Global North formal equity planning processes (e.g., indicator creation for disadvantages groups [@cuiSpatialAccessPublic2020]) and to avoid past mistakes (e.g., from entrenched car-centric development [@warrenDevelopingEquitableSustainable2015] to the disproportionate contribution to carbon-intense mobility [@chancelCarbonInequalityKyoto2015]. For the Global North, there is an opportunity to reckon with its own contributions to uneven development globally and environmental impacts as well as adopt relevant aspects from informal planning processes.

## "Who" are the subjects of transportation justice and their mobility tools {#sect3.2}

### Population groups

Turning to the question of "who?" in @fig-fig1, the focus of the literature tends to be on "Income", especially the lowest-income groups who simultaneously face lower mobility and accessibility _and_ higher costs and exposure [@peungnumsaiGridBasedSpatialAnalysis2020; @zhaoUnequableSpatialAccessibility2020; @falavignaAssessingInequalitiesPublic2016]. Evidence suggests that low household income is a significant determinant of transport-related inequities e.g., urban access to public transport [@peungnumsaiGridBasedSpatialAnalysis2020], access to urban employment opportunities [@boisjolyAccessibilityMeasurementsSao2020], and unfavorable rates of environmental noise, air pollution, and green space [@kruizeEnvironmentalEquityRole2007]. Yet, low-income is not universally associated with lower transport-related benefits for every object of justice. For instance, in Sheffield, UK, @mearsUnderstandingSocioeconomicEquity2019 demonstrate that historically working-class (lower-income) neighbourhoods have _more_ access to green space than other neighbourhoods, but of lower quality, likely due to historic urban planning approaches. Similarly, @bertrandMeasuringMappingDisparities2008 find that the granularity of the analysis matters, and lower income groups do not always have lower accessibility , something echoed in other spatial and temporal contexts [@fothEquitableTransitExamining2013a; @allenSizingTransportPoverty2019a]

"Age" is the second most common personal trait. In this regards, @martinez-jimenezAccessibilityCultureEducation2019 and @arranz-lopezSocialSpatialEquity2019 investigate travel distances/times to various opportunities, demonstrating how age is associated with differences in opportunity access. School-aged children and older populations are a common focus. For children, analysis of wellbeing [@laszkiewiczChildrenGreenWalk2020], safety and access to schools [@sharmaSpatialSocialInequities2022; @corazzaMethodologyEvidenceCase2020], and promoting active travel [@mackieOvercomingBarriersCycling2009; @mehdizadehWalkingTimeSchool2017] are common-place. Papers that focus on older adults similarly investigate transport-related wellbeing [@chenSpatialAnalysisFramework2020], access to age-specific destinations [@chengInvestigatingWalkingAccessibility2019], and options to reduce unmet travel needs [@nordbakkeTransportUnmetActivity2015].

Many papers focus on intersecting characteristics. As an exception, we classify some studies as focusing on  "(Dis)abilities" or "All Populations". Studies with a "(Dis)abilities" focus assess travel capabilities, namely through physical accessibility and universal design guidelines [@parkJourneyVisuallyImpaired2017; @chiscanoImprovingDesignUrban2021; @orellanaWalkRollMapping2020]. "All Populations" papers make no distinction in population. This is done, for instance in @kitaDoesLevelTransportation2020, who investigates disparities in accessibility to food-stores and self-reported capability/frequency of going outdoors. Often, the implicit or explicit motivation of these papers is access (to necessities and desired destinations) for all.

A large proportion of papers apply composite vulnerability indices that combine several individual traits like low income, unemployment, and/or immigrant status. These indices are generated from government sources or author-informed census data creation methods. As an example, the Neighbourhood Equity Index (NEI) is a measure of vulnerability created by the City of Toronto and used in @awuorInfluenceEnvironmentalHealth2019 to examine disparities in premature death. Other works use national census indicators such as the social and housing deprivation index [@pucciInequalitiesJobrelatedAccessibility2019] or explore transport disadvantage, equity in policy implementation, or transport-related mortality burden by means of census measures (e.g., household poverty) and transport-related accessibility indicators [@aldredEquityNewActive2021; @iungmanImpactUrbanTransport2021; @sunPublicTransportAvailability2021; @scheurerSpatialAccessibilityPublic2017]. Similarly, Environmental Justice (EJ) indicators have been used in the US literature to identify neighbourhoods that have a higher than average proportion of low-income and non-white populations and evaluate the equity impacts of transportation projects [@rowangouldIdentifyingEnvironmentalJustice2016; @parkTransitParksEnvironmental2021; @reddySafeguardingMinorityCivil2010].

Multi-dimensional considerations are so prevalent in the corpus that "Gender", "Race/ethnicity", "Education", or "Employment" are infrequently studied in isolation. Only a few papers focus exclusively on gendered differences in active transportation [@adlakhaMindGapGender2020, @xieWonCycleRoute2018], race/ethnicity's relationship to green space proximity [@silvaEnvironmentalJusticeAccessibility2018], and culturally-appropriate opportunities [@wangModelingSpatialAccessibility2011]. Papers that focus *solely* on "Employment" status or "Education" level are completely absent. Furthermore, "Other" population groups are also frequently considered in the corpus: this is a catch-all category that includes populations less commonly targeted by research e.g., veterans and access to specific-healthcare needs [@mooneyTravelDistanceBarrier2000], pregnant people and access to maternity services [@vadrevuMeasuringSpatialEquity2016], and youth who live in foster care [@batscheUsingGISEnhance2012]. Overall, the diversity of population groups considered in the corpus demonstrates the variety of transportation-equity concerns addressed in the literature.

### Modes (mobility tools)

Travel mode, though modifiable, is intertwined with individual identity, posing challenges for equity analysis. Here we choose to view the mode of travel primarily as those who use the mode, hence as subjects of justice (i.e., the “who”), but occasionally, this will be enjoined to an object of justice (i.e., a “what” such as inequities in rural transit [@giuffridaAddressingPublicTransport2021]).

Transit characterises the majority of the corpus (@fig-fig1). Despite transit being perceived unfavourably by some [e.g., @mellaDrivers2021], it is often viewed as the only and primary mobility option mode for many [@jacquesRescuing2012]. Papers that focus on transit assess a variety of topics, including food deserts [@mckeyCrowdsourcedMappingHealthy2020] and barrier-free transportation for people who face disabilities [@jimenez-espadaResearchProblemUniversal2021; @liuStatusQuoChallenges2019; @limFacilitatingIndependentCommuting2021; @feeleyValidationParatransitSkills2019]. We infer that transit is reckoned as a public good and hence a natural object of justice. Accordingly, public transportation can be modified to meet the demands of justice, by funding it sufficiently to provide barrier-free transport for most, despite challenges like low densities, fiscal constraints and political will [@markardUnsustainabilities2023].

Transit also plays a central role in multimodal comparisons for transport equity analysis, and often feature in papers dealing with "Walk", "Car", or some other category in @fig-fig1. For instance, @brusselAccessAccessibilityCritique2019 compares transit, pedestrian and road network accessibility measurements in the context of SDG 11.2. @renneWhatHasAmerica2018 reviews natural disaster evacuation plans, focusing on car-less households and emphasizing transit and pedestrian networks. A few papers also frame transit as a direct competitor of car travel or use it as a benchmark [@golubUsingPrinciplesJustice2014; @martensJusticetheoreticApproachDistribution2012]. For instance, @warrenDevelopingEquitableSustainable2015 propose car ownership standards while acknowledging the tension between mobility needs in transit under-served areas and emission reduction goals. However, this framing is not universal, and transit is sometimes seen as a mode to fulfill individual capabilities. As an example, @smithAccessibilityCapabilityMinimum2012 explores perspectives about transport needs and costs to achieve perceived sufficient living standards for those living in rural areas. Notably, papers vary in the importance they place on climate urgency, with some focusing more on satisfying *all* sufficient individual needs while planning for less car-dependent cities in the future.

After transit, pedestrian travel ("Walk") is the second most studied object of justice. Pedestrians represent a unique convergence of "what" and "who," utilizing their own bodies for mobility. Papers focusing solely on walking often use walkability scores to assess neighbourhood quality [@evansAccessibilityUserNeeds2015], mobility by different demographics [@towneUsingWalkScore2016; @kimNeighborhoodEnvironmentWalkability2016], or urban peripheral regions [@blecicCapabilitywiseWalkabilityEvaluation2021]. Some use walkability to gauge public health and urban vitality [@sungResidentialBuiltEnvironment2015; @mccormackRelationshipClusteranalysisDerived2012]. Papers with a pedestrian focus also often see walking as a bridge to connect multiple modes: they discuss 'walkability' as part of active transportation, which focuses on both walking, bicycle and/or transit. Concepts discussed include how active transport contributes to children's physical activity levels [@mammenSchoolTravelPlanning2014], walkability as an alternative to car predominance [@bertrandMeasuringMappingDisparities2008] or tension that exists between modes, creating unsafe conditions for walking [@siuAssessmentPhysicalEnvironment2019; @ferenchakEquityAnalysisProactively2019].

The third most studied mode is "Car". Unlike transit and walking, cars are rarely the sole focus of equity research. In our corpus they are often critically compared with transit or considered in areas with inadequate transit [@kimmelStructuralBarriersComprehensive2018; @aljoufie2016]. Upon the rise of automobility, when walking and cycling got in the way of faster travel, transit came to be seen as a social service and sometimes as a hindrance to the full realization of automobility. Our assessement of the focus on non-car modes in our corpus is that it represents a reckoning with automobility's legacy and ongoing demands for space, public subsidies, and government supports that underestimate the true cost of cars [@gosslingSocial2019; @timperleyFight2021]. In addition to comparing "Car" to other modes, some papers reflect on "Other" forms of mobility (@fig-fig1), including wheelchair accessible taxi and/or paratransit services [@marquezFactorsAffectingPersonal2019; @wilkinsonmeyersLiveOrdinaryLife2015], travel on waterways [@vadrevuMeasuringSpatialEquity2016; @caoIndicatorsSocioSpatialTransport2017; @parrySocialVulnerabilityClimatic2018], motorcycle or other micro-mobility [@berryInvestigatingFuelPoverty2016; @savingmothersDidSavingMothers2019; @tiwariDevelopmentPublicTransport2021], or emergency vehicles [@pedigoInvestigationDisparitiesGeographic2010; @patelDeterminingGeographicAreas2007].

Papers that pay no particular attention to any mode are "Unspecified": as examples, a focus on road infrastructure or road network distances [@wismadiSpatialPreferenceModelling2014; @mishraEIABasedComparative2014], travel needs generally [@titheridgeSocialExclusionAccessibility2008; @benevenutoExaminingTransportNeeds2020], realized travel [@abasoloEquityUtilizationAccess2001a], or externalities of realized travel [@iungmanImpactUrbanTransport2021].

### Who is missing in the "who"?

As shown, "who" is the subject of inequities provides ample but not exhaustive coverage of populations. But who is missing from the "Who" question? There is a need to go beyond the focus on low-income transit riders; positioning intersectionality of socio-economic and mode-use to create community-based definitions of equity and tailored standards are critical. How decision makers define an equity-deserving community will impact results e.g., [@rowangouldIdentifyingEnvironmentalJustice2015]. In this sense, standards need to be sensitive to changing community-based definitions of inequities. Are issues of economic-inequity at the root of transport inequities for a specific community? Are (dis)abilities? Are inequities in the service of transit the focus of a study because it can be improved to address transport inequities and are alternative modes driving these inequities? Access to what sorts of opportunities are driving transport inequities? How do populations, transit modes, and opportunities sought intersect to define the "who" of inequities? A community- informed understanding of inequities and tracking how it changes are needed.

## "What" are the objects of transportation justice {#sect3.3}

The most fundamental benefit of transportation systems is mobility (enabling or impeding movement). This benefit is sometimes valued by itself but is often seen by the literature as instrumental to achieve an ulterior goal (e.g., activity participation and associated benefits). For example, although vehicle kilometers traveled (VKT) is sometimes seen as a useful policy instrument [@zhaoRethinking2021], travelling more is not necessarily a sign of advantage when accessibility is low, and short trips may actually be a sign of advantage [@parkTransitParksEnvironmental2021]. For this reason, although the right to the road (and transportation systems more generally) is important, the literature leans heavily on the ulterior object, namely accessibility to (the ease of reaching) destinations.

### Mobility

Most papers in our corpus take a broad approach, with `r (dest_count$count[8]/165) |> percent()` focusing on "Unspecified" destinations (@fig-fig1). They examine various equity dimensions across different transportation modes. Some papers focus on factors that impact the trip itself like infrastructure quality, safety, and service level [@zheEvaluationSharedUse2008; @prasertsubpakijEvaluatingAccessibilityBangkok2012; @furstBestBadPractices2013 ; @lattmanPerceivedAccessibilityPublic2016]. Others analyze trips tailored for specific groups like women or people with disabilities [@russellPedallingEquityExploring2021; @wilkinsonmeyersLiveOrdinaryLife2015], often with a consideration of what constitutes 'sufficient' quality of life [@churchillTransportPovertySubjective2019]. In sum, these papers reveal the multifaceted nature of transportation systems: they serve utilitarian purposes while also shaping user experiences.

### Accessibility

In terms of destinations, "Healthcare" services (`r (dest_count$count[6]/165) |> percent()`) and "Employment" (`r (dest_count$count[5]/165) |> percent()`) are the most common in our corpus (@fig-fig1). Papers on healthcare often highlight disparities in services, like @wangModelingSpatialAccessibility2011, who assess access to Mandarin-speaking family physicians for Chinese speakers in Toronto. Similarly, papers focusing on employment are often aimed at identifying transportation-poor neighbourhoods [@allenSizingTransportPoverty2019a; @churchillTransportPovertySubjective2019]. Employment is frequently used as a proxy for overall accessibility since it is the most common trip purpose and employment is usually co-located with destinations like shops, recreation, and other services. These studies typically use travel surveys, census data, and point-of-interest databases, and benefit from well-developed institutional data. This especially holds in the Global North where these data are more readily available.

Other destinations have received less attention despite serving essential needs. "Shopping" destinations (`r (dest_count$count[7]/165) |> percent()`) often aim to identify food deserts [@choiFoodDesertsActivity2013; @jiaoHowIdentifyFood2012; @mckeyCrowdsourcedMappingHealthy2020; @kimAssessingSocialSpatial2020]. "Education"-related papers (`r (dest_count$count[3]/165) |> percent()`) explore children's active transportation to school [@larkinsAccessibleTransportationBuilt2011] or universal design [@larkinsAccessibleTransportationBuilt2011]. Places of "Leisure" (`r (dest_count$count[6]/165) |> percent()`) prompt questions about their spatial distribution [@xuSocialInequalitiesPark2017] and accessibility [@mavoaAreaLevelDisparitiesPublic2015]. Fewer papers cover "Community" destinations (e.g., public service centres, places of community support or worship) (`r (dest_count$count[2]/165) |> percent()`) or "Childcare" (`r (dest_count$count[1]/165) |> percent()`), but they are integral to holistically study activity participation [@albertsRebuildingWomenLivelihoods2016; @smithAccessibilityCapabilityMinimum2012]. The lack of information about community destinations, especially for children, is noticeable [@desjardinsChildren2022].

## "How" is fairness determined {#sect3.4}

### Conceptualisations of fairness

The conceptual foundations of fairness are not always made explicit by researchers and often had to be inferred from the corpus. Some categories emerged from our review and we must note that they are not exhaustive. In broad terms, they are (@fig-fig2 details definitions):

- Vertical equity (`r df_conc|> group_by(conc) |> summarise(total = n()) |> subset(conc == "Vertical equity", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Transportation related social exclusion (`r df_conc|> group_by(conc) |> summarise(total = n()) |> subset(conc == "TR exclusion", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Well-being (`r df_conc|> group_by(conc) |> summarise(total = n()) |> subset(conc == "Wellbeing", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Spatial equity (`r df_conc|> group_by(conc) |> summarise(total = n()) |> subset(conc == "Spatial equity", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Horizontal equity (`r df_conc|> group_by(conc) |> summarise(total = n()) |> subset(conc == "Horizontal equity", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Inequitable externalities (`r df_conc|> group_by(conc) |> summarise(total = n()) |> subset(conc == "Ineq. externalities", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Rights (`r df_conc|> group_by(conc) |> summarise(total = n()) |> subset(conc == "Rights", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Emerging theories (`r df_conc|> group_by(conc) |> summarise(total = n()) |> subset(conc == "Emerg. theories", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Utilitarian (`r df_conc|> group_by(conc) |> summarise(total = n()) |> subset(conc == "Utilitarian", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

A slice of the corpus is supported by broader "Rights" conceptualisations of fairness: these papers often focus on equity for people with disabilities or non-car users and associated challenges accessing transport infrastructure [@jimenez-espadaResearchProblemUniversal2021; @daamenAssessingGapPublic2008; @bharathyRevisitingClearFloor2018]. While many papers are underpinned with the right to the city (the _right_ to participate in the production of urban space [@lefebvreDroitVille1967]) [@adliRightCityApplying2018], others emphasize legal _Rights_ like ADA regulations [@bharathyRevisitingClearFloor2018] or the goal of _access for all_ in land-use transportation master plans [@limFacilitatingIndependentCommuting2021].

In another subset of the corpus, distributions are examined through concepts of "Horizontal" and "Spatial equity", often using quantitative methods to assess distributional disparities without explicit justice rationales. Examples include setting travel impedance thresholds [@shenMeasuringSpatialAccessibility2020] and mapping accessibility indices spatially across populations [@monzonEfficiencySpatialEquity2013] or population-groups [@sharmaPublicTransitAccessibility2021]. These papers may also address traffic-related air and noise pollution or urban temperatures. In these papers, equity is theoretically achieved if similar levels are attained for all populations (horizontal equity) or spatial areas (spatial equity). This egalitarian perspective rarely delves into minimum or maximum levels associated with harm or need satisfaction.

Papers that center "Well-being" assess what constitutes a satisfactory life in relation to transportation; mixed-methods are often used and, in contrast to the last set, the objects of injustice are often identified. Some papers use physical activity guidelines and surveys to understand the effect of active transportation infrastructure [@adlakhaMindGapGender2020; @auchinclossDesignBaselineDescription2020; @mccormackRelationshipClusteranalysisDerived2012]. Mixed or qualitative methods combined with health-related outcome standards (e.g., travel times for emergency treatment [@savingmothersDidSavingMothers2019], premature mortality [@awuorInfluenceEnvironmentalHealth2019]) lead to more concrete statements of fairness.

Another research branch, often quantitative with some qualitative or mixed-methods studies, focuses on "Transport-related social exclusion", "Vertical equity", and/or "Sufficientarian/capabilities". Here, the objects of justice are often clearly identified and link standards to tangible welfare-informed outcomes. They focus on groups from perspectives of disadvantage such as: social exclusion and transport poverty [@allenSizingTransportPoverty2019a; @delboscTransportProblemsThat2011; @churchillTransportPovertySubjective2019]], food deserts [@mckeyCrowdsourcedMappingHealthy2020], and energy poverty [@robinsonDoubleEnergyVulnerability2020; @berryInvestigatingFuelPoverty2016; @berryDistributionalEffectsCarbon2019].

### Standards and methods of measuring fairness

We identify several categories of standards that overlap with at least one of the conceptualisations of fairness discussed above (see @fig-fig2 for definitions):

- Opportunity standards (`r df_standard_mcat |> group_by(std_mcat) |> summarise(total = n()) |> subset(std_mcat == "Opportunity", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Population standards  (`r df_standard_mcat |> group_by(std_mcat) |> summarise(total = n()) |> subset(std_mcat == "Population", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Infrastructure standards (`r df_standard_mcat |> group_by(std_mcat) |> summarise(total = n()) |> subset(std_mcat == "Infrastructure", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

- Environment+ standards (`r df_standard_mcat |> group_by(std_mcat) |> summarise(total = n()) |> subset(std_mcat == "Environmental +", select = "total") |> mutate(total = total/165) |> pull(total)  |> percent()`)

Papers with "Opportunity" standards often employ quantitative methods to analyze disparities and assess distributional fairness. Many deal with travel impedance thresholds based on speed, distance, or cost [@chenImpactHighspeedRail2017; @yenisettyMeasuringAccessibilityVarious2020; @shenMeasuringSpatialAccessibility2020]. Inequality measures like the Gini coefficient and poverty measures are used to empirically define travel impedance thresholds [@vanderveenOperationalizingIndicatorSufficient2020; @tiznadoaitkenRoleAccessibilityPublic2018]. Further, methods tangential to travel impedance, like limiting transport expenditure to 10% of monthly income [@rivasHowAffordableTransportation2018], addressing spatial mismatch [@mulleyAreNetworkPlanning2015], or pinpointing areas of relative regional inequities are also used. Notably, many papers with "Opportunity" standards consider multiple dimensions, employing similar methods but tailored to different focal points. For instance @peungnumsaiGridBasedSpatialAnalysis2020 suggest service benchmarks of equal supply and demand of transit, revealing "Horizontal equities" as well. Others conceive the externalities of transportation system as trade-offs and aim to maximize transport-related benefits (i.e., time savings, emissions reductions, congestion reductions, user fares) through optimization/location-allocation methodologies [@fakhrmoosaviIncorporatingTravelTime2021; @zhengAreabasedEquitablePricing2020; @wismadiSpatialPreferenceModelling2014].

"Population" standards are often founded on "Well-being" conceptualisations from a variety of socio-demographic and spatial angles. Methods include: establishing thresholds based on questionnaires and comparisons to recommended physical activity levels [@auchinclossDesignBaselineDescription2020; @mccormackRelationshipClusteranalysisDerived2012; @kimNeighborhoodEnvironmentWalkability2016; @timperioSafetyNumbersDoes2015], region-relative comparisons in health outcomes in a spatial unit such as premature mortality rates [@awuorInfluenceEnvironmentalHealth2019], spatial access benchmarks based on population-related characteristics like supermarket access [@murphySupermarketAccessTransport2017] and hospital access @pereiraGeographicAccessCOVID192021], summative per capita benchmarks such as decent living energy consumption levels [@raoDecentLivingEmissions2012], and community-informed spatial boundaries like EJ defined communities [@rowangouldIdentifyingEnvironmentalJustice2016]. While most of these papers use quantitative or mixed-methods approaches, some employ exclusively qualitative methods [@berheAdaptationDissonanceQuality2014].

Papers that feature both "Population" and "Opportunity" standards are often founded on "Vertical equity", "Well-being", and/or "Transport-related social exclusion" conceptualisations. They feature mixed-methods, with questionnaires and qualitative approaches for "population" standards and quantitative methods like accessibility indices for "opportunity" standards. Census data and household estimates within specific travel distances or times to key destinations identify social exclusionary situations [@chenExploringTravelCharacteristics2010; @danielsProposalAccessibilityPlanning2011; @sunPublicTransportAvailability2021; @sharmaPublicTransitAccessibility2021], linkages between transport disadvantages [@delboscTransportProblemsThat2011], areas experiencing transport poverty [@allenSizingTransportPoverty2019a; @churchillTransportPovertySubjective2019], food deserts [@mckeyCrowdsourcedMappingHealthy2020], or transport-related energy poverty [@robinsonDoubleEnergyVulnerability2020; @berryInvestigatingFuelPoverty2016; @berryDistributionalEffectsCarbon2019]. They employ various methods, such as clustering techniques [@mohriClusteringMethodMeasuring2021]. Some exclusively use qualitative methods to analyze survey data on travel willingness/barriers or conduct interviews on unmet activity needs [@chenExploringTravelCharacteristics2010; @mehdizadehWalkingTimeSchool2017; @nordbakkeTransportUnmetActivity2015].

"Infrastructure" standards offer another perspective on fairness, commonly grounded in "Rights" conceptualisations, which are twice as frequent in this segment of the corpus compared to other conceptualisations. These papers most frequently address the rights of non-car users and populations with disabilities. Though various methods are applied, infrastructure and environmental audits as well as qualitative approaches are most prominent. Infrastructure audits compare existing infrastructure against universal design best practices [@odeckEconomicAppraisalUniversal2010; @larkinsAccessibleTransportationBuilt2011; @jimenez-espadaResearchProblemUniversal2021; @perezdelhoyoParticipatoryManagementImprove2021] or investigate elements correlating with mode use by specific population groups [@moniruzzamanInvestigation2016]. Qualitative methods include interviews/surveys on perceived access [@marquezFactorsAffectingPersonal2019; @mateo-babianoBicycleSharingAsia2017; @furstBestBadPractices2013; @velhoEffectTransportAccessibility2016; @parkJourneyVisuallyImpaired2017; @limFacilitatingIndependentCommuting2021; @stjernborgAccessibilityAllPublic2019; @desjardinsGoing2021] and assessment of standards under best-practice criteria [@daamenAssessingGapPublic2008; @velhoEffectTransportAccessibility2016; @bharathyRevisitingClearFloor2018].

Additionally, "Infrastructure" standards papers sometimes encompass multiple dimensions, moving beyond rights (to the infrastructure) to provide "Opportunity" and/or "Population" standards. These papers often employ "Vertical", "Horizontal", and "Spatial equity" lenses. These papers often refer to guidelines and propose composite indices. For example, [@racheleNeighbourhoodSocioeconomicTransport2017a] integrate various transport network properties to define an indicator supporting walkability and public transport access. Others evaluate infrastructure quality [@xuSocialInequalitiesPark2017], accident severity [@benevenutoExaminingTransportNeeds2020; @appleyardTransitCorridorLivability2017], and user-groups, especially disadvantaged ones with respect to vertical equity of multi-criteria indicators [@prasertsubpakijEvaluatingAccessibilityBangkok2012]. Some focus explicitly on affordability and barriers, proposing infrastructure enhancements for better inclusivity, especially for the most disadvantaged [@basuPracticalFrameworkBenchmarking2019; @songWomenWheelsGender2019; @welchEquityTransportDistribution2013], grounded in transport-related social exclusion [@kentPrioritizingLowstressEquitable2019] or capabilities approaches [@smithAccessibilityCapabilityMinimum2012].

"Environmental+" standards feature less prominently in our corpus, with studies that predominantly examine "inequitable externalities", possibly because the environmental burdens of transportation are addressed more broadly in other literature (e.g., environmental justice). However, the papers included in our corpus present some interesting insights. They often use traffic-related air and noise pollution, green space, urban design, urban air temperature, health outcomes, and physical activity guidelines to assess transport-related externalities. Methods employed are primarily quantitative or mixed-methods, identifying inequalities through Gini coefficients [@fengTradeoffsMobilityEquity2014] or composite indices [@agostfelipInclusiveModelAssessing2021; @mirandaBenchmarkingSustainableUrban2012; @corazzaMethodologyEvidenceCase2020], occasionally incorporating spatial analysis [@jephcoteGeospatialAnalysisNaturally2013; @carrierApplicationThreeMethods2014]. Many use established thresholds or health guidelines e.g., WHO guidelines for Active Aging and targets included in the United Nations’ SDG [@agostfelipInclusiveModelAssessing2021] and OECD or WHO standards for traffic noise level, $NO_2$ levels, available green space and $PM_{2.5}$ levels [@kruizeEnvironmentalEquityRole2007; @iungmanImpactUrbanTransport2021; @apparicioCyclingOneMost2021; @khomenkoLiveableCityHealthy2020; @muellerSocioeconomicInequalitiesUrban2018]. These metrics are sometimes criticized for their general applicability. Nonetheless, their use provides interpretable values for tracking progress, offering comparability across communities, unlike accessibility measures, which vary in methods and assumptions.

# Moving forward: calls for action {#sect4}

This paper makes two contributions to the literature. It outlines a conceptual and flexible "Why", Where", "When", "Who", "What" and "How" (5WH) framework with supporting definitions for approaching questions about transportation justice. It also applies the framework in the synthesis and appraisal of 165 academic articles drawn from the transportation justice literature; these studies use equity analysis as instruments to gauge proximity to their context's just situation.

In brief, justice is an end objective and standards are definitive statements of fairness. As such, they are essential to assess proximity to the objective. Context-specific evidence of transport inequities is necessary to acknowledge injustices and inspire creative interventions to advance the cause of justice. Armed with appropriate knowledge, policy can be designed to address inequities while remaining adaptable to the evolving demands of justice. 
Flexible frameworks are crucial. What is (in)formally considered (in)essential today, could change during the next pandemic or technological revolution. Creating adaptable transport justice frameworks that lend themselves to change is required to ensure resiliency to backsliding. From the diverse corpus reviewed in this paper, we argue for the necessity of setting standards from a systems approach, considering both _positive_ rights (e.g., a right to have access to sufficient quality of essential services) and _negative_ rights (e.g., mobility of cars must be limited to reduce their impact on air quality, health, safety, and other positive rights).

We conclude with the following five calls to action.

## Call 1: Develop standards based on rigorous concepts of justice {#sect4.1}

The conceptual grounding for standards is often left implicit within the literature: for example, @muellerSocioeconomicInequalitiesUrban2018 suggests the relative risk of mortality from transport-related air pollution should not be higher in deprived groups than the general population. While the conceptual justice foundations are not explicitly declared, we infer an egalitarian focus. But, what level of relative risk is acceptable for the general population? What purpose does mortality risk serve and who benefits from it? To move equity analysis outputs towards just transport futures, explicit inclusion of justice rationales (the "Why?") should become more common place. We must be clear with our terms--what is equity and for whom?

Standards are statements of fairness, yet some standards are seemingly arbitrary in the reviewed literature--in other words, set arbitrary justice goals. For example, @caoIndicatorsSocioSpatialTransport2017 proposes 20 ferries per day to avoid social exclusion for inter-island transport planning, though acknowledging the standard should be politically determined. It is unclear if selecting a different benchmark such as 10 or 30 ferries would make a difference in any specific object of justice (e.g., accessibility to particular destinations) or if that number is tied to funding or resource constraints. Another example is the conceptual underpinnings of the 15-minute city, where 15 minutes is the standard for travel times. Is this sufficientarism? Or egalitarianism? The standard can be interpreted in a multitude of ways. We argue that conceptualization of justice must guide the selection of the standard. As @martensFair2021 demonstrate, being explicit about "Why" is important. Recent work by @karnerAdvancesPitfallsMeasuring2024 provides several concrete examples for developing standards that follow directly from different justice conceptualizations.

For decision-makers, setting standards, measuring inequities, and developing flexible guidelines for standards that are also compatible with community-informed calls for justice, are the next steps for transportation equity planning. If we assume that one function of the academic literature is to recommend standards, researchers must connect compatible conceptualisations and standards to justice frameworks. Currently, these fundamental connections are mostly missing.

## Call 2: Develop creative methods for systems-thinking approaches to fairness {#sect4.2}

On the methodological side, transportation justice research would benefit from wider use of mixed-methods. Concepts and standards are often discussed from purely qualitative or quantitative approaches. This is a missed opportunity to combine the strengths of both approaches, whether by deep diving into some particular experiences or perceptions through qualitative methods or tailoring more meaningful quantitative analysis after qualitative explorations. As mixed-method examples, @xieWonCycleRoute2018 find through interviews and go-alongs with women cyclists that the standard Cycling Level of Service (CLS) tools used by engineers to plan cycling infrastructure misses a critical gendered perspective. Further, @somenahalliAgingTransportMobility2007 survey older adults to understand their mobility issues, revealing factors that are unseen in standard daily travel surveys.

While disparity analysis is frequently used, the resulting standards do not often align with practical applications. For example, metrics of accessibility (usually measured with travel impedance cut-offs of between 15 to 60 minutes depending on the destination, population group and mode) are used to show descriptive differences among areas and groups but with scarce implications as to the experience of travelers. How low must an accessibility index be before it is too low? Rellatedly, once sufficient accessibility levels are reached, excessively high accessibility values can result in high inequalities. Are inequalities of a certain level an issue? A lack of discussion poses a challenge to translating results from these analyses into policy and practice. Creative methods and discussions of quantitative accessibility metrics should be paired with results; they should yield interpretable results. The explicit discussion of minimum _and_ maximum values in the distribution of the object of justice (the "What?"), as applicable, is critical.

Other times, when analysis engages with metrics that may be tied to particular concepts of equity (like the Gini coefficient or Theil index), they fall short in assessing whether the results are good or bad [@mijaresEquityAnalysisUrban2013]. For example, a Gini coefficient of 0 would mean that all people have the same access to public transport stops. But is this level of access good enough or lacking? And what does it mean if the result is 0.2, 0.3, or 0.4? Is this good or bad news? How should a decision-maker interpret this? Are new policies needed to reduce that number to a certain threshold, orienting future interventions? These questions usually remain unanswered in the literature despite their importance. These measurements can also bring some challenges and pitfalls as summarized by @karnerAdvancesPitfallsMeasuring2024 but are necessary for more effective equity analysis.

## Call 3: Making data available is a matter of justice {#sect4.3}

In our review of the literature, we were left wondering specifically--what are the motivations for the use of some destinations and the choice to not include others? For instance, leisure destinations (e.g., green space, parks, recreation) are infrequently studied and some categories are missing altogether. We suspect that since methods used are predominately quantitative, the reliance on commonly used point of interest databases is also high. These databases typically include education, health and occasionally include aggregated categories for leisure and community, but are less generous in terms of 'community' and 'leisure' points of interest (community organizations, government services, care-giving destinations). Further, categories like childcare are often missing, typically daycare or facilities--domestic work, mobilities of care, and mobility interdependence. Not coincidentally, these categories are underrepresented compared to work and healthcare destinations.

We can only track what we have collected and compiled; and we know that transport systems are more than just facilitators to work or as a source of economic development (though in underdeveloped regions, transport systems as a force of economic development is pronounced, e.g., high-speed rail [@chenImpactHighspeedRail2017; @monzonEfficiencySpatialEquity2013; @kimImpactsHighspeedRail2015]). As such, data availability matters. In the context of the Global South and rural geographies fewer official data sources and public research resources exist relative to urban communities and areas in the Global North. As well as in the operationalization of emerging conceptual theories in equity analysis, such as sufficientariansm [@vanderveenOperationalizingIndicatorSufficient2020].

The calls for and relevant issues of data availability are not new, but they have at least three parts. What and who is the subject of justice? When/Where/How is it measured? And, who gets to consult and use the data? Deciding who is the subject of justice frames the data collection activity; if it is the mobility of those that do domestic work, the classification of who does this work and how it changes over time/space is fundamental. How we classify mobility and access has implications for our understanding of just who is a subject of justice, as illustrated by the history of racial classifications in the US [@leeRacial1993]. The methods we use, the spatio-temporal boundaries for data collection, and who has (and does not have) access to the collected data, can all impact how we think about justice. In the case of transportation, the issue of data collection/availability as a matter of justice is gaining traction as digitization of data casts a starker light on these questions [@sourbatiSmart2021; @behrendtData2023].

## Call 4: Develop more direct and explicit links between standards and lived experiences {#sect4.4}

Robust assessments of the implications of equity standards on lived experiences are still lacking in the literature, but are essential for equity analysis that translates into just practice. While the estimation of the benefits of increased mobility or accessibility, or reducing burdens is commonplace in the literature, there is a need to associate them to outcomes, including life and neighbourhood satisfaction, subjective well-being, mental and physical health, and social capital.

Composite measures such as the transport-land-use index proposed by @appleyardTransitCorridorLivability2017 is an example of systems-thinking approach that links findings to quality-of-life proxies; they ground their measure in the principle of _livability_ along corridors of varying levels of estimated transport-land-use integration. However, the methods could go further; they could be tied to absolute goals of integration or livability. Relatably, @higgsUrbanLiveabilityIndex2019 develops an urban livability index and demonstrates the relationship between the population's use of a certain mode in a neighbourhood and a one unit increase in the index. However, are there absolute minimums or maximums for the index or the mode-choice goals that should not be crossed?

More explicit discussion of the boundaries in the distribution of the object of justice (the "What?") alongside these creative methods are needed. These links may be used to track progress towards justice across time and space, a critical point for practitioners.

## Call 5: Rigorously evaluate interventions and policies {#sect4.5}

There is a need to evaluate equity interventions and policies: track their before, during, and after. In this review, barely 10% of studies assess specific transport-related policy interventions through an equity lens. Examples include mode-shift from driving to active school travel [@mammenSchoolTravelPlanning2014], transit fare restructures [@hickeyUsingQuantitativeMethods2010] and spatial analysis of Low Traffic neighbourhoods [@aldredEquityNewActive2021]. The assessment of interventions and tracking associated outcomes can be thought of as a key step towards transport justice. Similarly, for how long the burdens and benefits can still be associated to a specific transportation intervention remains a critical avenue to explore.

Outcomes of interventions can be compared within and between communities, and cross community comparisons can be created that may expedite the adoption of effective policies that move towards just outcomes; the presence of these synthesis and comparative studies could support brave decision-makers in the application of research into practice.

# Declarations

During the preparation of this work, the first and second authors used ChatGPT exclusively to reduce word count in the section "An appraisal of the lay of the land" while keeping the original writing style. After using this tool, all authors reviewed the manuscript and take full responsibility for the content of the publication.

\newpage

# References

::: {#refs}
:::

# Appendix {#sec-sect6}

<!-- ## Why? A framework extension to analyse questions of justice {#sec-sect61} -->

<!-- To give an example drawing from the Canadian Charter of Rights and Freedoms, a number of rights and fundamental freedoms [including the liberty to move freely; see @chartopediaArt72023] are recognized to exist in Canada "without discrimination by reason of race, national origin, colour, religion or sex". Notice that this declaration includes several individual traits that define the subjects of justice, but does not consider age as one aspect of the person. Does this mean that they apply universally irrespective of age? Surely not, since no reasonable person would consider a toddler's demands for freedom of association as absolute. So, when do these rights fully apply in the life of a person? In 2017, Adrian Crook of Vancouver, B.C., was warned by the province's Ministry of Children and Family Development that his kids could not be out of home in the community, alone or with other kids the same age, without supervision. The Ministry's argument was that children riding the bus unsupervised compromised the parent's ability to provide care and placed them at risk [@brendHere2017]. Crook argued that the goal of teaching his kids (then aged 7, 8, 9 and 11) to ride transit to school was to raise self-reliant children [@stueckYoung2017]. While both sides discovered to everyone's surprise that no clear rules existed regarding children's independent transit use [@brendHere2017], statistically there is evidence that travel by car is riskier than travel by bus [@morencyTraveling2018]. If the public good is to reduce risk, shouldn't children be banned from riding in cars? The Canadian legal system concerns itself with fundamental freedoms and rights, one of which is the liberty to move freely. This provides an answer (of many) as to "why" we would even consider children as subjects of justice. But the devil is in the details, and answering the other questions discussed above is essential to pin the devil by the tail. Are children under the age of 10 subjects of justice (i.e., are they a "who") and who decides if they are (i.e., the courts or the parents in this case)? If so, is the ability to use transit unsupervised an object of justice that should be justly distributed (i.e., the "what") and under what circumstances (e.g, to school, to a social event; during the day, or late at night?). "When" (e.g., for all children going forward, for only 5 years) and "where" (e.g., only in the suburbs, across the country) do these questions apply? Eventually, in 2020, a court of appeals ruled in Crook's favor, allowing his children to continue to ride the bus, thus providing much needed clarity with respect to the use of transit by children who desire and are encouraged to exercise independent mobility [@stueckCourt2020]. In actuality, the courts _created_ a standard of equity: starting at a certain age, children are objects of justice from the perspective of freedom of movement, and they are due the benefits of independent mobility by transit.  -->

<!-- Another perspective on the answer to "Why?" are cultural _norms_ that provide a rationale. In this case, the liberty of movement for all---including children---reigns supreme. But a competing concern is that the outside is dangerous, and it is so in part due to vehicular traffic [@morencyTraveling2018], and in part because of potential violence from others. Parents may have over-pronounced 'stranger danger' based on neighbourhood and socio-economic characteristics of neighbourhoods [@francisLlBeDriving2017] and children themselves may fear potential violence based on real threats of victimization [@painParanoidParentingRematerializing2006]. Whose voice, what perception of fear, and actions we take to intervene define and reproduce cultural norms, all play into the rationale behind the question "Why?".  -->

## Evidence search strategy {#sect5-2}

```{r PRISMA-figA1}
#| label: fig-figA1
#| fig-cap: "PRISMA flow diagram for the evidence selection process. ES signifies equity standard and EC signifies equity conceptualization."
knitr::include_graphics("figures/Pflow_chart.tiff")
```

The evidence selection process is also represented using the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) flow diagram [@pagePRISMA2020Statement2021] in @fig-figA1. Notably, two rounds of exclusion occurred during the assessment for full-text eligibility. 1710 studies entered step 2, 1223 were excluded and the remaining 487 papers entered step 3. The data extraction template used by the reviewers (authorship team) in step 3 revealed that, as expected, inclusion was initially too generous, and some papers were not sufficiently relevant, because of a lack of content on standards and/or conceptual/theoretical elements. In this fashion, 322 papers were further excluded and data extraction was completed to give a final corpus of 165 papers. A summary of the reasons for exclusion of the 1545 papers (between steps 2 and 3) are included in @fig-figA1.

```{r Search-query-A2}
#| label: fig-figA2
#| fig-cap: "The search query. TS = topic search (keywords, abstract, title). TASCA = subject categories. Green text area transportation system related terms, blue text are equity dimension related terms, purple text are equity/justice conceptualization related terms, and orange text are standards related terms. Hits corresponds the number of papers that the search yielded and was retained into the evidence selection process."
knitr::include_graphics("figures/Search-query.png")
```

Definitions of the population-concept context (PCC) used in the creation of the inclusion and exclusion criteria for the search strategy.

-   **Population**: the focus of the included studies should be on individuals, groups, communities, or entire regional areas that are impacted by passenger transportation infrastructure and systems (i.e., all modes and flows) from the perspective of equity (i.e., fair distribution, production, and re-production of burdens and benefits). This criteria is reflected in the creation of the first set of topic search terms that relate to transportation modes (e.g., "walking" OR "cycling" OR "transit" - see green text in @fig-figA2 for the full list).
-   **Concept**: the included studies should also include equity dimensions and conceptualizes equity as discussed in the previous section. This inclusion criteria is reflected in the second and third set of topic search terms developed in the search strategy. These terms relate to types of equity dimensions (e.g., "accessibility" OR "mobility" or "transport-related air pollution" - see blue text in the @fig-figA2 for the full list) and equity conceptualizations (e.g., "Justice" OR "equity" - see purple text in @fig-figA2 for the full list).
-   **Context**: the included studies should also be limited to publications that include equity standards. Context can be more difficult to explicitly search for with key terms so synonyms for 'standards' were added to the query as a four set of topic search terms (e.g., threshold, indicator, criteria - see orange text in @fig-figA1 for full list). Additionally, journal article and conference papers, English-language literature from any country, any study design (e.g., quantitative, qualitative, or mixed-method studies, or conceptual frameworks), and any record published within the past 30 years are included (January 1992 to March 2022). The time period is selected as the first (to the authors knowledge) peer-reviewed article which operationalized equity standards and equity conceptualization was published in 1996 [@khistyOperationalizingConceptsEquity1996]; we are broadening the search by a few years for completeness. English is selected as it is the common language spoken across the authorship team. Furthermore, papers that explicitly fall within the Transportation or related topic/category is included in the query (e.g., "Transportation", "Social Sciences", "Geography", "Civil Engineering", "Philosophy" - see the @fig-figA2 for full query).

The **exclusion criteria** for the search are papers that are not within the inclusion criteria.
Specifically:

-   Literature published before January 1992.
-   Papers which do not include transportation equity dimensions.
-   Grey literature, as concepts contained within are frequently published in a more developed form in journals.

## Evidence selection and data extraction {#sect5-3}

```{r methods_framework-fig1}
#| label: fig-figA3
#| fig-cap: "Evidence selection workflow. Step 1 (orange) is title and abstract screening, step 2 (green) is full-text review, and step 3 (purple) is data extraction."
knitr::include_graphics("figures/Methods_framework.png")
```

The following steps summarize the evidence selection process:

1. The first step (orange box in @fig-figA3) included screening all titles and abstracts of papers on whether they included transportation equity as defined by the PCC. Each paper was screened by two independent reviewers who then voted for inclusion, exclusion, or uncertain inclusion. All uncertain papers, conflicting papers, and papers missing abstracts were reviewed by a third person for inclusion or exclusion.

2. The second step (green box in @fig-figA3) included scanning all full-text papers which passed step 1. These papers were reviewed to determine if they included a relevant "how", i.e., an standard and/or relevant theoretical or conceptual discussion. At this stage, papers were evaluated again by two independent reviewers who voted for inclusion or exclusion. If an article was voted to be excluded, it was tagged with one of five possible reasons for exclusion, namely (1) no standards included; (2) no relevant conceptual elements included; (3) no standard and no conceptual elements included; (4) send back -- QA issue; or (5) other. Discrepancies were resolved by a third reviewer.

3. In the last step, a data extraction template for each record was filled by one reviewer (purple box in @fig-figA3). The data extraction template was created with the aim of striking a balance between the complexity of categories and the simplicity of summary; information related to "Where?" (the geographical context and sphere of life), "When?" (temporal circumstances for the application of justice), "Who?" (the subject of justice), "What?" (the object of justice), and "How?" (equity standard(s) and concepts) was filled out for each study. The following table contains the template that was input into *Covidence* and used throughout.

Data extraction for each document that passed through all three steps was then extracted using this template:
```{r data-extraction-template-A4}
#| label: fig-figA4
#| fig-cap: "The data extraction template with associated defintions."
knitr::include_graphics("figures/Data-extract-template.png")
```

:::landscape

To provide further example of data extraction, a table of the resulting literature looked like:
```{r}
dat <- data.frame(matrix(c("What?: **mobility and accessibilty**",
                           "Where? [@rivasHowAffordableTransportation2018] - South America (select cities)",
                           "Analyses how affordable urban public transportation is in select Latin American and Caribbean countries. They look at the estimated average monthly cost of transit trips and average monthly household income and conceptualize **transportation-related** **affordability**, especially for the most economically vulnerable (**vertical equity**).",
                           "How?: The financial burden of a basket of urban public transportation trips (60 trip fares, representing 30 round-trips per month) should not exceed 10% of household monthly income.",
                           "What?: **mobility and accessibilty**",
                           "Where?: [@bharathyRevisitingClearFloor2018] - North America (USA - National)",
                           "This study designed a web-based tool and took a representative sample of wheeled mobility device (WhMD) users anthropometry measurements to determine if the minimum standard suggested by the ADA is sufficient. We understand this conceptualization as a type of **Rights** conceptualization that WhMD should have minimum clear floor space (as described the guidelines in line with the American Disabilities Act) to access bus shelters, bus stop pads, and transit terminals.",
                           "How?: The clear floor area for wheelchairs: 760 mm (30 in.) wide by 1220 mm (48 in.) in length as described by the ADA standards. Of note, this minimum clear floor area is insufficient for a variety of the WhMD users.",
                           "What?: **mobility and accessibilty**",
                           "[@ryanWhatAreWe2021] - Europe (Stockholm, Gothenburg and Malmo cities in Sweden)",
                           "Investigates what the literature and planning process is missing when we measure accessibility by comparing objective and self-reported accounts of accessibility among older people. This paper conceptualizes accessibility as from the position of the **capabilities approach** and **vertical equity** (particularly acknowledging that older people have capabilities that differ from the general population).",
                           "How?: Specifically for older populations (aged 65+), the following travel distances are suggested as equitable trip lengths to grocery stores per mode: Walking: less than or equal to 1500m, Combined transit and walking (less than or equal to 1000m (walking element)), Combined car and walking: less than or equal to 1000m less than or equal to 1000m (walking element)), Bicycle: less than or equal to 3000m in addition to travel time threshold of less than 15 mins.",
                           "What?: **mobility and accessibilty**",
                           "Where?: [@wismadiSpatialPreferenceModelling2014] - Asia (Yogyakarta, Indonesia)",
                           "Explores the equitable provision of transport infrastructure provision: an application of Sen's **capability approach**. Conceptualizes equity through Sen's capability approach and spatial equity.",
                           "How?: Areas below the relative poverty line (of its neighbours) can only be located transport resources (i.e., measure in person*kms that can be travelled at car speed, i.e., mobility) based on the following 2 benchmarks (they can be considered, together as the floor/minmum access): 1) Global: standard deviation (SD) distance to mean should be minimized. 2) Local: priority to minimise the differences with its neighbourhood",
                           "What?: **mobility and accessibilty**",
                           "Where?: [@zhengAreabasedEquitablePricing2020] - North America",
                           "This paper conceptualizies equity in the multimodal network (transit, car) being fair toll-pricing across differences in populatins value of time (VOT). VOT is determined based on household income, with lower income households having lower VOT and thus deserving of lower tolls (vertical equity). From this perspective, a utilitarian perspective that seeks to minimize multimodal traffic congestion through introducing toll-pricing based on VOT is implemented.",
                           "How?: suggest that a toll-pricing scheme based on individuals travel value-of-time (lower income people have a lower VOT) is equitable.",

                           "What?: **environmental pollution**",
                           "Where?: [@carrierApplicationThreeMethods2014] - North America (Montreal, Canada) ",
                           "This work examines the statistical association between different social groups and the concentration of air pollutants. They frame their work from the perspective of environmental equity. We interpret the conceptualizations to be along the lines of **inequitable externalities**, **spatial** and **vertical equity** - transport-related air pollution is a product of road transport and it impacts the air of residents in unequal spatial ways. The paper then frames this impact as unfair, particularly from the perspective of disproportionately disadvantaged residents",
                           "How?: The literature suggests that the health implications from the transport-related air pollution from major roadways is most acute at residential distance locations of 200 m or less. Residential locations should not be located within this distance threshold from the perspective of human health. **Environmental+** and **Population standards**: Uses the WHO NO² threshold as a point of comparison (annual concentrations of NO² should not exceed 40 μg/m-3). They argue that even through no neighbourhood, even those disproportionately low income, exceed the WHO limit in this case study, they still suggest that air pollution should not be disproportionately impacting disadvantaged neighbourhoods. It can be interpreted that they use the WHO threshold as a minimum threshold and suggest that air pollution levels should not be impacting disadvantaged populations disproportionately ( a relative population standard)",
                           "What?: **environmental pollution**",
                           "Where?: [@jephcoteGeospatialAnalysisNaturally2013] - Europe (Leicester, UK)",
                           "Geospatial analysis of naturally occurring boundaries in road-transport emissions and childrens respiratory health across a demographically diverse cityscape. Emperically identifies at what distance away from major roadways children are most impacted by transport-related pollution. This is framed in the perspective of children’s **well-being**. Children are at most risk for acute respiratory distress from elevated levels of air pollution, and as such planning should consider this point of public health.",
                           "How?: Finds that children (most vulnerable to air pollution - related to motoized traffic) are most impacted by air pollution within 283 m of a road way. This should be the distance threshold that schools and other childrens facilities are located.",
                           "What?: **health impacts**",
                           "Where? [@adlakhaMindGapGender2020] - Asia (Chennai, India)",
                           "From the perspective of disparity in gendered physical activity, this paper focuses on women's cycling as both transport and exercise. They advocate for all people achieving physical activity thresholds (**horizontal equity**) but prioritize women and especially women in neighbourhoods with low-walkability and socio-economic status (**vertical equity**).",
                           "How?: All people should get 150 min of moderate activity a week or 75 min of vigorous physical activity per week.",
                           "What?: **health impacts**",
                           "Where? [@savingmothersDidSavingMothers2019] - Africa (Select urban and rural regions in Uganda)",
                           "The **well-being** of mothers, this paper examines the timely access to emergency obsteric and newborn care for child-bearing aged women in Uganda.",
                           "How?: 2 hours to the nearest facility with surgical capacity with anesthesia services - this threshold is determined through the onset of bleeding to death if a women with obstetric hemorrhage does not receive adequate treatment).",
                           "What?: **health impacts**",
                           "Where?: [@iungmanImpactUrbanTransport2021] - Europe (Madrid and Barcelona, Spain)",
                           "They use environmental pollution guidelines, but from the position of health. They investigate the impact of urban and transport planning on attributable mortality burden in Madrid and Barcelona and its distribution by socioeconomic status . Pre-mature mortality is linked to the exposure to pollution and motorized vehicles (**inequitable externalities**). These externalities should not be impacting people disproportionately (**vertical equity**) and should be even across space (**spatial equity**).",
                           "How?: All minimum thresholds, if exceeded this is inequitable: NO² concentration 40 μg/m³; PM 2.5 concentration 10 μg/m³; Noise 53dB for average 24 hours; Living with 300 m crow-flies distance from at least .5 hectares of greenspace; and a Change of air temperature of at least 1 ⁰C. ",
                           "What?: **health impacts**",
                           "Where?: [@mehdizadehWalkingTimeSchool2017] - Asia (Rasht, Iran)",
                           "From the perspective of children’s **well-being**, assesses the walking time to school. They frame walking to school as health-related.",
                           "How?: perceived walking time to school for students aged 7-9 yrs is 10 mins, and the longer the PWTS the less likely they were to use an active mode to travel to school.",
                           "What?: **health impacts**",
                           "Where? [@murphySupermarketAccessTransport2017] - Oceania (Melbourne, Australia)",
                           "Assesses the relationship between supermarket access and transport mode used, the body mass index (BMI) of the mode-user (**wellbeing**) and the equity in access distribution by income (**vertical equity**).",
                           "How?: all households should be sufficiently active (greater than 150 min and at least 5 sessions) and households should be within 1 km euclidean distance to supermarket (80-90% of the dwellings should meet this). Planners should prioritize socially disadvantaged areas to meeting these standards first.",

                           "What?: **transport-related safety**",
                           "Where?: [@ferenchakEquityAnalysisProactively2019] - North America (Denver, USA)",
                           "Operationalizes and compares an equity analysis of proactively- and reactively-identified traffic safety issues from the perspective of **Spatial equity**, **Vertical equity** and **Inequitable exposure to externalities**.",
                           "How?: standards are suggested for both reactive and proactive analysis. First, the lower the number of collisions on the road with pedestrians/cyclists (i.e., reactive safety analysis), the better. No/minimal inequalities for general population vs. equity seeking groups (high proportion of POC and/or low income in tract). Second, the lower the perceived safety, the better (i.e., if travel to school by ped. or bike is unsafe due to traffic conditions). No/minimal inequalities for general population vs. equity seeking groups (high proportion of POC and/or low income in tract).",
                           "What?: **transport-related safety**",
                           "Where?: [@zheEvaluationSharedUse2008] - Asia (Tokyo, Takamatsu, and Tokushima)",
                           "Evaluates the observed safety of  shared use pedestrian and bicycle paths from the perspective of **well-being**.",
                           "How?: the study suggests that the safety threshold for bicycles and pedestrians to coexist on shared infrastructure is less than 0.5 pedestrians/minute per metre of sidewalk (width) and less than 3.0 cyclists/minute per metre of sidewalk (width). The standard for pedestrian/bicycle share use in terms of hourly traffic volume is less than 26 pedestrians / hour and 108 cyclists / hour for 2m wide sidewalks.",

                           "What?: **mobility/accessibility and health impacts**",
                           "Where?: [@aldertonWhatMeaningUrban2019] - Asia (Bangkok, Thailand) – **Mobility/ accessibility** and **health**",
                           "Establishes short-, medium-, and long-term goals for the city in collaboration with technical leaders within the municipal government for the perspective of **well-being** (urban livability): the standards included in this table relate directly to transportation systems. Indicators are inspired by the Sustainable Development Goals (SDGs) as well other global planning standards.",
                           "How?: 1) Green space: % of residents living < 400 m from public open space, a large park (> 1.5ha), and/or local park, 2) transit access: % of residents living < 400 m of a local bus stop and <800 m of train station, 3) Facilities: % of residents living < 400 m of a community centre. The following **Infrastructure standard** is suggested: Canal water quality - dissolved oxygen content of equal to or less than 2.0 mL/L",
                           "What?: **Mobility/ accessibility** and **health impacts**",
                           "Where?: [@berheAdaptationDissonanceQuality2014] - Africa (Mekelle, Ethiopia)",
                         "Examines adaption and dissonance in the quality of life (QoL) of residents. QoL is conceptualized along the lines of **well-being** and aspects of QoL directly tie into transport systems. They conduct a qualitative QoL survey of residents on the topic of three QoL domains: housing quality, access to important destinations, and affordability. They also measure quantitative indicators associated with these domains. We assume the equity goal for this paper is that subjective and objective QoL measures should not be mismatched: as discussed by the authors of this study, subjective QoL is higher than objective QoL the participant is experiencing adaption and in the reverse scenario the participant is experience dissonance.",
                           "How: 1 & 2) Access to primary or secondary education facility, percentage of households living within 1 km or 2km (walking distance), respectively from a primary school or secondary school. 3) Access to health facility, percentage of households within 40 min walking time from a health facility. 4) Access to public transport, percentage of households within a distance of 500 m from a mini-bus stop. **Population standards**: 1) Adequate family income, percentage of households earning more than the official poverty line. 2) Subjective QoL is constructed based on the households level of satisfaction for each of the eight indicators using a six point Likert-scale (1=very satisfied to 6=very dissatisfied).",
                           "What?: **mobility/accessibility, health impacts, and safety**",
                           "Where? [@agostfelipInclusiveModelAssessing2021] - Europe (Castellon, Spain)",
                           "Conceptualizes equity through age-friendly urban spaces that reduce (and eliminate) conditions for **transportation-related social exclusion** for older populations and prioritize those who are economically vulnerable (**vertical equity**). These guidelines are inspired by the SDGs in addition to planning guidelines used national, regional, and local guidelines used in Spain.",
                           "How: 1) Access to facilities needed for old age health. Minimum distance thresholds from the geometric center of neighbourhood are suggested: at least: 1000 m from health facilities (600 m or less is preferred), elderly-specific care facilities and shops should be 600 m (300 m or less is preferred). **Population standards**: 1) Certain neighbourhoods should be prioritized above others. From this papers focus on age-friendly urban environment, they suggest that if the neighbourhood has an average old age indicator (i.e., greater than 64 years, and/or greater than 79 years, and/or aging ratio of persons aged greater than 64 relative to 15 to 64 age) should be prioritized. 2) Economic vulnerable and non-civically engaged neighbourhoods should also be prioritized. If the neighbourhood has a lower percentage of civic associations within the neighbourhood than average, and/or household income, and/or a higher than average interventions for dependency and/or social subsidies, they should be priorized. **Infrastructure standards **: 1) Green space: should be at least 10 m2 per inhabitant in the neighbourhood, greater than 15 m2 per inhab. is the goal. 2) As related to sidewalk infrastructure at least 50% of all sidewalks (preferably 75% or greater) should: have a width of 1.5m or larger, ramps should have a grade of 8% or less, be well maintained (free from deficiencies), be paved for pedestrian use, and cover transit stops. 3) Lighting is critical for traffic-safety and a sense of safety overall. As such, at least 50% roads should: have a min. of 35 lux (road traffic) and 20 lux (pedestrian streets), and adapted traffic lights. 4) Buildings should be age-friendly. As a proxy for the quality of residential living space quality, at least 50% of residential buildings in a neighbourhood should be built within the last 50 years (preferably 75% or more). In terms of physical access into the buildings, at least 10% should have elevators and accessible entrances (preferably 25% or more). **Environment + standards **: 1) Noise at the street level should be less than 55 dB and 45 dB (but preferably less than 50 dB and 40 dB) in the daytime and nighttime, respectively.",
                           "What?: **mobility/accessibility and safety**",
                           "Where?: [@mateobabianoPedestrianNeedsMatter2016] - Asian (Manila, Philippine)",
                           "The perception of pedestrians’ walking environments should be sufficient across 6 themes. Equity is conceptualized around **spatial equity** (equally fair walking environments for all locations) and **rights** (the right to mobility/accessibility for pedestrians)",
                           "How?: perceived pedestrian perception on protection, ease, equitable access, mobility, identity, and enjoyment must be met."
                           ), ncol = 4, byrow = T))
```

```{r}
dat |>
  gt() |>
  tab_options(table.font.size = "xx-small") |>
  cols_label(X1 = "Dimension",
             X2 = "Continent",
             X3 = "Conceptualization",
             X4 = "Standard") |>
  tab_caption("Examples of Data Extraction")
```
:::